---
title: "Investigate data"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library, echo = FALSE}
library(tidyverse)
library(labelled)
library(gtsummary)
library(survminer)
library(survival)
library(car)
theme_gtsummary_compact()
theme_set(theme_classic())
```

```{r load}
path_raw <- fs::path("","Volumes","Peres_Research", "AACES")

spatial_data <- 
  read_rds(paste0(here::here(), 
                  "/data/raw_data",
                  # path_raw, "/Panel 2 Immune Markers/Survival Analysis",
                  "/compartment_spatial_dataset_06Mar2026.rds")) %>% 
  janitor::clean_names()
# immune_subtypes <- 
#   read_csv(paste0(path_raw, 
#                   "/Panel 2 Immune Markers/Clinical data",
#                   "/panel2_bothphases_11072024.csv")) %>% 
#   select(suid, subtype)
# race_dat <- 
#   read_csv(paste0(path_raw, 
#                   "/AACES_NCOCS_data/clinical data/processed_data",
#                   "/case_ctrl_data_2024-08-16.csv")) %>% 
#   select(suid, site, race)
```

```{r fix grade}
spatial_data <- spatial_data %>% 
  # collapse histotype/grade, assume missing grade is HIGH
  mutate(histo_grade = case_when(histotype2=="high-grade serous" ~ histotype2,
                                 grade_cat == "Low grade"   ~ "other, low-grade",
                                 TRUE  ~ "other, high-grade"))
```

```{r Labeling data}
var_label(spatial_data) <- list(refage = "Patient age",
                                diagyear = "Diagnosis Year", 
                                histotype = "Histotype",
                                histotype2 = "Histotype, collapsed", 
                                histo_grade = "Histotype/Grade",
                                # sex = "Sex",
                                # race_cr_src_desc_1	 = "Race (raw)",
                                # race = "Race", ethnicity_src_desc = "Ethnicity",
                                # histology_desc = "Histology",
                                # bmi = "BMI (continuous)",
                                # bmi_cat = "BMI (categorical)",
                                grade_cat = "Grade",
                                new_dblkstat_treat_ca125 = "debulking status",
                                
                                int_to_event_mos = "Follow-up time from interview (months)",
                                vital_status_fin = "Vital Status"#,
                                # os_event = "Deaths"
)
```

```{r}
# interface_cd3plus_cells
# interface_percent_cd3plus_cells
# stroma_cd3_cells
# stroma_percent_cd3_cells
# tumor_cd3_cells
# tumor_percent_cd3_cells
# 
# "interface_tumor_cells"                                    
#  [21] "interface_stroma_cells"                                    "interface_cells"
# 
# "stroma_tumor_cells"                                        "stroma_stroma_cells"                                      
#  [57] "stroma_total_cells"
# 
# "tumor_tumor_cells"                                        
#  [91] "tumor_stroma_cells"                                        "tumor_total_cells"
# 
# [109] "boundary_length"
# 
# [111] "interface_cd3plus_do_ce_k_border"                          
# "tumor_cd3plus_do_ce_k_border"                             
# [129] "stroma_cd3plus_do_ce_k_border"                                   
# [137] "cd3 high/low"                                              "cd3+ cd8+ high/low"                                       
# [139] "Stroma cd3 high/low"                                       "Stroma cd3+ cd8+ high/low"                                
# [141] "Interface cd3 high/low"                                    "Interface cd3+ cd8+ high/low"
```

```{r}
spatial_data <- spatial_data %>% 
  select(suid, 
         contains("percent_"),
         # interface_cd3plus_cells,
         # interface_percent_cd3plus_cells,
         # stroma_cd3_cells, stroma_percent_cd3_cells,
         tumor_cd3_cells, tumor_cd3plus_cd8plus_cells,
         # tumor_percent_cd3_cells,
         ends_with("do_ce_k_border"),
         
         interfacekernel_tumorhalo_cells = interface_tumor_cells,
         interfacekernel_stromahalo_cells = interface_stroma_cells, 
         interface_total_cells = interface_cells,
         
         stromakernel_tumorhalo_cells = stroma_tumor_cells, 
         stromakernel_stromahalo_cells = stroma_stroma_cells,
         stroma_total_cells,
         
         tumorkernel_tumorhalo_cells = tumor_tumor_cells,
         tumorkernel_stromahalo_cells = tumor_stroma_cells, 
         tumor_total_cells,
         
         boundary_length,
         
         tumor_cd3_high_low = cd3_high_low, tumor_cd3_cd8_high_low = cd3_cd8_high_low,
         stroma_cd3_high_low, stroma_cd3_cd8_high_low,
         interface_cd3_high_low, interface_cd3_cd8_high_low,
         
         tumor_cd3_cells, tumor_cd3plus_cd8plus_cells, tumor_cd3plus_cd8minus_cells, 
         stroma_cd3_cells, stroma_cd3plus_cd8plus_cells, stroma_cd3plus_cd8minus_cells, 
         set : ncol(.)
  ) %>% 
  mutate(across(c(ends_with("_high_low"), -contains("cd8")), 
                ~ factor(., levels = 
                           c("Low (<=2%)", "High (>2%)")))) %>% 
  mutate(across(c(ends_with("_high_low") & contains("cd8")), 
                ~ factor(., levels = 
                           c("Low (<=1%)", "High (>1%)")))) %>% 
  mutate(across(c(contains("tumor") & contains("percent"), 
                  contains("stroma") & contains("percent"),
                  contains("tumor") & contains("do_ce_k_border"), 
                  contains("stroma") & contains("do_ce_k_border"),
                  "interface_cd3plus_do_ce_k_border", "interface_cd3plus_cd8plus_do_ce_k_border"
  ), ~ . / sd(., na.rm = TRUE) , .names = "{.col}_sdhr"))
```


# Inclusion
```{r}
spatial_data <- spatial_data %>% 
  filter(phase == 1 & #sample_type=="Compartment" & 
           (neoadj_treat != "yes" | is.na(neoadj_treat)))
```






# Data available
```{r}
colnames(spatial_data)[2:92]
```



# I. Patient characteristics
```{r Patient characteristics}
spatial_data %>% 
  select(refage, diagyear, stage_cat, grade, grade_cat, behavior, 
         histotype, histotype2, histo_grade,
         subtype, new_dblkstat_treat_ca125, vital_status_fin, int_to_event_mos) %>%
  tbl_summary() %>%
  bold_labels() %>%
  modify_header(label ~ "**Patient characteristics**")
```

```{r reset stage_cat}
spatial_data <- spatial_data %>% 
  mutate(stage_cat = case_when(
    is.na(stage_cat)                  ~ "Late", 
    TRUE                              ~ stage_cat)) #%>% 
  # set_variable_labels(stage_cat="Stage")
```

```{r}
spatial_data %>% 
  select(stage_cat) %>%
  tbl_summary() %>%
  bold_labels() 
```




# II. Immune abundance
## A. Distributions
```{r Quick look at immune abundance}
tbl1 <- spatial_data %>% 
  select(boundary_length) %>% 
  tbl_summary() %>%
  bold_labels() %>%
  modify_header(label ~ "**Immune abundance**")
tbl2 <- spatial_data %>% 
  select(starts_with("tumor_percent_"), 
         starts_with("tumorkernel_"),
         tumor_total_cells, 
         -ends_with("_sdhr")) %>% 
  tbl_summary() %>%
  bold_labels()
tbl3 <- spatial_data %>% 
  select(starts_with("stroma_percent_"),
         starts_with("stromakernel_"),
         stroma_total_cells, 
         -ends_with("_sdhr")) %>% 
  tbl_summary() %>%
  bold_labels()
tbl4 <- spatial_data %>% 
  select(starts_with("interface_percent_"),
         starts_with("interfacekernel_"),
         interface_total_cells) %>% 
  tbl_summary() %>%
  bold_labels()

tbl_stack(list(tbl1, tbl2, tbl3, tbl4), group_header = c("Boundary", "Tumor",
                                                         "Stroma", "Interface")) %>% 
  modify_footnote_header(
    footnote = "**Zero tumor cells from HALO were categorized as stroma using Kernel <br>
    Zero stroma cells from HALO were categorized as tumor using Kernel**",
    columns = all_stat_cols(),
    replace = FALSE
  ) %>% 
  as_gt() %>%
  gt::tab_options(row_group.border.top.color = "black",
                  row_group.border.top.width = 3,
                  row_group.border.bottom.color = "darkgrey",
                  row_group.font.size = 15,
                  row_group.font.weight = "bold",
                  row_group.text_transform = "uppercase",
                  row_group.background.color = "grey")
```

```{r}
spatial_data %>% 
  select(suid, contains("percent_"),
         -c(contains("tim3"), contains("pd1"), contains("minus"), 
            ends_with("_sdhr"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("tumor", "stroma", "interface"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|percent_|_cells|plus")) %>% 
  mutate(unit = "percent") %>% 
  ggplot(aes(y = value, x = unit, fill = compartment)) + 
  geom_violin(alpha = 0.5) +
  ggtitle("Compartment differences in immune cell abundance")+
  labs(x = "", y = "Immune cells (%)", fill = "Compartment")+
  facet_wrap(. ~ cell, scales = "free")


library(ggridges)
spatial_data %>% 
  select(suid, contains("percent_"),
         -c(contains("tim3"), contains("pd1"), contains("minus"), 
            ends_with("_sdhr"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("interface", "stroma", "tumor"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|percent_|_cells|plus")) %>% 
  

  ggplot(aes(x = value, y = compartment, fill = cell)) +
  geom_density_ridges(alpha=0.2, panel_scaling = TRUE) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "bottom"
  )+
  scale_y_discrete(expand = c(0, 0.1))+
  facet_wrap(~ cell, nrow = 1, scales = "free_x")
  

spatial_data %>% 
  select(suid, contains("percent_"),
         -c(contains("tim3"), contains("pd1"), contains("minus"), 
         ends_with("_sdhr"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("interface", "stroma", "tumor"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|percent_|_cells|plus")) %>% 
  ggplot(aes(x = value, y = compartment, fill = cell)) +
  geom_density_ridges(alpha=0.2, panel_scaling = TRUE) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "bottom"
  )+
  xlim(0,40)+
  scale_y_discrete(expand = c(0, 0.1))+
  facet_wrap(~ cell, nrow = 1)
```

## B. HR
Increased immune cell abundance is associated with poorer survival when using 1-SD normalization or categorical variables. 
```{r}
tbl_uvregression(
  spatial_data %>% select(int_to_event_mos, vital_status_fin,
                          contains("percent") & contains("_sdhr"),
                          contains("high_low")),
  method = coxph,
  y = Surv(time = int_to_event_mos, event = vital_status_fin),
  exponentiate = TRUE#,
  # include = c("tumor_percent_cd3_cells_sdhr", "tumor_percent_cd3plus_cd8plus_cells_sdhr", 
  #             "stroma_percent_cd3_cells_sdhr", "stroma_percent_cd3plus_cd8plus_cells_sdhr", 
  #             "interface_percent_cd3plus_cells_sdhr", "interface_percent_cd3plus_cd8plus_cells_sdhr",
  #             "tumor_cd3_high_low", "tumor_cd3_cd8_high_low", 
  #             "stroma_cd3_high_low", "stroma_cd3_cd8_high_low", 
  #             "interface_cd3_high_low", "interface_cd3_cd8_high_low"  )
) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**UVA**")
```

## C. HGSOC
Just checking if i get the same Ns as Brett
```{r}
hgsoc_data <- spatial_data %>% 
  filter(histotype2=="high-grade serous")

hgsoc_data %>% 
  select(tumor_percent_cd3_cells) %>% 
  tbl_summary() %>%
  bold_labels()
```

# III. Spatial clustering
## A. Distributions
```{r Spatial clustering}
tbl1 <- spatial_data %>% 
  select(starts_with("tumor_") & ends_with("_do_ce_k_border"), vital_status_fin) %>% 
  tbl_summary(by = vital_status_fin) %>%
  bold_labels() %>% 
  add_p() %>% bold_p() %>% add_overall() %>% 
  modify_spanning_header(c(stat_1, stat_2) ~ "**Death**")
tbl2 <- spatial_data %>% 
  select(starts_with("stroma_") & ends_with("_do_ce_k_border"), vital_status_fin) %>% 
  tbl_summary(by = vital_status_fin) %>%
  bold_labels() %>% 
  add_p() %>% bold_p() %>% add_overall()
tbl3 <- spatial_data %>% 
  select(starts_with("interface_") & ends_with("_do_ce_k_border"), vital_status_fin) %>% 
  tbl_summary(by = vital_status_fin) %>%
  bold_labels() %>% 
  add_p() %>% bold_p() %>% add_overall()

tbl_stack(list(tbl1, tbl2, tbl3), group_header = c("Tumor",
                                                   "Stroma", "Interface")) %>% 
  as_gt() %>%
  gt::tab_options(row_group.border.top.color = "black",
                  row_group.border.top.width = 3,
                  row_group.border.bottom.color = "darkgrey",
                  row_group.font.size = 15,
                  row_group.font.weight = "bold",
                  row_group.text_transform = "uppercase",
                  row_group.background.color = "grey")

```

```{r}
spatial_data %>% 
  select(suid, ends_with("_do_ce_k_border"),
         -c(contains("tim3"), contains("pd1"), contains("minus"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("tumor", "stroma", "interface"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|_do_ce_k_border|_cells|plus")) %>% 
  mutate(unit = "ripley") %>% 
  ggplot(aes(y = value, x = unit, fill = compartment)) + 
  geom_violin(alpha = 0.5) +
  ggtitle("Compartment differences in immune cell clustering")+
  labs(x = "", y = "Ripley's K", fill = "Compartment")+
  facet_wrap(. ~ cell, scales = "free")


library(ggridges)
spatial_data %>% 
  select(suid, ends_with("_do_ce_k_border"),
         -c(contains("tim3"), contains("pd1"), contains("minus"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("interface", "stroma", "tumor"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|_do_ce_k_border|_cells|plus")) %>% 
  

  ggplot(aes(x = value, y = compartment, fill = cell)) +
  geom_density_ridges(alpha=0.2, panel_scaling = TRUE) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "bottom"
  )+
  scale_y_discrete(expand = c(0, 0.1))+
  facet_wrap(~ cell, nrow = 1, scales = "free_x")
  

spatial_data %>% 
  select(suid, ends_with("_do_ce_k_border"),
         -c(contains("tim3"), contains("pd1"), contains("minus"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("interface", "stroma", "tumor"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|_do_ce_k_border|_cells|plus")) %>% 
  ggplot(aes(x = value, y = compartment, fill = cell)) +
  geom_density_ridges(alpha=0.2, panel_scaling = TRUE) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "bottom"
  )+
  scale_x_continuous(limits = c(-0.00001,3e7), 
                     labels = function(x) format(x, scientific = TRUE))+
  scale_y_discrete(expand = c(0, 0.1))+
  facet_wrap(~ cell, nrow = 1)
```

## B. HR
Increased immune cell abundance is associated with poorer survival when using 1-SD normalization or categorical variables. 
```{r}
tbl_uvregression(
  spatial_data %>% select(int_to_event_mos, vital_status_fin,
                          contains("percent") & contains("_sdhr")),
  method = coxph,
  y = Surv(time = int_to_event_mos, event = vital_status_fin),
  exponentiate = TRUE#,
  # include = c("tumor_percent_cd3_cells_sdhr", "tumor_percent_cd3plus_cd8plus_cells_sdhr", 
  #             "stroma_percent_cd3_cells_sdhr", "stroma_percent_cd3plus_cd8plus_cells_sdhr", 
  #             "interface_percent_cd3plus_cells_sdhr", "interface_percent_cd3plus_cd8plus_cells_sdhr",
  #             "tumor_cd3_high_low", "tumor_cd3_cd8_high_low", 
  #             "stroma_cd3_high_low", "stroma_cd3_cd8_high_low", 
  #             "interface_cd3_high_low", "interface_cd3_cd8_high_low"  )
) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**UVA**")

spat_vars <- spatial_data %>% select(
                          contains("percent") & contains("_sdhr")) %>%
  names()

# Create a list to store the results
cox_mva_fct <- lapply(spat_vars, function(var) {
  # Build the formula dynamically
  formula <- as.formula(paste("Surv(int_to_event_mos, vital_status_fin) ~", 
                              var, 
                              # "+ logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio",
                              "+ set + refage + stage_cat + histo_grade"))
  
  # Fit the Cox model
  cox_model <- coxph(formula, data = spatial_data)
  
  # Create a regression table for each model
  tbl_regression(cox_model, exponentiate = TRUE, include = var) %>%
    bold_p(t = 0.05) %>%
    add_nevent(location = "level") %>%
    add_n(location = "level") %>%
    bold_labels() %>%
    italicize_levels()
})
tbl_stack(cox_mva_fct) %>%
  modify_header(label ~ "**MVA <br>
                set + refage + stage_cat + histo_grade**")
  
```

# IV. Interaction between Immune abundance and clustering
## A. Distributions - create Dispersed and Clustered categories

Use the abundance categories definitions:

`cd3 high/low` = cut-off for tumor_percent_cd3_cells > 2   
`cd3+ cd8+ high/low` = cut-off for tumor_percent_cd3plus_cd8plus_cells > 1   
`Stroma cd3 high/low` = cut-off for stroma_percent_cd3_cells > 2   
`Stroma cd3+ cd8+ high/low` = cut-off for stroma_percent_cd3plus_cd8plus_cells > 1   
`Interface cd3 high/low` = cut-off for interface_percent_cd3plus_cells > 2   
`Interface cd3+ cd8+ high/low` = cut-off for interface_percent_cd3plus_cd8plus_cells > 1   

```{r}
spatial_data1 <- spatial_data %>% 
  # Tumor
  mutate(tumor_cd3_cluster_cat = case_when(
    tumor_percent_cd3_cells == 0                               ~ "No cell",
    tumor_cd3plus_do_ce_k_border < 
      median(tumor_cd3plus_do_ce_k_border, na.rm = TRUE)       ~ "Dispersed",
    tumor_cd3plus_do_ce_k_border >= 
      median(tumor_cd3plus_do_ce_k_border, na.rm = TRUE)       ~ "Clustered",
  )) %>% 
  mutate(tumor_cd3_abundance_cluster_cat = case_when(
    tumor_cd3_cluster_cat == "No cell"                         ~ "No cell",
    tumor_cd3_high_low == "Low (<=2%)" & 
      tumor_cd3_cluster_cat == "Dispersed"                     ~ "Low Dispersed cells",
    tumor_cd3_high_low == "Low (<=2%)" & 
      tumor_cd3_cluster_cat == "Clustered"                     ~ "Low Clustered cells",
    tumor_cd3_high_low == "High (>2%)" & 
      tumor_cd3_cluster_cat == "Dispersed"                     ~ "High Dispersed cells",
    tumor_cd3_high_low == "High (>2%)" & 
      tumor_cd3_cluster_cat == "Clustered"                     ~ "High Clustered cells"
  )) %>% 
  mutate(tumor_cd3cd8_cluster_cat = case_when(
    tumor_percent_cd3plus_cd8plus_cells == 0                           ~ "No cell",
    tumor_cd3plus_cd8plus_do_ce_k_border < 
      median(tumor_cd3plus_cd8plus_do_ce_k_border, na.rm = TRUE)       ~ "Dispersed",
    tumor_cd3plus_cd8plus_do_ce_k_border >= 
      median(tumor_cd3plus_cd8plus_do_ce_k_border, na.rm = TRUE)       ~ "Clustered",
  )) %>% 
  mutate(tumor_cd3cd8_abundance_cluster_cat = case_when(
    tumor_cd3cd8_cluster_cat == "No cell"                         ~ "No cell",
    tumor_cd3_cd8_high_low == "Low (<=1%)" & 
      tumor_cd3cd8_cluster_cat == "Dispersed"                     ~ "Low Dispersed cells",
    tumor_cd3_cd8_high_low == "Low (<=1%)" & 
      tumor_cd3cd8_cluster_cat == "Clustered"                     ~ "Low Clustered cells",
    tumor_cd3_cd8_high_low == "High (>1%)" & 
      tumor_cd3cd8_cluster_cat == "Dispersed"                     ~ "High Dispersed cells",
    tumor_cd3_cd8_high_low == "High (>1%)" & 
      tumor_cd3cd8_cluster_cat == "Clustered"                     ~ "High Clustered cells"
  )) %>% 
  # Stroma
  mutate(stroma_cd3_cluster_cat = case_when(
    stroma_percent_cd3_cells == 0                               ~ "No cell",
    stroma_cd3plus_do_ce_k_border < 
      median(stroma_cd3plus_do_ce_k_border, na.rm = TRUE)       ~ "Dispersed",
    stroma_cd3plus_do_ce_k_border >= 
      median(stroma_cd3plus_do_ce_k_border, na.rm = TRUE)       ~ "Clustered",
  )) %>% 
  mutate(stroma_cd3_abundance_cluster_cat = case_when(
    stroma_cd3_cluster_cat == "No cell"                         ~ "No cell",
    stroma_cd3_high_low == "Low (<=2%)" & 
      stroma_cd3_cluster_cat == "Dispersed"                     ~ "Low Dispersed cells",
    stroma_cd3_high_low == "Low (<=2%)" & 
      stroma_cd3_cluster_cat == "Clustered"                     ~ "Low Clustered cells",
    stroma_cd3_high_low == "High (>2%)" & 
      stroma_cd3_cluster_cat == "Dispersed"                     ~ "High Dispersed cells",
    stroma_cd3_high_low == "High (>2%)" & 
      stroma_cd3_cluster_cat == "Clustered"                     ~ "High Clustered cells"
  )) %>% 
  mutate(stroma_cd3cd8_cluster_cat = case_when(
    stroma_percent_cd3plus_cd8plus_cells == 0                           ~ "No cell",
    stroma_cd3plus_cd8plus_do_ce_k_border < 
      median(stroma_cd3plus_cd8plus_do_ce_k_border, na.rm = TRUE)       ~ "Dispersed",
    stroma_cd3plus_cd8plus_do_ce_k_border >= 
      median(stroma_cd3plus_cd8plus_do_ce_k_border, na.rm = TRUE)       ~ "Clustered",
  )) %>% 
  mutate(stroma_cd3cd8_abundance_cluster_cat = case_when(
    stroma_cd3cd8_cluster_cat == "No cell"                         ~ "No cell",
    stroma_cd3_cd8_high_low == "Low (<=1%)" & 
      stroma_cd3cd8_cluster_cat == "Dispersed"                     ~ "Low Dispersed cells",
    stroma_cd3_cd8_high_low == "Low (<=1%)" & 
      stroma_cd3cd8_cluster_cat == "Clustered"                     ~ "Low Clustered cells",
    stroma_cd3_cd8_high_low == "High (>1%)" & 
      stroma_cd3cd8_cluster_cat == "Dispersed"                     ~ "High Dispersed cells",
    stroma_cd3_cd8_high_low == "High (>1%)" & 
      stroma_cd3cd8_cluster_cat == "Clustered"                     ~ "High Clustered cells"
  )) %>% 
  # Interface
  mutate(interface_cd3_cluster_cat = case_when(
    interface_percent_cd3plus_cells == 0                               ~ "No cell",
    interface_cd3plus_do_ce_k_border < 
      median(interface_cd3plus_do_ce_k_border, na.rm = TRUE)       ~ "Dispersed",
    interface_cd3plus_do_ce_k_border >= 
      median(interface_cd3plus_do_ce_k_border, na.rm = TRUE)       ~ "Clustered",
  )) %>% 
  mutate(interface_cd3_abundance_cluster_cat = case_when(
    interface_cd3_cluster_cat == "No cell"                         ~ "No cell",
    interface_cd3_high_low == "Low (<=2%)" & 
      interface_cd3_cluster_cat == "Dispersed"                     ~ "Low Dispersed cells",
    interface_cd3_high_low == "Low (<=2%)" & 
      interface_cd3_cluster_cat == "Clustered"                     ~ "Low Clustered cells",
    interface_cd3_high_low == "High (>2%)" & 
      interface_cd3_cluster_cat == "Dispersed"                     ~ "High Dispersed cells",
    interface_cd3_high_low == "High (>2%)" & 
      interface_cd3_cluster_cat == "Clustered"                     ~ "High Clustered cells"
  )) %>% 
  mutate(interface_cd3cd8_cluster_cat = case_when(
    interface_percent_cd3plus_cd8plus_cells == 0                           ~ "No cell",
    interface_cd3plus_cd8plus_do_ce_k_border < 
      median(interface_cd3plus_cd8plus_do_ce_k_border, na.rm = TRUE)       ~ "Dispersed",
    interface_cd3plus_cd8plus_do_ce_k_border >= 
      median(interface_cd3plus_cd8plus_do_ce_k_border, na.rm = TRUE)       ~ "Clustered",
  )) %>% 
  mutate(interface_cd3cd8_abundance_cluster_cat = case_when(
    interface_cd3cd8_cluster_cat == "No cell"                         ~ "No cell",
    interface_cd3_cd8_high_low == "Low (<=1%)" & 
      interface_cd3cd8_cluster_cat == "Dispersed"                     ~ "Low Dispersed cells",
    interface_cd3_cd8_high_low == "Low (<=1%)" & 
      interface_cd3cd8_cluster_cat == "Clustered"                     ~ "Low Clustered cells",
    interface_cd3_cd8_high_low == "High (>1%)" & 
      interface_cd3cd8_cluster_cat == "Dispersed"                     ~ "High Dispersed cells",
    interface_cd3_cd8_high_low == "High (>1%)" & 
      interface_cd3cd8_cluster_cat == "Clustered"                     ~ "High Clustered cells"
  )) %>% 
  mutate(across(c(ends_with("_cluster_cat"), -ends_with("_abundance_cluster_cat")), 
                ~ factor(., levels = 
                           c("Clustered", "Dispersed", "No cell")))) %>% 
  mutate(across(c(ends_with("_abundance_cluster_cat")), 
                ~ factor(., levels = 
                           c("Low Clustered cells", "Low Dispersed cells", 
                             "High Dispersed cells", "High Clustered cells", "No cell")))) %>% 
  mutate(clustering_cd3_tumor_stroma = case_when(
    tumor_cd3_cluster_cat == "Dispersed" |
      stroma_cd3_cluster_cat == "Dispersed"        ~ "Both dispersed",
    tumor_cd3_cluster_cat == "Clustered" |
      stroma_cd3_cluster_cat == "Clustered"        ~ "Both clustered",
    
    tumor_cd3_cluster_cat == "Dispersed" |
      stroma_cd3_cluster_cat == "Clustered"        ~ "Cluster stroma/dispersed tumore",
    tumor_cd3_cluster_cat == "Clustered" |
      stroma_cd3_cluster_cat == "Dispersed"        ~ "Cluster tumor/dispersed stroma"
  )) %>% 
  mutate(across(c(ends_with("do_ce_k_border_sdhr")), ~ case_when(
    ntile(., 5) == 1 ~ "1",
    ntile(., 5) == 2 ~ "2",
    ntile(., 5) == 3 ~ "3",
    ntile(., 5) == 4 ~ "4",
    ntile(., 5) == 5 ~ "High clustered"),
    .names = "{.col}_5cat"))



spatial_data1 %>% 
  select(ends_with("cluster_cat"), #ends_with("_5cat"), 
         clustering_cd3_tumor_stroma
         ) %>% 
  tbl_summary() %>%
  bold_labels() 
```

## B. HR
When using categorical variables :

- Increased immune cell clustering is associated with better survival or Dispersed cells is associated with worst survival. 
- High abundance / high cluster is associated with better survival
```{r}
tbl_uvregression(
  spatial_data1 %>% select(int_to_event_mos, vital_status_fin,
                           ends_with("cluster_cat")),
  method = coxph,
  y = Surv(time = int_to_event_mos, event = vital_status_fin),
  exponentiate = TRUE#,
  # include = c("tumor_cd3_cluster_cat",
  #             "tumor_cd3_abundance_cluster_cat",
  #             "stroma_cd3_cluster_cat",
  #             "stroma_cd3_abundance_cluster_cat",
  #             # "interface_cd3_cluster_cat",
  #             # "interface_cd3_abundance_cluster_cat",
  #             "clustering_cd3_tumor_stroma",
  #             # "tumor_cd3plus_do_ce_k_border_sdhr_5cat",
  #             # "tumor_cd3plus_cd8plus_do_ce_k_border_sdhr_5cat",
  #             # "tumor_cd3plus_cd8minus_do_ce_k_border_sdhr_5cat",
  #             # "stroma_cd3plus_do_ce_k_border_sdhr_5cat",
  #             # "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr_5cat",
  #             # "stroma_cd3plus_cd8minus_do_ce_k_border_sdhr_5cat"
  # )
  ) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**UVA**")
```





# WTS eligible???
```{r}
wts_eligible <- spatial_data %>% 
  filter(tumorkernel_tumorhalo_cells >= 500 & stromakernel_stromahalo_cells >= 500)
```


# V. Logit approach

logit_prop (T cell % in Tumor – logit transformed)
logit_stroma_prop (T cell % in stroma – logit transformed)
cd8p_cd8n_logratio (Tumor CD8+/CD8- ratio – log transformed) 
cd8p_cd8n_stroma_logratio (Stroma CD8+/CD8- ratio – log transformed)

```{r Brett logit transf}
eps <- 1e-7
spatial_log_data <- spatial_data1 %>% 
  mutate(
  #        cyto_stroma_z = scale(stroma_cyto_tcell, center = TRUE, scale = TRUE),
  #        cyto_tumor_z =  scale(tumor_cyto_tcell, center = TRUE, scale = TRUE),
  #        cyto_interface_z =  scale(interface_cyto_tcell, center = TRUE, scale = TRUE),
  #        stroma_z = scale(stroma_tcell, center = TRUE, scale = TRUE),
  #        tumor_z =  scale(tumor_tcell, center = TRUE, scale = TRUE),
  #        interface_z =  scale(interface_tcell, center = TRUE, scale = TRUE),
  #        stroma_ex_z = scale(stroma_percent_cd3plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        tumor_ex_z =  scale(tumor_percent_cd3plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        interface_ex_z =  scale(interface_percent_cd3plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        stroma_cytoex_z = scale(stroma_percent_cd3plus_cd8plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        tumor_cytoex_z =  scale(tumor_percent_cd3plus_cd8plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        interface_cytoex_z =  scale(interface_percent_cd3plus_cd8plus_tim3plus_cells, center = TRUE, scale = TRUE)) %>%
  # #same for CD8- subtype for models with CD8- and CD8+ both included - sensitivity analyses
  # mutate(nocyto_tumor_z = scale(tumor_percent_cd3plus_cd8minus_cells, center=TRUE, scale=TRUE ),
  #        tumor_nocytoex_z = scale(tumor_percent_cd3plus_cd8minus_tim3plus_cells, center=TRUE, scale=TRUE)) %>%
  # mutate(nocyto_stroma_z = scale(stroma_percent_cd3plus_cd8minus_cells, center=TRUE, scale=TRUE ),
  #        stroma_nocytoex_z = scale(stroma_percent_cd3plus_cd8minus_tim3plus_cells, center=TRUE, scale=TRUE)) %>%
  # mutate(tumor_noncyto_tcell_timexhausted = (  ifelse(tumor_cd3plus_cd8minus_cells != 0,
  #                                                     (tumor_cd3plus_cd8minus_tim3plus_cells  / 
  #                                                        tumor_cd3plus_cd8minus_cells)*100, 0) )) %>%
  # mutate(stroma_noncyto_tcell_timexhausted = (  ifelse(stroma_cd3plus_cd8minus_cells != 0,
  #                                                      (stroma_cd3plus_cd8minus_tim3plus_cells  / 
  #                                                         stroma_cd3plus_cd8minus_cells)*100, 0) )) %>%
  # mutate(interface_noncyto_tcell_timexhausted = (  ifelse(interface_cd3plus_cd8minus_cells != 0,
  #                                                         (interface_cd3plus_cd8minus_tim3plus_cells  / 
  #                                                            interface_cd3plus_cd8minus_cells)*100, 0) )
  ) %>%
  # scale
  mutate(tumor_cd3plus_do_ce_k_border_scale = scales::rescale(tumor_cd3plus_do_ce_k_border, to=c(0,1)),
         stroma_cd3plus_do_ce_k_border_scale = scales::rescale(stroma_cd3plus_do_ce_k_border, to=c(0,1)),
         tumor_cd3plus_cd8plus_do_ce_k_border_scale = scales::rescale(tumor_cd3plus_cd8plus_do_ce_k_border, to=c(0,1)),
         stroma_cd3plus_cd8plus_do_ce_k_border_scale = scales::rescale(stroma_cd3plus_cd8plus_do_ce_k_border, to=c(0,1))) %>% 
  ## RATIOS
  mutate(
         # cd3p_cd3n_ratio = (tumor_cd3_cells + eps) / ((tumor_total_cells - tumor_cd3_cells) + eps),
         # cd3p_cd3n_logratio = log(cd3p_cd3n_ratio),
         # cd3p_cd3n_stroma_ratio = (stroma_cd3_cells + eps) / ((stroma_total_cells - stroma_cd3_cells) + eps),
         # cd3p_cd3n_stroma_logratio = log(cd3p_cd3n_stroma_ratio),
         # cd3p_cd3n_interface_ratio = (interface_cd3plus_cells + eps) / ((interface_total_cells - interface_cd3plus_cells) + eps),
         # cd3p_cd3n_interface_logratio = log(cd3p_cd3n_interface_ratio),
         cd8p_cd8n_ratio = (tumor_cd3plus_cd8plus_cells + eps) / (tumor_cd3plus_cd8minus_cells + eps),
         cd8p_cd8n_logratio = log(cd8p_cd8n_ratio),
         cd8p_cd8n_stroma_ratio = (stroma_cd3plus_cd8plus_cells + eps) / (stroma_cd3plus_cd8minus_cells + eps),
         cd8p_cd8n_stroma_logratio = log(cd8p_cd8n_stroma_ratio),
         #    cd8p_cd8n_interface_ratio = (interface_cd3plus_cd8plus_cells + eps) / (interface_cd3plus_cd8minus_cells + eps),
         #    cd8p_cd8n_interface_logratio = log(cd8p_cd8n_interface_ratio),
         logit_prop = qlogis(pmin(pmax(tumor_percent_cd3_cells/100, eps), 1-eps)),
         logit_stroma_prop = qlogis(pmin(pmax(stroma_percent_cd3_cells/100, eps), 1-eps)),
         #    logit_interface_prop = qlogis(pmin(pmax(interface_percent_cd3plus_cells/100, eps), 1-eps)),
         logit_cytoprop = qlogis(pmin(pmax(tumor_percent_cd3plus_cd8plus_cells/100, eps), 1-eps)),
         logit_stroma_cytoprop = qlogis(pmin(pmax(stroma_percent_cd3plus_cd8plus_cells/100, eps), 1-eps)),
         #    logit_interface_cytoprop = qlogis(pmin(pmax(interface_percent_cd3plus_cd8plus_cells/100, eps), 1-eps)),
         #    logit_cytoprop_relative = qlogis(pmin(pmax(tumor_cyto_tcell_relative/100, eps), 1-eps)),
         #    logit_stroma_cytoprop_relative = qlogis(pmin(pmax(stroma_cyto_tcell_relative/100, eps), 1-eps)),
         #    logit_interface_cytoprop_relative = qlogis(pmin(pmax(interface_cyto_tcell_relative/100, eps), 1-eps)),
         #    tumor_cd3plus_tim3minus_cells = tumor_cd3_cells -  tumor_cd3plus_tim3plus_cells,
         #    stroma_cd3plus_tim3minus_cells = stroma_cd3_cells -  stroma_cd3plus_tim3plus_cells,
         #    interface_cd3plus_tim3minus_cells = interface_cd3plus_cells -  interface_cd3plus_tim3plus_cells,
         #    timp_timn_ratio =  (tumor_cd3plus_tim3plus_cells + eps) / (tumor_cd3plus_tim3minus_cells + eps),
         #    timp_timn_logratio = log(timp_timn_ratio),
         #    timp_timn_stroma_ratio =  (stroma_cd3plus_tim3plus_cells + eps) / (stroma_cd3plus_tim3minus_cells + eps),
         #    timp_timn_stroma_logratio = log(timp_timn_stroma_ratio),
         #    timp_timn_interface_ratio =  (interface_cd3plus_tim3plus_cells + eps) / (interface_cd3plus_tim3minus_cells + eps),
         #    timp_timn_interface_logratio = log(timp_timn_interface_ratio),
         #    logit_exhprop = qlogis(pmin(pmax(tumor_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_stroma_exhprop = qlogis(pmin(pmax(stroma_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_interface_exhprop = qlogis(pmin(pmax(interface_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_exh2prop = qlogis(pmin(pmax(tumor_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_stroma_exh2prop = qlogis(pmin(pmax(stroma_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_interface_exh2prop = qlogis(pmin(pmax(interface_tcell_bothexhausted/100, eps), 1-eps)),
         #   logit_exh2prop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_pd1plus_tim3plus_cells /100, eps), 1-eps)),
         #    logit_stroma_exh2prop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_pd1plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exh2prop_total = qlogis(pmin(pmax(interface_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         # logit_exh2cytoprop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_cd8plus_pd1plus_tim3plus_cells /100, eps), 1-eps)),
         #    logit_stroma_exh2cytoprop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_cd8plus_pd1plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exh2cytoprop_total = qlogis(pmin(pmax(interface_percent_cd3plus_cd8plus_pd1plus_tim3plus_cells/100, eps), 1-eps)),
         # logit_exh2prop_relative = qlogis(pmin(pmax(tumor_pd1_relative/100, eps), 1-eps)),
         #    logit_stroma_exh2prop_relative = qlogis(pmin(pmax(stroma_pd1_relative/100, eps), 1-eps)),
         #    logit_interface_exh2prop_relative = qlogis(pmin(pmax(interface_pd1_relative/100, eps), 1-eps)),
         #   logit_exh2cytoprop_relative = qlogis(pmin(pmax(tumor_pd1_cyto_relative/100, eps), 1-eps)),
         #    logit_stroma_exh2cytoprop_relative = qlogis(pmin(pmax(stroma_pd1_cyto_relative/100, eps), 1-eps)),
         #    logit_interface_exh2cytoprop_relative = qlogis(pmin(pmax(interface_pd1_cyto_relative/100, eps), 1-eps)),
         #    tumor_cd3plus_cd8plus_tim3minus_cells = tumor_cd3plus_cd8plus_cells - tumor_cd3plus_cd8plus_tim3plus_cells,
         #    stroma_cd3plus_cd8plus_tim3minus_cells = stroma_cd3plus_cd8plus_cells - stroma_cd3plus_cd8plus_tim3plus_cells,
         #    interface_cd3plus_cd8plus_tim3minus_cells = interface_cd3plus_cd8plus_cells - interface_cd3plus_cd8plus_tim3plus_cells,
         #    timp_timn_cd8_ratio = (tumor_cd3plus_cd8plus_tim3plus_cells + eps) / (tumor_cd3plus_cd8plus_tim3minus_cells + eps),
         #    timp_timn_cd8_logratio = log(timp_timn_cd8_ratio),
         #    timp_timn_cd8_stroma_ratio = (stroma_cd3plus_cd8plus_tim3plus_cells + eps) / (stroma_cd3plus_cd8plus_tim3minus_cells + eps),
         #    timp_timn_cd8_stroma_logratio = log(timp_timn_cd8_stroma_ratio),
         #    timp_timn_cd8_interface_ratio = (interface_cd3plus_cd8plus_tim3plus_cells + eps) / (interface_cd3plus_cd8plus_tim3minus_cells + eps),
         #    timp_timn_cd8_interface_logratio = log(timp_timn_cd8_interface_ratio),
         # logit_exhprop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_stroma_exhprop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exhprop_total = qlogis(pmin(pmax(interface_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         #   logit_exhcytoprop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_cd8plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_stroma_exhcytoprop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_cd8plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exhcytoprop_total = qlogis(pmin(pmax(interface_percent_cd3plus_cd8plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_exhcytoprop = qlogis(pmin(pmax(tumor_cyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_stroma_exhcytoprop = qlogis(pmin(pmax(stroma_cyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_interface_exhcytoprop = qlogis(pmin(pmax(interface_cyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_exh2cytoprop = qlogis(pmin(pmax(tumor_cyto_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_stroma_exh2cytoprop = qlogis(pmin(pmax(stroma_cyto_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_interface_exh2cytoprop = qlogis(pmin(pmax(interface_cyto_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_exhnocytoprop = qlogis(pmin(pmax(tumor_noncyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_stroma_exhnocytoprop = qlogis(pmin(pmax(stroma_noncyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_interface_exhnocytoprop = qlogis(pmin(pmax(interface_noncyto_tcell_timexhausted/100, eps), 1-eps)) 
         
         
         logit_cd3tumor_ripley = qlogis(pmin(pmax(tumor_cd3plus_do_ce_k_border_scale, eps), 1-eps)),
         logit_cd3stroma_ripley = qlogis(pmin(pmax(stroma_cd3plus_do_ce_k_border_scale, eps), 1-eps)),
         logit_cd8tumor_ripley = qlogis(pmin(pmax(tumor_cd3plus_cd8plus_do_ce_k_border_scale, eps), 1-eps)),
         logit_cd8stroma_ripley = qlogis(pmin(pmax(stroma_cd3plus_cd8plus_do_ce_k_border_scale, eps), 1-eps))
  ) %>%
  #Paired Differences
  # mutate(cd3_total_diff = logit_prop - logit_stroma_prop,
                     # cd8_total_diff = logit_cytoprop - logit_stroma_cytoprop,
                     # cd8_ratio_diff = cd8p_cd8n_logratio - cd8p_cd8n_stroma_logratio,
                     # exh_total_diff = logit_exhprop_total - logit_stroma_exhprop_total,
                     # exhcd8_total_diff = logit_exhcytoprop_total - logit_stroma_exhcytoprop_total,
                     # exh_ratio_diff = timp_timn_logratio - timp_timn_stroma_logratio, 
                     # exhcd8_ratio_diff = timp_timn_cd8_logratio - timp_timn_cd8_stroma_logratio,
                     # exh2_total_diff = logit_exh2prop_total - logit_stroma_exh2prop_total,
                     # exh2cd8_total_diff = logit_exh2cytoprop_total - logit_stroma_exh2cytoprop_total,
                     # exh2_ratio_diff = logit_exh2prop - logit_stroma_exh2prop,
                     # exh2cd8_ratio_diff = logit_exh2cytoprop - logit_stroma_exh2cytoprop) %>%
  # set_variable_labels(cd3_total_diff = "CD3+ Difference (T-S)",
  #                    cd8_total_diff = "CD8+ Difference (T-S)",
  #                    cd8_ratio_diff = "CD8 Ratio Difference (T-S)",
  #                    exh_total_diff = "TIM3+ Difference (T-S)",
  #                    exhcd8_total_diff = "TIM3+CD8+ Difference (T-S)",
  #                    exh_ratio_diff =  "TIM3 Ratio Difference (T-S)", 
  #                    exhcd8_ratio_diff = "TIM3 CD8 Ratio Difference (T-S)",
  #                    exh2_total_diff = "TIM3+PD1+ Difference (T-S)",
  #                    exh2cd8_total_diff = "TIM3+PD1+CD8+ Difference (T-S)",
  #                    exh2_ratio_diff = "TIM3+PD1 Ratio Difference (T-S)",
  #                    exh2cd8_ratio_diff = "TIM3+PD1 CD8 Ratio Difference (T-S)") %>%
  
  set_variable_labels(
                      # tumor_tcell="Tumor T cell",
                      # tumor_cyto_tcell="Tumor Cytotoxic T cell",
                      # tumor_cd8minus_tcell="Tumor CD8- T cell",
                      # tumor_bcell = "Tumor B cell",
                      # tumor_cyto_tcell_relative= "Tumor Cytotoxic T cell (% of T cells)",
                      # tumor_cd8minus_tcell_relative= "Tumor CD8- T cell (% of T cells)",
                      # tcell_exh2_grp = "Tumor TIM3+PD1+ Abundance",
                      # tcell_exh2_stroma_grp = "Stroma TIM3+PD1+ Abundance",
                      # tcell_exh2_interface_grp = "Interface TIM3+PD1+ Abundance",
                      # tcell_cyto_exh2_grp = "Tumor Cyto TIM3+PD1+ Abundance",
                      # tcell_cyto_exh2_stroma_grp = "Stroma Cyto TIM3+PD1+ Abundance",
                      # tcell_cyto_exh2_interface_grp = "Interface Cyto TIM3+PD1+ Abundance",
                      # 
                      # stroma_tcell="Stroma T cell",
                      # stroma_cyto_tcell="Stroma Cytotoxic T cell",
                      # stroma_cd8minus_tcell="Stroma CD8- T cell",
                      # stroma_bcell = "Stroma B cell",
                      # stroma_cyto_tcell_relative= "Stroma Cytotoxic T cell (% of T cells)",
                      # stroma_cd8minus_tcell_relative= "Stroma CD8- T cell (% of T cells)",
                      # cd3p_cd3n_logratio = "Tumor CD3+/CD3- Ratio (Log)",
                      # cd3p_cd3n_stroma_logratio = "Stroma CD3+/CD3- Ratio (Log)",
                      # cd3p_cd3n_interface_logratio = "Interface CD3+/CD3- Ratio (Log)",
                      cd8p_cd8n_logratio = "Tumor CD8+/CD8- Ratio (Log)",
                      cd8p_cd8n_stroma_logratio = "Stroma CD8+/CD8- Ratio (Log)",
                      # cd8p_cd8n_interface_logratio = "Interface CD8+/CD8- Ratio (Log)",
                      # timp_timn_logratio = "Tumor TIM+/TIM- Ratio (Log)",
                      # timp_timn_stroma_logratio = "Stroma TIM+/TIM- Ratio (Log)",
                      # timp_timn_interface_logratio = "Interface TIM+/TIM- Ratio (Log)",
                      # timp_timn_cd8_logratio = "Tumor TIM+/TIM- CD8+ Ratio (Log)",
                      # timp_timn_cd8_stroma_logratio = "Stroma TIM+/TIM- CD8+ Ratio (Log)",
                      # timp_timn_cd8_interface_logratio = "Interface TIM+/TIM- CD8+ Ratio (Log)",
                      
                      logit_prop = "Tumor T cell proportion (Logit)",
                      logit_stroma_prop = "Stroma T cell proportion (Logit)",
                      #  logit_interface_prop = "Interface T cell proportion (Logit)",
                      logit_cytoprop = "Tumor CD8+ T cell proportion (Logit)",
                      #  logit_exhprop = "Tumor Exhausted T cell proportion (Logit)",
                      #  logit_exhcytoprop = "Tumor Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_exhnocytoprop = "Tumor Exhausted CD8- T cell proportion (Logit)",
                      logit_stroma_cytoprop = "Stroma CD8+ T cell proportion (Logit)",
                      #  logit_stroma_exhprop ="Stroma Exhausted T cell proportion (Logit)",
                      #  logit_stroma_exhcytoprop = "Stroma Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_stroma_exhnocytoprop = "Stroma Exhausted CD8- T cell proportion (Logit)",
                      #  logit_interface_cytoprop = "Interface CD8+ T cell proportion (Logit)",
                      #  logit_interface_exhprop ="Interface Exhausted T cell proportion (Logit)",
                      #  logit_interface_exhcytoprop = "Interface Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_interface_exhnocytoprop = "Interface Exhausted CD8- T cell proportion (Logit)",
                      #  
                      #  logit_exh2prop = "Tumor PD1 Exhausted T cell proportion (Logit)",
                      #  logit_exh2cytoprop = "Tumor PD1 Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_stroma_exh2prop ="Stroma PD1 Exhausted T cell proportion (Logit)",
                      #  logit_stroma_exh2cytoprop = "Stroma PD1 Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_interface_exh2prop ="Interface PD1 Exhausted T cell proportion (Logit)",
                      #  logit_interface_exh2cytoprop = "Interface PD1 Exhausted CD8+ T cell proportion (Logit)",
                      #  
                      #  tumor_z="Tumor T cell abundance (1 SD)",
                      #  stroma_z ="Stroma T cell abundance (1 SD)",
                      #  interface_z ="Interface T cell abundance (1 SD)",
                      #  cyto_tumor_z="Tumor CD8+ abundance (1 SD)",
                      #  cyto_stroma_z ="Stroma CD8+ abundance (1 SD)",
                      #  cyto_interface_z ="Interface CD8+ abundance (1 SD)",
                      #  nocyto_tumor_z="Tumor CD8- abundance (1 SD)",
                      #  nocyto_stroma_z ="Stroma CD8- abundance (1 SD)",
                      #  tumor_noncyto_tcell_timexhausted= "Tumor CD8- Exhausted Proportion",
                      #  stroma_noncyto_tcell_timexhausted = "Stroma CD8- Exhausted Proportion",
                      #  interface_noncyto_tcell_timexhausted = "Interface CD8- Exhausted Proportion",
                      #  
                      #  interface_tcell="Interface T cell",
                      #  interface_cyto_tcell="Interface Cytotoxic T cell",
                      #  interface_bcell = "Interface B cell",
                      #  interface_cyto_tcell_relative= "Interface Cytotoxic T cell (% of T cells)",
                      # 
                      # tumor_tcell_timexhausted= "Tumor Exhausted T cell - TIM3+",
                      # stroma_tcell_timexhausted= "Stroma Exhausted T cell - TIM3+",
                      # interface_tcell_timexhausted= "Interface Exhausted T cell - TIM3+",
                      # tumor_cyto_tcell_timexhausted= "Tumor Exhausted Cyto T cell - TIM3+",
                      # stroma_cyto_tcell_timexhausted= "Stroma Exhausted Cyto T cell - TIM3+",
                      # interface_cyto_tcell_timexhausted= "Interface Exhausted Cyto T cell - TIM3+",
                      # 
                      # tumor_tcell_bothexhausted= "Tumor Terminal Exhausted T cell - PD1+TIM3+",
                      # stroma_tcell_bothexhausted= "Stroma Terminal Exhausted T cell - PD1+TIM3+",
                      # interface_tcell_bothexhausted= "Interface Terminal Exhausted T cell - PD1+TIM3+",
                      # tumor_cyto_tcell_bothexhausted= "Tumor Terminal Exhausted Cyto T cell - PD1+TIM3+",
                      # stroma_cyto_tcell_bothexhausted= "Stroma Terminal Exhausted Cyto T cell - PD1+TIM3+",
                      # interface_cyto_tcell_bothexhausted= "Interface Terminal Exhausted Cyto T cell - PD1+TIM3+"
  )
```

```{r Brett create exhausted}
## Create  datasets - all histotypes and limited to HGSOC datasets 
## Rerunning median groups so reset thresholds to the limited subjects distributions

# data <- compartment %>%
# #reset median cutoff for this dataset distribution
#   mutate(tcell_exhausted =
#           ifelse(tumor_tcell_timexhausted > median(tumor_tcell_timexhausted, na.rm = TRUE), "Above Median", "Below Median")) %>%
#   mutate(tcell_exhausted_stroma = 
#         ifelse(stroma_tcell_timexhausted > median(stroma_tcell_timexhausted, na.rm = TRUE),  "Above Median", "Below Median")) %>%
#   mutate(tcell_exhausted_interface = 
#         ifelse(interface_tcell_timexhausted > median(interface_tcell_timexhausted, na.rm = TRUE),  "Above Median", "Below Median")) %>%
#   mutate(tcell_cyto_exhausted =
#           ifelse(tumor_cyto_tcell_timexhausted > median(tumor_cyto_tcell_timexhausted, na.rm = TRUE), "Above Median", "Below Median")) %>%
#   mutate(tcell_cyto_exhausted_stroma = 
#         ifelse(stroma_cyto_tcell_timexhausted > median(stroma_cyto_tcell_timexhausted, na.rm = TRUE),  "Above Median", "Below Median")) %>%
#   mutate(tcell_cyto_exhausted_interface = 
#         ifelse(interface_cyto_tcell_timexhausted > median(interface_cyto_tcell_timexhausted, na.rm = TRUE),  "Above Median", "Below Median")) %>%
#   
# #exhaustion variable with low t cells by exhaustion  
#   mutate(tcell_cat = case_when (
#             `cd3 high/low` == "Low (<=2%)" & tcell_exhausted == "Above Median"     ~ "Low T cells-High Exhaustion",
#              `cd3 high/low` == "Low (<=2%)" & tcell_exhausted == "Below Median"   ~ "Low T cells-Low Exhaustion",
#             `cd3 high/low` != "Low (<=2%)" & tcell_exhausted == "Above Median"     ~ "High T cells-High Exhaustion",
#             `cd3 high/low` != "Low (<=2%)" & tcell_exhausted == "Below Median"     ~ "High T cells-Low Exhaustion"
#   )) %>%
#   mutate(tcell_cat = factor(tcell_cat, levels=c("Low T cells-Low Exhaustion", "Low T cells-High Exhaustion", 
#                                                 "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#   mutate(tcell_cat_stroma = case_when (
#         `Stroma cd3 high/low` == "Low (<=2%)" & tcell_exhausted_stroma == "Above Median" ~ "Low T cells-High Exhaustion",
#         `Stroma cd3 high/low` == "Low (<=2%)" & tcell_exhausted_stroma == "Below Median" ~ "Low T cells-Low Exhaustion",
#       `Stroma cd3 high/low` != "Low (<=2%)" & tcell_exhausted_stroma == "Above Median" ~ "High T cells-High Exhaustion",
#       `Stroma cd3 high/low` != "Low (<=2%)" & tcell_exhausted_stroma == "Below Median" ~ "High T cells-Low Exhaustion"
#   )) %>%
#   mutate(tcell_cat_stroma = factor(tcell_cat_stroma, levels=c("Low T cells-Low Exhaustion", 
#               "Low T cells-High Exhaustion","High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#   mutate(tcell_cat_interface = case_when (
#       interface_percent <1 ~ "No interface",
#       `Interface cd3 high/low` == "Low (<=2%)" & tcell_exhausted_interface == "Above Median" ~ "Low T cells-High Exhaustion",
#       `Interface cd3 high/low` == "Low (<=2%)" & tcell_exhausted_interface == "Below Median" ~ "Low T cells-Low Exhaustion",
#       `Interface cd3 high/low` != "Low (<=2%)" & tcell_exhausted_interface == "Above Median" ~ "High T cells-High Exhaustion",
#       `Interface cd3 high/low` != "Low (<=2%)" & tcell_exhausted_interface == "Below Median" ~ "High T cells-Low Exhaustion"
#   )) %>%
#   mutate(tcell_cat_interface = factor(tcell_cat_interface, levels=c("No interface", "Low T cells-Low Exhaustion", "Low T cells-High Exhaustion", "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#   
# #cytotoxic
#      mutate(tcell_cyto_cat = case_when (
#             `cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted == "Above Median"     ~ "Low T cells-High Exhaustion",
#              `cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted == "Below Median"   ~ "Low T cells-Low Exhaustion",
#             `cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted == "Above Median"     ~ "High T cells-High Exhaustion",
#             `cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted == "Below Median"     ~ "High T cells-Low Exhaustion"
#             
#   )) %>%
#   mutate(tcell_cyto_cat = factor(tcell_cyto_cat, levels=c("Low T cells-Low Exhaustion", "Low T cells-High Exhaustion", 
#                                                 "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#   mutate(tcell_cyto_cat_stroma = case_when (
#             `Stroma cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted_stroma == "Above Median"   ~ "Low T cells-High Exhaustion",
#              `Stroma cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted_stroma == "Below Median"   ~ "Low T cells-Low Exhaustion",
#             `Stroma cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted_stroma == "Above Median"     ~ "High T cells-High Exhaustion",
#             `Stroma cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted_stroma == "Below Median"     ~ "High T cells-Low Exhaustion"
#             
#   )) %>%
#   mutate(tcell_cyto_cat_stroma = factor(tcell_cyto_cat_stroma, levels=c("Low T cells-Low Exhaustion", "Low T cells-High Exhaustion",  "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#     mutate(tcell_cyto_cat_interface = case_when (
#       interface_percent <1 ~ "No interface",
#       `Interface cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted_interface == "Above Median" ~ "Low T cells-High Exhaustion",
#       `Interface cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted_interface == "Below Median" ~ "Low T cells-Low Exhaustion",
#       `Interface cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted_interface == "Above Median" ~ "High T cells-High Exhaustion",
#       `Interface cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted_interface == "Below Median" ~ "High T cells-Low Exhaustion"
#   )) %>%
#   mutate(tcell_cat_interface = factor(tcell_cat_interface, levels=c("No interface", "Low T cells-Low Exhaustion", "Low T cells-High Exhaustion", "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#   
#   
# ## T cell Exhaustion Category - MEDIAN (HIGH TCELL samples only)
#   mutate(high_tcell_exhausted = case_when(
#     grepl("High", `cd3 high/low`)  ~      ifelse(tumor_tcell_timexhausted > 
#                                        median(tumor_tcell_timexhausted[grepl("High", `cd3 high/low`)], na.rm = TRUE), 
#                                            "Above Median", "Below Median")), 
#         high_tcell_exhausted_stroma = case_when(
#           grepl("High", `Stroma cd3 high/low`) ~      ifelse(stroma_tcell_timexhausted > 
#                                        median(stroma_tcell_timexhausted[grepl("High", `Stroma cd3 high/low`)], na.rm = TRUE), 
#                                            "Above Median", "Below Median")),
#         high_tcell_cyto_exhausted = case_when(
#           grepl("High", `cd3+ cd8+ high/low`)  ~      ifelse(tumor_tcell_timexhausted > 
#                                        median(tumor_tcell_timexhausted[grepl("High", `cd3+ cd8+ high/low`)], na.rm = TRUE), 
#                                            "Above Median", "Below Median")),
#         high_tcell_cyto_exhausted_stroma = case_when(
#           grepl("High", `Stroma cd3+ cd8+ high/low`) ~      ifelse(stroma_tcell_timexhausted > 
#                                        median(stroma_tcell_timexhausted[grepl("High", `Stroma cd3+ cd8+ high/low`)], na.rm = TRUE), 
#                                            "Above Median", "Below Median")) ) %>%
# ## T CELL CATEGORY - MEDIANS v2
#   mutate(tcell_cat2 = case_when (
#             `cd3 high/low` == "Low (<=2%)"   ~ "Low T cells",
#             `cd3 high/low` != "Low (<=2%)" & high_tcell_exhausted == "Above Median"     ~ "High T cells-High Exhaustion",
#             `cd3 high/low` != "Low (<=2%)" & high_tcell_exhausted == "Below Median"     ~ "High T cells-Low Exhaustion"),
#           tcell_cat2 = factor(tcell_cat2, levels=c("Low T cells", "High T cells-Low Exhaustion", "High T cells-High Exhaustion")),
#           tcell_cat_stroma2 = case_when (
#               `Stroma cd3 high/low` == "Low (<=2%)" ~ "Low T cells",
#               `Stroma cd3 high/low` != "Low (<=2%)" & high_tcell_exhausted_stroma == "Above Median" ~ "High T cells-High Exhaustion",
#               `Stroma cd3 high/low` != "Low (<=2%)" & high_tcell_exhausted_stroma == "Below Median" ~ "High T cells-Low Exhaustion"),
#           tcell_cat_stroma2 = factor(tcell_cat_stroma2, levels=c("Low T cells", "High T cells-Low Exhaustion", "High T cells-High Exhaustion")),
#           tcell_cyto_cat2 = case_when (
#             `cd3+ cd8+ high/low` == "Low (<=1%)"   ~ "Low T cells",
#             `cd3+ cd8+ high/low` != "Low (<=1%)" & high_tcell_cyto_exhausted == "Above Median"     ~ "High T cells-High Exhaustion",
#             `cd3+ cd8+ high/low` != "Low (<=1%)" & high_tcell_cyto_exhausted == "Below Median"     ~ "High T cells-Low Exhaustion"),
#           tcell_cyto_cat2 = factor(tcell_cyto_cat2, levels=c("Low T cells","High T cells-Low Exhaustion", "High T cells-High Exhaustion")),
#           tcell_cyto_cat_stroma2 = case_when (
#             `Stroma cd3+ cd8+ high/low` == "Low (<=1%)"    ~ "Low T cells",
#             `Stroma cd3+ cd8+ high/low` != "Low (<=1%)" & high_tcell_cyto_exhausted_stroma == "Above Median"     ~ "High T cells-High Exhaustion",
#             `Stroma cd3+ cd8+ high/low` != "Low (<=1%)" & high_tcell_cyto_exhausted_stroma == "Below Median"     ~ "High T cells-Low Exhaustion"),
#             tcell_cyto_cat_stroma2 = factor(tcell_cyto_cat_stroma2, 
#                                          levels=c("Low T cells", "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#   mutate(tumor_stroma_tcellpheno = case_when(tcell_cat2=="Low T cells" & tcell_cat_stroma2 != "Low T cells" ~ "Immune excluded",
#                                                     tcell_cat2=="Low T cells" | tcell_cat_stroma2 == "Low T cells"  ~ "Low T cells",
#                                                     TRUE ~ "High T cells"),
#                 tumor_stroma_exhpheno = case_when(tcell_cat2=="Low T cells" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "Immune excluded-Not Exhausted",
#                                                   tcell_cat2=="Low T cells" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "Immune excluded-Exhausted",
#                                                   tcell_cat2=="Low T cells" | tcell_cat_stroma2 == "Low T cells"  ~ "Low T cells",
#                                                   tcell_cat2=="High T cells-Low Exhaustion" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "Neither Exhausted",
#                                                   tcell_cat2=="High T cells-Low Exhaustion" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "Stroma Exhausted",
#                                                   tcell_cat2=="High T cells-High Exhaustion" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "Tumor Exhausted",
#                                                   tcell_cat2=="High T cells-High Exhaustion" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "Both Exhausted",
#                                                     TRUE ~ NA_character_),
#                  tumor_stroma_exhpheno2 = case_when(tcell_cat2=="Low T cells" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "Immune excluded-Not Exhausted",
#                                                   tcell_cat2=="Low T cells" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "Immune excluded-Exhausted",
#                                                   tcell_cat2=="Low T cells" | tcell_cat_stroma2 == "Low T cells"  ~ "Low T cells",
#                                                   tcell_cat2=="High T cells-Low Exhaustion" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "High T-Not Exhausted",
#                                                   tcell_cat2=="High T cells-Low Exhaustion" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "High T-Exhausted",
#                                                   tcell_cat2=="High T cells-High Exhaustion" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "High T-Exhausted",
#                                                   tcell_cat2=="High T cells-High Exhaustion" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "High T-Exhausted",
#                                                     TRUE ~ NA_character_) ) 
# 
# 
# 
# data.limited <- compartment %>% filter(histotype2=="high-grade serous") %>%
# #reset median cutoff for this dataset distribution
#   mutate(tcell_exhausted =
#           ifelse(tumor_tcell_timexhausted > median(tumor_tcell_timexhausted, na.rm = TRUE), "Above Median", "Below Median")) %>%
#   mutate(tcell_exhausted_stroma = 
#         ifelse(stroma_tcell_timexhausted > median(stroma_tcell_timexhausted, na.rm = TRUE),  "Above Median", "Below Median")) %>%
#   mutate(tcell_exhausted_interface = 
#         ifelse(interface_tcell_timexhausted > median(interface_tcell_timexhausted, na.rm = TRUE),  "Above Median", "Below Median")) %>%
#   mutate(tcell_cyto_exhausted =
#           ifelse(tumor_cyto_tcell_timexhausted > median(tumor_cyto_tcell_timexhausted, na.rm = TRUE), "Above Median", "Below Median")) %>%
#   mutate(tcell_cyto_exhausted_stroma = 
#         ifelse(stroma_cyto_tcell_timexhausted > median(stroma_cyto_tcell_timexhausted, na.rm = TRUE),  "Above Median", "Below Median")) %>%
#   mutate(tcell_cyto_exhausted_interface = 
#         ifelse(interface_cyto_tcell_timexhausted > median(interface_cyto_tcell_timexhausted, na.rm = TRUE),  "Above Median", "Below Median")) %>%
#   
# #exhaustion variable with low t cells by exhaustion  
#   mutate(tcell_cat = case_when (
#             `cd3 high/low` == "Low (<=2%)" & tcell_exhausted == "Above Median"     ~ "Low T cells-High Exhaustion",
#              `cd3 high/low` == "Low (<=2%)" & tcell_exhausted == "Below Median"   ~ "Low T cells-Low Exhaustion",
#             `cd3 high/low` != "Low (<=2%)" & tcell_exhausted == "Above Median"     ~ "High T cells-High Exhaustion",
#             `cd3 high/low` != "Low (<=2%)" & tcell_exhausted == "Below Median"     ~ "High T cells-Low Exhaustion"
#   )) %>%
#   mutate(tcell_cat = factor(tcell_cat, levels=c("Low T cells-Low Exhaustion", "Low T cells-High Exhaustion", 
#                                                 "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#   mutate(tcell_cat_stroma = case_when (
#         `Stroma cd3 high/low` == "Low (<=2%)" & tcell_exhausted_stroma == "Above Median" ~ "Low T cells-High Exhaustion",
#         `Stroma cd3 high/low` == "Low (<=2%)" & tcell_exhausted_stroma == "Below Median" ~ "Low T cells-Low Exhaustion",
#       `Stroma cd3 high/low` != "Low (<=2%)" & tcell_exhausted_stroma == "Above Median" ~ "High T cells-High Exhaustion",
#       `Stroma cd3 high/low` != "Low (<=2%)" & tcell_exhausted_stroma == "Below Median" ~ "High T cells-Low Exhaustion"
#   )) %>%
#   mutate(tcell_cat_stroma = factor(tcell_cat_stroma, levels=c("Low T cells-Low Exhaustion", 
#               "Low T cells-High Exhaustion","High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#   mutate(tcell_cat_interface = case_when (
#       interface_percent <1 ~ "No interface",
#       `Interface cd3 high/low` == "Low (<=2%)" & tcell_exhausted_interface == "Above Median" ~ "Low T cells-High Exhaustion",
#       `Interface cd3 high/low` == "Low (<=2%)" & tcell_exhausted_interface == "Below Median" ~ "Low T cells-Low Exhaustion",
#       `Interface cd3 high/low` != "Low (<=2%)" & tcell_exhausted_interface == "Above Median" ~ "High T cells-High Exhaustion",
#       `Interface cd3 high/low` != "Low (<=2%)" & tcell_exhausted_interface == "Below Median" ~ "High T cells-Low Exhaustion"
#   )) %>%
#   mutate(tcell_cat_interface = factor(tcell_cat_interface, levels=c("No interface", "Low T cells-Low Exhaustion", "Low T cells-High Exhaustion", "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#   
# #cytotoxic
#      mutate(tcell_cyto_cat = case_when (
#             `cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted == "Above Median"     ~ "Low T cells-High Exhaustion",
#              `cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted == "Below Median"   ~ "Low T cells-Low Exhaustion",
#             `cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted == "Above Median"     ~ "High T cells-High Exhaustion",
#             `cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted == "Below Median"     ~ "High T cells-Low Exhaustion"
#             
#   )) %>%
#   mutate(tcell_cyto_cat = factor(tcell_cyto_cat, levels=c("Low T cells-Low Exhaustion", "Low T cells-High Exhaustion", 
#                                                 "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#   mutate(tcell_cyto_cat_stroma = case_when (
#             `Stroma cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted_stroma == "Above Median"   ~ "Low T cells-High Exhaustion",
#              `Stroma cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted_stroma == "Below Median"   ~ "Low T cells-Low Exhaustion",
#             `Stroma cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted_stroma == "Above Median"     ~ "High T cells-High Exhaustion",
#             `Stroma cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted_stroma == "Below Median"     ~ "High T cells-Low Exhaustion"
#             
#   )) %>%
#   mutate(tcell_cyto_cat_stroma = factor(tcell_cyto_cat_stroma, levels=c("Low T cells-Low Exhaustion", "Low T cells-High Exhaustion",  "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#     mutate(tcell_cyto_cat_interface = case_when (
#       interface_percent <1 ~ "No interface",
#       `Interface cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted_interface == "Above Median" ~ "Low T cells-High Exhaustion",
#       `Interface cd3+ cd8+ high/low` == "Low (<=1%)" & tcell_cyto_exhausted_interface == "Below Median" ~ "Low T cells-Low Exhaustion",
#       `Interface cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted_interface == "Above Median" ~ "High T cells-High Exhaustion",
#       `Interface cd3+ cd8+ high/low` != "Low (<=1%)" & tcell_cyto_exhausted_interface == "Below Median" ~ "High T cells-Low Exhaustion"
#   )) %>%
#   mutate(tcell_cat_interface = factor(tcell_cat_interface, levels=c("No interface", "Low T cells-Low Exhaustion", "Low T cells-High Exhaustion", "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#    
# ## T cell Exhaustion Category - MEDIAN (HIGH TCELL samples only)
#   mutate(high_tcell_exhausted = case_when(
#     grepl("High", `cd3 high/low`)  ~      ifelse(tumor_tcell_timexhausted > 
#                                        median(tumor_tcell_timexhausted[grepl("High", `cd3 high/low`)], na.rm = TRUE), 
#                                            "Above Median", "Below Median")), 
#         high_tcell_exhausted_stroma = case_when(
#           grepl("High", `Stroma cd3 high/low`) ~      ifelse(stroma_tcell_timexhausted > 
#                                        median(stroma_tcell_timexhausted[grepl("High", `Stroma cd3 high/low`)], na.rm = TRUE), 
#                                            "Above Median", "Below Median")),
#         high_tcell_cyto_exhausted = case_when(
#           grepl("High", `cd3+ cd8+ high/low`)  ~      ifelse(tumor_tcell_timexhausted > 
#                                        median(tumor_tcell_timexhausted[grepl("High", `cd3+ cd8+ high/low`)], na.rm = TRUE), 
#                                            "Above Median", "Below Median")),
#         high_tcell_cyto_exhausted_stroma = case_when(
#           grepl("High", `Stroma cd3+ cd8+ high/low`) ~      ifelse(stroma_tcell_timexhausted > 
#                                        median(stroma_tcell_timexhausted[grepl("High", `Stroma cd3+ cd8+ high/low`)], na.rm = TRUE), 
#                                            "Above Median", "Below Median")) ) %>%
# ## T CELL CATEGORY - MEDIANS v2
#   mutate(tcell_cat2 = case_when (
#             `cd3 high/low` == "Low (<=2%)"   ~ "Low T cells",
#             `cd3 high/low` != "Low (<=2%)" & high_tcell_exhausted == "Above Median"     ~ "High T cells-High Exhaustion",
#             `cd3 high/low` != "Low (<=2%)" & high_tcell_exhausted == "Below Median"     ~ "High T cells-Low Exhaustion"),
#           tcell_cat2 = factor(tcell_cat2, levels=c("Low T cells", "High T cells-Low Exhaustion", "High T cells-High Exhaustion")),
#           tcell_cat_stroma2 = case_when (
#               `Stroma cd3 high/low` == "Low (<=2%)" ~ "Low T cells",
#               `Stroma cd3 high/low` != "Low (<=2%)" & high_tcell_exhausted_stroma == "Above Median" ~ "High T cells-High Exhaustion",
#               `Stroma cd3 high/low` != "Low (<=2%)" & high_tcell_exhausted_stroma == "Below Median" ~ "High T cells-Low Exhaustion"),
#           tcell_cat_stroma2 = factor(tcell_cat_stroma2, levels=c("Low T cells", "High T cells-Low Exhaustion", "High T cells-High Exhaustion")),
#           tcell_cyto_cat2 = case_when (
#             `cd3+ cd8+ high/low` == "Low (<=1%)"   ~ "Low T cells",
#             `cd3+ cd8+ high/low` != "Low (<=1%)" & high_tcell_cyto_exhausted == "Above Median"     ~ "High T cells-High Exhaustion",
#             `cd3+ cd8+ high/low` != "Low (<=1%)" & high_tcell_cyto_exhausted == "Below Median"     ~ "High T cells-Low Exhaustion"),
#           tcell_cyto_cat2 = factor(tcell_cyto_cat2, levels=c("Low T cells","High T cells-Low Exhaustion", "High T cells-High Exhaustion")),
#           tcell_cyto_cat_stroma2 = case_when (
#             `Stroma cd3+ cd8+ high/low` == "Low (<=1%)"    ~ "Low T cells",
#             `Stroma cd3+ cd8+ high/low` != "Low (<=1%)" & high_tcell_cyto_exhausted_stroma == "Above Median"     ~ "High T cells-High Exhaustion",
#             `Stroma cd3+ cd8+ high/low` != "Low (<=1%)" & high_tcell_cyto_exhausted_stroma == "Below Median"     ~ "High T cells-Low Exhaustion"),
#             tcell_cyto_cat_stroma2 = factor(tcell_cyto_cat_stroma2, 
#                                          levels=c("Low T cells", "High T cells-Low Exhaustion", "High T cells-High Exhaustion"))) %>%
#   mutate(tumor_stroma_tcellpheno = case_when(tcell_cat2=="Low T cells" & tcell_cat_stroma2 != "Low T cells" ~ "Immune excluded",
#                                                     tcell_cat2=="Low T cells" | tcell_cat_stroma2 == "Low T cells"  ~ "Low T cells",
#                                                     TRUE ~ "High T cells"),
#                 tumor_stroma_exhpheno = case_when(tcell_cat2=="Low T cells" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "Immune excluded-Not Exhausted",
#                                                   tcell_cat2=="Low T cells" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "Immune excluded-Exhausted",
#                                                   tcell_cat2=="Low T cells" | tcell_cat_stroma2 == "Low T cells"  ~ "Low T cells",
#                                                   tcell_cat2=="High T cells-Low Exhaustion" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "Neither Exhausted",
#                                                   tcell_cat2=="High T cells-Low Exhaustion" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "Stroma Exhausted",
#                                                   tcell_cat2=="High T cells-High Exhaustion" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "Tumor Exhausted",
#                                                   tcell_cat2=="High T cells-High Exhaustion" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "Both Exhausted",
#                                                     TRUE ~ NA_character_),
#                  tumor_stroma_exhpheno2 = case_when(tcell_cat2=="Low T cells" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "Immune excluded-Not Exhausted",
#                                                   tcell_cat2=="Low T cells" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "Immune excluded-Exhausted",
#                                                   tcell_cat2=="Low T cells" | tcell_cat_stroma2 == "Low T cells"  ~ "Low T cells",
#                                                   tcell_cat2=="High T cells-Low Exhaustion" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "High T-Not Exhausted",
#                                                   tcell_cat2=="High T cells-Low Exhaustion" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "High T-Exhausted",
#                                                   tcell_cat2=="High T cells-High Exhaustion" & tcell_cat_stroma2 == "High T cells-Low Exhaustion" ~ "High T-Exhausted",
#                                                   tcell_cat2=="High T cells-High Exhaustion" & tcell_cat_stroma2 == "High T cells-High Exhaustion" ~ "High T-Exhausted",
#                                                     TRUE ~ NA_character_) ) 
# 
# 
# data <- data %>% set_variable_labels(  
#                                              
#                                       tumor_stroma_tcellpheno = "Tumor-Stroma Abundance",
#                                       tumor_stroma_exhpheno = "Tumor-Stroma Exhaustion",
#                                       tumor_stroma_exhpheno2 = "Tumor-Stroma Exhaustion",
#                                       
#                                      tcell_exhausted= "Tumor Exhausted T cell Grouped",
#                                      tcell_exhausted_stroma= "Stroma Exhausted T cell Grouped",
#                                      tcell_exhausted_interface= "Interface Exhausted T cell Grouped",
#                                      tcell_cyto_exhausted= "Tumor Exhausted Cyto T cell Grouped",
#                                      tcell_cyto_exhausted_stroma= "Stroma Exhausted Cyto T cell Grouped",
#                                      tcell_cyto_exhausted_interface= "Interface Exhausted Cyto T cell Grouped",
#                                     
#                                      high_tcell_exhausted= "Tumor Exhausted T cell Grouped, Median v2",
#                                      high_tcell_exhausted_stroma= "Stroma Exhausted T cell Grouped, Median v2",
#                                      high_tcell_exhausted_interface= "Interface Exhausted T cell Grouped, Median v2",
#                                      high_tcell_cyto_exhausted= "Tumor Exhausted Cyto T cell Grouped, Median v2",
#                                      high_tcell_cyto_exhausted_stroma= "Stroma Exhausted Cyto T cell Grouped, Median v2",
#                                      high_tcell_cyto_exhausted_interface= "Interface Exhausted Cyto T cell Grouped, Median v2",
#                                 
#                                      
#                                      tcell_cat = "Tumor T cell Exhaustion (TIM3+)",
#                                      tcell_cyto_cat = "Tumor Cyto T cell Exhaustion (TIM3+)",
#                                      tcell_cat_stroma = "Stroma T cell Exhaustion (TIM3+)",
#                                      tcell_cyto_cat_stroma = "Stroma Cyto T cell Exhaustion (TIM3+)",
#                                      tcell_cat_interface = "Interface T cell Exhaustion (TIM3+)",
#                                      tcell_cyto_cat_interface = "Interface Cyto T cell Exhaustion (TIM3+)",
#                                      
#                                      tcell_cat2 = "Tumor T cell Exhaustion (Median v2)",
#                                      tcell_cyto_cat2 = "Tumor Cyto T cell Exhaustion (Median v2)",
#                                      tcell_cat_stroma2 = "Stroma T cell Exhaustion (Median v2)",
#                                      tcell_cyto_cat_stroma2 = "Stroma Cyto T cell Exhaustion (Median v2)",
#                                      tcell_cat_interface2 = "Interface T cell Exhaustion (Median v2)",
#                                      tcell_cyto_cat_interface2 = "Interface Cyto T cell Exhaustion (Median v2)")
# 
# data.limited <- data.limited %>% set_variable_labels(         
#                                       tumor_stroma_tcellpheno = "Tumor-Stroma Abundance",
#                                       tumor_stroma_exhpheno = "Tumor-Stroma Exhaustion",
#                                       tumor_stroma_exhpheno2 = "Tumor-Stroma Exhaustion",
#                                      tcell_exhausted= "Tumor Exhausted T cell Grouped",
#                                      tcell_exhausted_stroma= "Stroma Exhausted T cell Grouped",
#                                      tcell_exhausted_interface= "Interface Exhausted T cell Grouped",
#                                      tcell_cyto_exhausted= "Tumor Exhausted Cyto T cell Grouped",
#                                      tcell_cyto_exhausted_stroma= "Stroma Exhausted Cyto T cell Grouped",
#                                      tcell_cyto_exhausted_interface= "Interface Exhausted Cyto T cell Grouped",
#                                     
#                                      high_tcell_exhausted= "Tumor Exhausted T cell Grouped, Median v2",
#                                      high_tcell_exhausted_stroma= "Stroma Exhausted T cell Grouped, Median v2",
#                                      high_tcell_exhausted_interface= "Interface Exhausted T cell Grouped, Median v2",
#                                      high_tcell_cyto_exhausted= "Tumor Exhausted Cyto T cell Grouped, Median v2",
#                                      high_tcell_cyto_exhausted_stroma= "Stroma Exhausted Cyto T cell Grouped, Median v2",
#                                      high_tcell_cyto_exhausted_interface= "Interface Exhausted Cyto T cell Grouped, Median v2",
#                                 
#                                      
#                                      tcell_cat = "Tumor T cell Exhaustion (TIM3+)",
#                                      tcell_cyto_cat = "Tumor Cyto T cell Exhaustion (TIM3+)",
#                                      tcell_cat_stroma = "Stroma T cell Exhaustion (TIM3+)",
#                                      tcell_cyto_cat_stroma = "Stroma Cyto T cell Exhaustion (TIM3+)",
#                                      tcell_cat_interface = "Interface T cell Exhaustion (TIM3+)",
#                                      tcell_cyto_cat_interface = "Interface Cyto T cell Exhaustion (TIM3+)",
#                                      
#                                      tcell_cat2 = "Tumor T cell Exhaustion (Median v2)",
#                                      tcell_cyto_cat2 = "Tumor Cyto T cell Exhaustion (Median v2)",
#                                      tcell_cat_stroma2 = "Stroma T cell Exhaustion (Median v2)",
#                                      tcell_cyto_cat_stroma2 = "Stroma Cyto T cell Exhaustion (Median v2)",
#                                      tcell_cat_interface2 = "Interface T cell Exhaustion (Median v2)",
#                                      tcell_cyto_cat_interface2 = "Interface Cyto T cell Exhaustion (Median v2)")
```


```{r}
spatial_log_data %>% 
  select(logit_prop, logit_stroma_prop, cd8p_cd8n_logratio, cd8p_cd8n_stroma_logratio,
         set, refage, stage_cat) %>%
  tbl_summary() %>%
  bold_labels() 
```


## Continuous Models
## A. T cell abundance
### 1. CD3
```{r}
coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_prop  +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_prop")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")

coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_stroma_prop  +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_stroma_prop")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")
```

### 2. CD8
```{r}
coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_prop  + cd8p_cd8n_logratio +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_prop", "cd8p_cd8n_logratio")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")

coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_stroma_prop  + cd8p_cd8n_stroma_logratio +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_stroma_prop", "cd8p_cd8n_stroma_logratio")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")
```

## B. Tumor+Stroma Models
### 1. CD3
```{r}
fitwts <- coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_prop + logit_stroma_prop +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) 
vif(fitwts)
fitwts %>% tbl_regression(exponentiate = TRUE,
                          include = c("logit_prop",  "logit_stroma_prop")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")
```


### 2. CD8
```{r}
fitwts <- coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_cytoprop + logit_stroma_cytoprop +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE)
#vif(fitwts)
fitwts %>% tbl_regression(exponentiate = TRUE,
                          include = c("logit_cytoprop",  "logit_stroma_cytoprop")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")
```

### 3. CD8+/CD8- Ratio
```{r}
coxph(Surv(int_to_event_mos, vital_status_fin) ~  
        cd8p_cd8n_logratio + cd8p_cd8n_stroma_logratio +
        set + refage + stage_cat + histo_grade, 
      data = spatial_log_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("cd8p_cd8n_logratio",  "cd8p_cd8n_stroma_logratio")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")
```


## C. Final abundance multivariable - no interactions
```{r Bretts final model}
fitwts <- coxph(Surv(int_to_event_mos, vital_status_fin) ~ 
                  logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) 
#vif(fitwts)
fitwts %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_prop",  "logit_stroma_prop", 
                             "cd8p_cd8n_logratio", "cd8p_cd8n_stroma_logratio")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")
```

## D. T cell clustering
```{r}
spatial_log_data %>% 
  select(ends_with("_ripley")) %>%
  tbl_summary() %>%
  bold_labels() 
```

### 1. CD3
```{r}
coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_cd3tumor_ripley  +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_cd3tumor_ripley")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")

coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_cd3stroma_ripley  +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_cd3stroma_ripley")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")

coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_cd3tumor_ripley + logit_cd3stroma_ripley+
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_cd3tumor_ripley", "logit_cd3stroma_ripley")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")
```

### 2. CD8
```{r}
coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_cd8tumor_ripley  +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_cd8tumor_ripley")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")

coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_cd8stroma_ripley  +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_cd8stroma_ripley")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")

coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_cd8tumor_ripley + logit_cd8stroma_ripley +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_cd8tumor_ripley", "logit_cd8stroma_ripley")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")
```

## E. Tumor+Stroma Models
```{r}
fitwts <- coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_cd3tumor_ripley + logit_cd3stroma_ripley +
                  logit_cd8tumor_ripley + logit_cd8stroma_ripley +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) 

fitwts %>% tbl_regression(exponentiate = TRUE,
                          include = c("logit_cd3tumor_ripley",  "logit_cd3stroma_ripley",
                                      "logit_cd8tumor_ripley", "logit_cd8stroma_ripley")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")

fitwts <- coxph(Surv(int_to_event_mos, vital_status_fin) ~ logit_cd3tumor_ripley + 
                  logit_cd8tumor_ripley + logit_cd8stroma_ripley +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) 

fitwts %>% tbl_regression(exponentiate = TRUE,
                          include = c("logit_cd3tumor_ripley",  
                                      "logit_cd8tumor_ripley", "logit_cd8stroma_ripley")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")





fitwts <- coxph(Surv(int_to_event_mos, vital_status_fin) ~ tumor_cd3plus_do_ce_k_border_sdhr + 
                  tumor_cd3plus_cd8plus_do_ce_k_border_sdhr + stroma_cd3plus_cd8plus_do_ce_k_border_sdhr +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) 

fitwts %>% tbl_regression(exponentiate = TRUE,
                          include = c("tumor_cd3plus_do_ce_k_border_sdhr",  
                                      "tumor_cd3plus_cd8plus_do_ce_k_border_sdhr", "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")
```





???
```{r}
fitwts <- coxph(Surv(int_to_event_mos, vital_status_fin) ~ 
                  logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio +
                  logit_cd3tumor_ripley + logit_cd3stroma_ripley +
                  logit_cd8tumor_ripley + logit_cd8stroma_ripley +
                  set + refage + stage_cat + histo_grade, 
                data = spatial_log_data, x = TRUE) 
#vif(fitwts)
fitwts %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_prop",  "logit_stroma_prop", 
                             "cd8p_cd8n_logratio", "cd8p_cd8n_stroma_logratio",
                             "logit_cd3tumor_ripley",  "logit_cd3stroma_ripley",
                             "logit_cd8tumor_ripley", "logit_cd8stroma_ripley")) %>% 
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA**")
```


















# Stratify by Immune Infiltration 





<!-- ## T cell abundance /difference by clinical/tumor characteristics -->



<!-- ```{r} -->

<!-- tbl_uvregression( -->
<!--   data, -->
<!--   method = lm, -->
<!--   y = logit_cytoprop, -->
<!--   include = c("stage_cat", "grade_cat", "histotype2", "subtype", "smokever", "smokcurrent" , "famhxbrov", "diab", "cancer", "BMI_recent") -->
<!-- ) %>% modify_caption("Logit Cyto Prop") -->

<!-- tbl_uvregression( -->
<!--   data, -->
<!--   method = lm, -->
<!--   y = logit_stroma_cytoprop, -->
<!--   include = c("stage_cat", "grade_cat", "histotype2", "subtype", "smokever", "smokcurrent" , "famhxbrov", "diab", "cancer", "BMI_recent") -->
<!-- ) %>% modify_caption(" Stroma Logit Cyto Prop") -->

<!-- tbl_uvregression( -->
<!--   data, -->
<!--   method = lm, -->
<!--   y = logit_cytoprop_relative, -->
<!--   include = c("stage_cat", "grade_cat", "histotype2", "subtype", "smokever", "smokcurrent" , "famhxbrov", "diab", "cancer", "BMI_recent") -->
<!-- ) %>% modify_caption("Logit Cyto Prop Relative") -->

<!-- tbl_uvregression( -->
<!--   data, -->
<!--   method = lm, -->
<!--   y = logit_stroma_cytoprop_relative, -->
<!--   include = c("stage_cat", "grade_cat", "histotype2", "subtype", "smokever", "smokcurrent" , "famhxbrov", "diab", "cancer", "BMI_recent") -->
<!-- ) %>% modify_caption(" Stroma Logit Cyto Prop Relative") -->
<!-- ``` -->



# Effect of boundary length on immune abundance or ripley
```{r ggcorrplot}
library(ggcorrplot)

spat_corr <- spatial_data %>%
  select(contains("_percent"),
         -c(contains("tim3"), contains("pd1"), contains("minus")), 
         boundary_length
         )
  # `colnames<-`(str_remove(colnames(.), "l3midavg_") %>% str_remove(., "l3mid"))
  
spat_corr <- cor(spat_corr, use = "pairwise.complete.obs")
spat_corr2 <- spat_corr %>% as_tibble(rownames = "features")

spat_corr3 <- spat_corr2 %>%
  select(features, boundary_length) %>% 
  filter(features != "boundary_length") %>% 
  column_to_rownames(var = "features") %>% 
  as.matrix()

ggcorrplot(spat_corr3, #hc.order = TRUE, 
           method = "circle", 
           # type = "lower", 
           lab = TRUE, 
           title = "Correlation between boundary length and immune abundance",
           show.legend = TRUE, legend.title = "Correlation", show.diag = TRUE,
           lab_col = "darkblue", lab_size = 3, 
           sig.level = 0.05, insig = c("pch", "blank"), pch = 4, pch.col = "black", pch.cex = 10, 
           tl.cex = 10, tl.col = "green", tl.srt = 40,
           digits = 2
)


spat_corr <- spatial_data %>%
  select(ends_with("_do_ce_k_border"),
         -c(contains("tim3"), contains("pd1"), contains("minus")), 
         boundary_length
         )
  # `colnames<-`(str_remove(colnames(.), "l3midavg_") %>% str_remove(., "l3mid"))
  
spat_corr <- cor(spat_corr, use = "pairwise.complete.obs")
spat_corr2 <- spat_corr %>% as_tibble(rownames = "features")

spat_corr3 <- spat_corr2 %>%
  select(features, boundary_length) %>% 
  filter(features != "boundary_length") %>% 
  column_to_rownames(var = "features") %>% 
  as.matrix()

ggcorrplot(spat_corr3, #hc.order = TRUE, 
           method = "circle", 
           # type = "lower", 
           lab = TRUE, 
           title = "Correlation between boundary_length and ripley",
           show.legend = TRUE, legend.title = "Correlation", show.diag = TRUE,
           lab_col = "darkblue", lab_size = 3, 
           sig.level = 0.05, insig = c("pch", "blank"), pch = 4, pch.col = "black", pch.cex = 10, 
           tl.cex = 10, tl.col = "green", tl.srt = 40,
           digits = 2
)
```




# Reverse analysis
## Distribution
```{r}
spatial_data1 %>% 
  select(ends_with("_5cat")) %>% 
  tbl_summary() %>%
  bold_labels() 
```


## HR
```{r}
tbl_uvregression(
  spatial_data1 %>% select(int_to_event_mos, vital_status_fin, 
                           ends_with("_5cat")),
  method = coxph,
  y = Surv(time = int_to_event_mos, event = vital_status_fin),
  exponentiate = TRUE
) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**UVA**")
```

Increased HR in case of clustering of PD1+ cells but it might only mean there is more cells.
```{r fig.height=40}
spatial_data1 %>% 
  select(suid, 
         tumor_cd19plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8plus_pd1plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8minus_pd1plus_do_ce_k_border_sdhr_5cat,
         tumor_cd8plus_do_ce_k_border_sdhr_5cat,
         tumor_pd1plus_do_ce_k_border_sdhr_5cat,
         # stroma_cd19plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8plus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8minus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd3plus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd8plus_do_ce_k_border_5cat,
         tumor_percent_cd19_cells,
         tumor_percent_cd3plus_cd8plus_cells,
         tumor_percent_cd3plus_cd8plus_pd1plus_cells,
         tumor_percent_cd3plus_cd8minus_pd1plus_cells#,
         # tumor_percent_cd8_cells,
         # tumor_percent_pd1_cells
         ) %>% 

  pivot_longer(cols = c(contains("_5cat")), 
               values_to = "cluster_cat",
               names_to = "cluster_type") %>% 

  pivot_longer(cols = c(contains("percent")), 
               values_to = "cell_abundance",
               names_to = "cell_type") %>% 
  mutate(cluster_cell_type = str_remove_all(cluster_type, "tumor_|stroma_|interface_|_do_ce_k_border_sdhr_5cat")) %>% 
  mutate(cell_type_abundance = str_remove_all(cell_type, "tumor_|stroma_|interface_|percent_|_cells")) %>% 
  filter(!is.na(cluster_cat)) %>% 
  

  ggplot(aes(x = cell_abundance, y = cluster_cat, color = cell_type_abundance, fill = cell_type_abundance)) +
  geom_density_ridges(alpha=0.1, scale = 1) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "inside",
        legend.position.inside = c(0.3, 0.4)
  )+
  scale_y_discrete(expand = c(0, 0.1))+
  facet_wrap(~ cluster_cell_type, ncol = 1, scales = "free_x")
```

```{r ridges 5cat, fig.height=20, fig.width=10}
spatial_data1 %>% 
  select(suid, 
         tumor_cd19plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8plus_pd1plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8minus_pd1plus_do_ce_k_border_sdhr_5cat,
         tumor_cd8plus_do_ce_k_border_sdhr_5cat,
         tumor_pd1plus_do_ce_k_border_sdhr_5cat,
         # stroma_cd19plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8plus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8minus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd3plus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd8plus_do_ce_k_border_5cat,
         tumor_percent_cd19_cells,
         tumor_percent_cd3plus_cd8plus_cells,
         tumor_percent_cd3plus_cd8plus_pd1plus_cells,
         tumor_percent_cd3plus_cd8minus_pd1plus_cells#,
         # tumor_percent_cd8_cells,
         # tumor_percent_pd1_cells
         ) %>% 

  pivot_longer(cols = c(contains("_5cat")), 
               values_to = "cluster_cat",
               names_to = "cluster_type") %>% 

  pivot_longer(cols = c(contains("percent")), 
               values_to = "cell_abundance",
               names_to = "cell_type") %>% 
  mutate(cluster_cell_type = str_remove_all(cluster_type, "tumor_|stroma_|interface_|_do_ce_k_border_sdhr_5cat")) %>% 
  mutate(cell_type_abundance = str_remove_all(cell_type, "tumor_|stroma_|interface_|percent_|_cells")) %>% 
  filter(!is.na(cluster_cat)) %>% 
  

  ggplot(aes(x = cell_abundance, y = cluster_cat, color = cell_type_abundance, fill = cell_type_abundance)) +
  geom_density_ridges(alpha=0.1, scale = 1) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "bottom"
  )+
  scale_y_discrete(expand = c(0, 0.1))+
  scale_color_discrete(guide = guide_legend(nrow = 2))+
  facet_wrap(cell_type_abundance ~ cluster_cell_type, ncol = 6, scales = "free_x")
```

## Plot death?
```{r plot death vs abundance-ripley, fig.height=15, fig.width=10}
spatial_data1 %>% 
  select(suid, vital_status_fin,
         tumor_cd19plus_do_ce_k_border,
         tumor_cd3plus_cd8plus_do_ce_k_border,
         tumor_cd3plus_cd8plus_pd1plus_do_ce_k_border,
         tumor_cd3plus_cd8minus_pd1plus_do_ce_k_border,
         tumor_cd8plus_do_ce_k_border,
         tumor_pd1plus_do_ce_k_border,
         tumor_percent_cd19_cells,
         tumor_percent_cd3plus_cd8plus_cells,
         tumor_percent_cd3plus_cd8plus_pd1plus_cells,
         tumor_percent_cd3plus_cd8minus_pd1plus_cells,
         -c(contains("tim3"))) %>% 
  mutate(vital_status_fin = factor(vital_status_fin)) %>% 
  pivot_longer(cols = c(contains("_do_ce_k_border")), 
               values_to = "cluster_dist",
               names_to = "cluster_type") %>% 
  pivot_longer(cols = c(contains("percent")), 
               values_to = "cell_abundance",
               names_to = "cell_type") %>% 
  
  mutate(cluster_cell_type = str_remove_all(cluster_type, "tumor_|stroma_|interface_|_do_ce_k_border")) %>% 
  mutate(cell_type_abundance = str_remove_all(cell_type, "tumor_|stroma_|interface_|_cells")) %>% 
  
  ggplot(aes(y = cell_abundance, x = cluster_dist, color = vital_status_fin)) + 
  geom_point(alpha = 0.5) +
  ggtitle("Death in Immune cell clustering and abundance")+
  scale_x_log10()+ scale_y_log10()+
  labs(x = "Ripley's K", y = "Abundance", color = "Death")+
  facet_wrap(cluster_cell_type ~ cell_type_abundance, ncol = 4, scales = "free", strip.position = c("top"))+
  theme(legend.position = "bottom"
  )
```




# Ratio ripley / abundance
```{r Ratio ripley - abundance}
spatial_data2 <- spatial_data1 %>% 
  mutate(tumor_cd19plus_cluster_ratio = tumor_cd19plus_do_ce_k_border / tumor_percent_cd19_cells,
         tumor_cd3plus_cluster_ratio = tumor_cd3plus_do_ce_k_border / tumor_percent_cd3_cells,
         tumor_cd8plus_cluster_ratio = tumor_cd8plus_do_ce_k_border / tumor_percent_cd8_cells,
         tumor_pd1plus_cluster_ratio = tumor_pd1plus_do_ce_k_border / tumor_percent_pd1_cells,
         tumor_cd3plus_cd8plus_cluster_ratio = tumor_cd3plus_cd8plus_do_ce_k_border / tumor_percent_cd3plus_cd8plus_cells,
         tumor_cd3plus_cd8plus_pd1plus_cluster_ratio = tumor_cd3plus_cd8plus_pd1plus_do_ce_k_border / tumor_percent_cd3plus_cd8plus_pd1plus_cells,
         tumor_cd3plus_cd8minus_cluster_ratio = tumor_cd3plus_cd8minus_do_ce_k_border / tumor_percent_cd3plus_cd8minus_cells,
         tumor_cd3plus_cd8minus_pd1plus_cluster_ratio = tumor_cd3plus_cd8minus_pd1plus_do_ce_k_border / tumor_percent_cd3plus_cd8minus_pd1plus_cells,
         tumor_cd3plus_pd1plus_cluster_ratio = tumor_cd3plus_pd1plus_do_ce_k_border / tumor_percent_cd3plus_pd1plus_cells
         ) %>% 
  mutate(across(c(contains("cluster_ratio")), ~ . / sd(., na.rm = TRUE) , .names = "{.col}_sdhr"))
```

```{r UVA Ratio ripley - abundance}
spatial_data2 %>% select(ends_with("cluster_ratio"),
                         ends_with("cluster_ratio_sdhr")) %>% 
  tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})")) %>% 
  bold_labels()


tbl_uvregression(
  spatial_data2 %>% select(int_to_event_mos, vital_status_fin, 
                           ends_with("cluster_ratio_sdhr")),
  method = coxph,
  y = Surv(time = int_to_event_mos, event = vital_status_fin),
  exponentiate = TRUE
) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**UVA**")
```



















