---
title: "Immune Spatial–Abundance Crosstalk Analysis"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library, echo = FALSE}
library(tidyverse)
library(labelled)
library(ggcorrplot)
library(gtsummary)
library(survminer)
library(survival)
library(car)
library(gt)
theme_gtsummary_compact()
theme_set(theme_classic())
```

```{r load}
path_raw <- fs::path("","Volumes","Peres_Research", "AACES")

spatial_data <- 
  read_rds(paste0(here::here(), 
                  "/data/raw_data",
                  # path_raw, "/Panel 2 Immune Markers/Survival Analysis",
                  "/compartment_spatial_dataset_06Mar2026.rds")) %>% 
  janitor::clean_names()
# immune_subtypes <- 
#   read_csv(paste0(path_raw, 
#                   "/Panel 2 Immune Markers/Clinical data",
#                   "/panel2_bothphases_11072024.csv")) %>% 
#   select(suid, subtype)
# race_dat <- 
#   read_csv(paste0(path_raw, 
#                   "/AACES_NCOCS_data/clinical data/processed_data",
#                   "/case_ctrl_data_2024-08-16.csv")) %>% 
#   select(suid, site, race)
```

```{r fix grade}
spatial_data <- spatial_data %>% 
  # collapse histotype/grade, assume missing grade is HIGH
  mutate(histo_grade = case_when(histotype2=="high-grade serous" ~ histotype2,
                                 grade_cat == "Low grade"   ~ "other, low-grade",
                                 TRUE  ~ "other, high-grade"))
```

```{r Labeling data}
var_label(spatial_data) <- list(refage = "Patient age",
                                diagyear = "Diagnosis Year", 
                                histotype = "Histotype",
                                histotype2 = "Histotype, collapsed", 
                                histo_grade = "Histotype/Grade",
                                # sex = "Sex",
                                # race_cr_src_desc_1	 = "Race (raw)",
                                # race = "Race", ethnicity_src_desc = "Ethnicity",
                                # histology_desc = "Histology",
                                # bmi = "BMI (continuous)",
                                # bmi_cat = "BMI (categorical)",
                                grade_cat = "Grade",
                                new_dblkstat_treat_ca125 = "debulking status",
                                
                                int_to_event_mos = "Follow-up time from interview (months)",
                                vital_status_fin = "Vital Status"#,
                                # os_event = "Deaths"
)
```


```{r}
# interface_cd3plus_cells
# interface_percent_cd3plus_cells
# stroma_cd3_cells
# stroma_percent_cd3_cells
# tumor_cd3_cells
# tumor_percent_cd3_cells
# 
# "interface_tumor_cells"                                    
#  [21] "interface_stroma_cells"                                    "interface_cells"
# 
# "stroma_tumor_cells"                                        "stroma_stroma_cells"                                      
#  [57] "stroma_total_cells"
# 
# "tumor_tumor_cells"                                        
#  [91] "tumor_stroma_cells"                                        "tumor_total_cells"
# 
# [109] "boundary_length"
# 
# [111] "interface_cd3plus_do_ce_k_border"                          
# "tumor_cd3plus_do_ce_k_border"                             
# [129] "stroma_cd3plus_do_ce_k_border"                                   
# [137] "cd3 high/low"                                              "cd3+ cd8+ high/low"                                       
# [139] "Stroma cd3 high/low"                                       "Stroma cd3+ cd8+ high/low"                                
# [141] "Interface cd3 high/low"                                    "Interface cd3+ cd8+ high/low"
```

```{r reorder vars}
spatial_data <- spatial_data %>% 
  select(suid, 
         starts_with("tumor") & ends_with("do_ce_k_border"),
         starts_with("stroma") & ends_with("do_ce_k_border"),
         starts_with("interface") & ends_with("do_ce_k_border"),
         starts_with("tumor") & contains("percent_"),
         starts_with("stroma") & contains("percent_"),
         starts_with("interface") & contains("percent_"),
         
         starts_with("tumor") & ends_with("_cells"),
         tumorkernel_tumorhalo_cells = tumor_tumor_cells,
         tumorkernel_stromahalo_cells = tumor_stroma_cells, 
         tumor_total_cells,
         starts_with("stroma") & ends_with("_cells"),
         stromakernel_tumorhalo_cells = stroma_tumor_cells, 
         stromakernel_stromahalo_cells = stroma_stroma_cells,
         stroma_total_cells,
         starts_with("interface") & ends_with("_cells"),
         interfacekernel_tumorhalo_cells = interface_tumor_cells,
         interfacekernel_stromahalo_cells = interface_stroma_cells, 
         interface_total_cells = interface_cells,
         
         boundary_length,
         
         image_tag,
         
         tumor_cd3_high_low = cd3_high_low, tumor_cd3_cd8_high_low = cd3_cd8_high_low,
         stroma_cd3_high_low, stroma_cd3_cd8_high_low,
         interface_cd3_high_low, interface_cd3_cd8_high_low,
         
         set : ncol(.)
  )
```

# Data available
```{r}
colnames(spatial_data)[2:134]
```

```{r create sdhr vars}
spatial_data <- spatial_data %>% 
  mutate(across(c(ends_with("_high_low"), -contains("cd8")), 
                ~ factor(., levels = 
                           c("Low (<=2%)", "High (>2%)")))) %>% 
  mutate(across(c(ends_with("_high_low") & contains("cd8")), 
                ~ factor(., levels = 
                           c("Low (<=1%)", "High (>1%)")))) %>% 
  mutate(across(c(contains("tumor") & contains("percent"), 
                  contains("stroma") & contains("percent"),
                  contains("tumor") & contains("do_ce_k_border"), 
                  contains("stroma") & contains("do_ce_k_border"),
                  "interface_cd3plus_do_ce_k_border", "interface_cd3plus_cd8plus_do_ce_k_border"
  ), ~ . / sd(., na.rm = TRUE) , .names = "{.col}_sdhr")) %>% 
  mutate(across(c(ends_with("do_ce_k_border_sdhr")), ~ case_when(
    ntile(., 5) == 1 ~ "1",
    ntile(., 5) == 2 ~ "2",
    ntile(., 5) == 3 ~ "3",
    ntile(., 5) == 4 ~ "4",
    ntile(., 5) == 5 ~ "High clustered"),
    .names = "{.col}_5cat"))
```


# Inclusion
```{r}
spatial_data <- spatial_data %>% 
  filter(phase == 1 & #sample_type=="Compartment" & 
           (neoadj_treat != "yes" | is.na(neoadj_treat)))
```

<!-- Inclusion of WTS with more than 2% (tumor_percent_cd3_cells > 2) of T cell in the tumor compartment -->
<!-- ```{r} -->
<!-- spatial_data <- spatial_data %>%  -->
<!--   filter(tumor_percent_cd3_cells > 2) -->
<!-- ``` -->


```{r Brett logit transf}
eps <- 1e-7
spatial_data <- spatial_data %>%
  mutate(
  #        cyto_stroma_z = scale(stroma_cyto_tcell, center = TRUE, scale = TRUE),
  #        cyto_tumor_z =  scale(tumor_cyto_tcell, center = TRUE, scale = TRUE),
  #        cyto_interface_z =  scale(interface_cyto_tcell, center = TRUE, scale = TRUE),
  #        stroma_z = scale(stroma_tcell, center = TRUE, scale = TRUE),
  #        tumor_z =  scale(tumor_tcell, center = TRUE, scale = TRUE),
  #        interface_z =  scale(interface_tcell, center = TRUE, scale = TRUE),
  #        stroma_ex_z = scale(stroma_percent_cd3plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        tumor_ex_z =  scale(tumor_percent_cd3plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        interface_ex_z =  scale(interface_percent_cd3plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        stroma_cytoex_z = scale(stroma_percent_cd3plus_cd8plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        tumor_cytoex_z =  scale(tumor_percent_cd3plus_cd8plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        interface_cytoex_z =  scale(interface_percent_cd3plus_cd8plus_tim3plus_cells, center = TRUE, scale = TRUE)) %>%
  # #same for CD8- subtype for models with CD8- and CD8+ both included - sensitivity analyses
  # mutate(nocyto_tumor_z = scale(tumor_percent_cd3plus_cd8minus_cells, center=TRUE, scale=TRUE ),
  #        tumor_nocytoex_z = scale(tumor_percent_cd3plus_cd8minus_tim3plus_cells, center=TRUE, scale=TRUE)) %>%
  # mutate(nocyto_stroma_z = scale(stroma_percent_cd3plus_cd8minus_cells, center=TRUE, scale=TRUE ),
  #        stroma_nocytoex_z = scale(stroma_percent_cd3plus_cd8minus_tim3plus_cells, center=TRUE, scale=TRUE)) %>%
  # mutate(tumor_noncyto_tcell_timexhausted = (  ifelse(tumor_cd3plus_cd8minus_cells != 0,
  #                                                     (tumor_cd3plus_cd8minus_tim3plus_cells  /
  #                                                        tumor_cd3plus_cd8minus_cells)*100, 0) )) %>%
  # mutate(stroma_noncyto_tcell_timexhausted = (  ifelse(stroma_cd3plus_cd8minus_cells != 0,
  #                                                      (stroma_cd3plus_cd8minus_tim3plus_cells  /
  #                                                         stroma_cd3plus_cd8minus_cells)*100, 0) )) %>%
  # mutate(interface_noncyto_tcell_timexhausted = (  ifelse(interface_cd3plus_cd8minus_cells != 0,
  #                                                         (interface_cd3plus_cd8minus_tim3plus_cells  /
  #                                                            interface_cd3plus_cd8minus_cells)*100, 0) )
  ) %>%
  # scale
  mutate(tumor_cd3plus_do_ce_k_border_scale = scales::rescale(tumor_cd3plus_do_ce_k_border, to=c(0,1)),
         stroma_cd3plus_do_ce_k_border_scale = scales::rescale(stroma_cd3plus_do_ce_k_border, to=c(0,1)),
         tumor_cd3plus_cd8plus_do_ce_k_border_scale = scales::rescale(tumor_cd3plus_cd8plus_do_ce_k_border, to=c(0,1)),
         stroma_cd3plus_cd8plus_do_ce_k_border_scale = scales::rescale(stroma_cd3plus_cd8plus_do_ce_k_border, to=c(0,1))) %>%
  ## RATIOS
  mutate(
         # cd3p_cd3n_ratio = (tumor_cd3_cells + eps) / ((tumor_total_cells - tumor_cd3_cells) + eps),
         # cd3p_cd3n_logratio = log(cd3p_cd3n_ratio),
         # cd3p_cd3n_stroma_ratio = (stroma_cd3_cells + eps) / ((stroma_total_cells - stroma_cd3_cells) + eps),
         # cd3p_cd3n_stroma_logratio = log(cd3p_cd3n_stroma_ratio),
         # cd3p_cd3n_interface_ratio = (interface_cd3plus_cells + eps) / ((interface_total_cells - interface_cd3plus_cells) + eps),
         # cd3p_cd3n_interface_logratio = log(cd3p_cd3n_interface_ratio),
         cd8p_cd8n_ratio = (tumor_cd3plus_cd8plus_cells + eps) / (tumor_cd3plus_cd8minus_cells + eps),
         cd8p_cd8n_logratio = log(cd8p_cd8n_ratio),
         cd8p_cd8n_stroma_ratio = (stroma_cd3plus_cd8plus_cells + eps) / (stroma_cd3plus_cd8minus_cells + eps),
         cd8p_cd8n_stroma_logratio = log(cd8p_cd8n_stroma_ratio),
         #    cd8p_cd8n_interface_ratio = (interface_cd3plus_cd8plus_cells + eps) / (interface_cd3plus_cd8minus_cells + eps),
         #    cd8p_cd8n_interface_logratio = log(cd8p_cd8n_interface_ratio),
         logit_prop = qlogis(pmin(pmax(tumor_percent_cd3_cells/100, eps), 1-eps)),
         logit_stroma_prop = qlogis(pmin(pmax(stroma_percent_cd3_cells/100, eps), 1-eps)),
         #    logit_interface_prop = qlogis(pmin(pmax(interface_percent_cd3plus_cells/100, eps), 1-eps)),
         logit_cytoprop = qlogis(pmin(pmax(tumor_percent_cd3plus_cd8plus_cells/100, eps), 1-eps)),
         logit_stroma_cytoprop = qlogis(pmin(pmax(stroma_percent_cd3plus_cd8plus_cells/100, eps), 1-eps)),
         #    logit_interface_cytoprop = qlogis(pmin(pmax(interface_percent_cd3plus_cd8plus_cells/100, eps), 1-eps)),
         #    logit_cytoprop_relative = qlogis(pmin(pmax(tumor_cyto_tcell_relative/100, eps), 1-eps)),
         #    logit_stroma_cytoprop_relative = qlogis(pmin(pmax(stroma_cyto_tcell_relative/100, eps), 1-eps)),
         #    logit_interface_cytoprop_relative = qlogis(pmin(pmax(interface_cyto_tcell_relative/100, eps), 1-eps)),
         #    tumor_cd3plus_tim3minus_cells = tumor_cd3_cells -  tumor_cd3plus_tim3plus_cells,
         #    stroma_cd3plus_tim3minus_cells = stroma_cd3_cells -  stroma_cd3plus_tim3plus_cells,
         #    interface_cd3plus_tim3minus_cells = interface_cd3plus_cells -  interface_cd3plus_tim3plus_cells,
         #    timp_timn_ratio =  (tumor_cd3plus_tim3plus_cells + eps) / (tumor_cd3plus_tim3minus_cells + eps),
         #    timp_timn_logratio = log(timp_timn_ratio),
         #    timp_timn_stroma_ratio =  (stroma_cd3plus_tim3plus_cells + eps) / (stroma_cd3plus_tim3minus_cells + eps),
         #    timp_timn_stroma_logratio = log(timp_timn_stroma_ratio),
         #    timp_timn_interface_ratio =  (interface_cd3plus_tim3plus_cells + eps) / (interface_cd3plus_tim3minus_cells + eps),
         #    timp_timn_interface_logratio = log(timp_timn_interface_ratio),
         #    logit_exhprop = qlogis(pmin(pmax(tumor_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_stroma_exhprop = qlogis(pmin(pmax(stroma_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_interface_exhprop = qlogis(pmin(pmax(interface_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_exh2prop = qlogis(pmin(pmax(tumor_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_stroma_exh2prop = qlogis(pmin(pmax(stroma_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_interface_exh2prop = qlogis(pmin(pmax(interface_tcell_bothexhausted/100, eps), 1-eps)),
         #   logit_exh2prop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_pd1plus_tim3plus_cells /100, eps), 1-eps)),
         #    logit_stroma_exh2prop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_pd1plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exh2prop_total = qlogis(pmin(pmax(interface_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         # logit_exh2cytoprop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_cd8plus_pd1plus_tim3plus_cells /100, eps), 1-eps)),
         #    logit_stroma_exh2cytoprop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_cd8plus_pd1plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exh2cytoprop_total = qlogis(pmin(pmax(interface_percent_cd3plus_cd8plus_pd1plus_tim3plus_cells/100, eps), 1-eps)),
         # logit_exh2prop_relative = qlogis(pmin(pmax(tumor_pd1_relative/100, eps), 1-eps)),
         #    logit_stroma_exh2prop_relative = qlogis(pmin(pmax(stroma_pd1_relative/100, eps), 1-eps)),
         #    logit_interface_exh2prop_relative = qlogis(pmin(pmax(interface_pd1_relative/100, eps), 1-eps)),
         #   logit_exh2cytoprop_relative = qlogis(pmin(pmax(tumor_pd1_cyto_relative/100, eps), 1-eps)),
         #    logit_stroma_exh2cytoprop_relative = qlogis(pmin(pmax(stroma_pd1_cyto_relative/100, eps), 1-eps)),
         #    logit_interface_exh2cytoprop_relative = qlogis(pmin(pmax(interface_pd1_cyto_relative/100, eps), 1-eps)),
         #    tumor_cd3plus_cd8plus_tim3minus_cells = tumor_cd3plus_cd8plus_cells - tumor_cd3plus_cd8plus_tim3plus_cells,
         #    stroma_cd3plus_cd8plus_tim3minus_cells = stroma_cd3plus_cd8plus_cells - stroma_cd3plus_cd8plus_tim3plus_cells,
         #    interface_cd3plus_cd8plus_tim3minus_cells = interface_cd3plus_cd8plus_cells - interface_cd3plus_cd8plus_tim3plus_cells,
         #    timp_timn_cd8_ratio = (tumor_cd3plus_cd8plus_tim3plus_cells + eps) / (tumor_cd3plus_cd8plus_tim3minus_cells + eps),
         #    timp_timn_cd8_logratio = log(timp_timn_cd8_ratio),
         #    timp_timn_cd8_stroma_ratio = (stroma_cd3plus_cd8plus_tim3plus_cells + eps) / (stroma_cd3plus_cd8plus_tim3minus_cells + eps),
         #    timp_timn_cd8_stroma_logratio = log(timp_timn_cd8_stroma_ratio),
         #    timp_timn_cd8_interface_ratio = (interface_cd3plus_cd8plus_tim3plus_cells + eps) / (interface_cd3plus_cd8plus_tim3minus_cells + eps),
         #    timp_timn_cd8_interface_logratio = log(timp_timn_cd8_interface_ratio),
         # logit_exhprop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_stroma_exhprop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exhprop_total = qlogis(pmin(pmax(interface_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         #   logit_exhcytoprop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_cd8plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_stroma_exhcytoprop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_cd8plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exhcytoprop_total = qlogis(pmin(pmax(interface_percent_cd3plus_cd8plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_exhcytoprop = qlogis(pmin(pmax(tumor_cyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_stroma_exhcytoprop = qlogis(pmin(pmax(stroma_cyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_interface_exhcytoprop = qlogis(pmin(pmax(interface_cyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_exh2cytoprop = qlogis(pmin(pmax(tumor_cyto_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_stroma_exh2cytoprop = qlogis(pmin(pmax(stroma_cyto_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_interface_exh2cytoprop = qlogis(pmin(pmax(interface_cyto_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_exhnocytoprop = qlogis(pmin(pmax(tumor_noncyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_stroma_exhnocytoprop = qlogis(pmin(pmax(stroma_noncyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_interface_exhnocytoprop = qlogis(pmin(pmax(interface_noncyto_tcell_timexhausted/100, eps), 1-eps))


         # logit_cd3tumor_ripley = qlogis(pmin(pmax(tumor_cd3plus_do_ce_k_border_scale, eps), 1-eps)),
         # logit_cd3stroma_ripley = qlogis(pmin(pmax(stroma_cd3plus_do_ce_k_border_scale, eps), 1-eps)),
         # logit_cd8tumor_ripley = qlogis(pmin(pmax(tumor_cd3plus_cd8plus_do_ce_k_border_scale, eps), 1-eps)),
         # logit_cd8stroma_ripley = qlogis(pmin(pmax(stroma_cd3plus_cd8plus_do_ce_k_border_scale, eps), 1-eps))
  ) %>%
  #Paired Differences
  # mutate(cd3_total_diff = logit_prop - logit_stroma_prop,
                     # cd8_total_diff = logit_cytoprop - logit_stroma_cytoprop,
                     # cd8_ratio_diff = cd8p_cd8n_logratio - cd8p_cd8n_stroma_logratio,
                     # exh_total_diff = logit_exhprop_total - logit_stroma_exhprop_total,
                     # exhcd8_total_diff = logit_exhcytoprop_total - logit_stroma_exhcytoprop_total,
                     # exh_ratio_diff = timp_timn_logratio - timp_timn_stroma_logratio,
                     # exhcd8_ratio_diff = timp_timn_cd8_logratio - timp_timn_cd8_stroma_logratio,
                     # exh2_total_diff = logit_exh2prop_total - logit_stroma_exh2prop_total,
                     # exh2cd8_total_diff = logit_exh2cytoprop_total - logit_stroma_exh2cytoprop_total,
                     # exh2_ratio_diff = logit_exh2prop - logit_stroma_exh2prop,
                     # exh2cd8_ratio_diff = logit_exh2cytoprop - logit_stroma_exh2cytoprop) %>%
  # set_variable_labels(cd3_total_diff = "CD3+ Difference (T-S)",
  #                    cd8_total_diff = "CD8+ Difference (T-S)",
  #                    cd8_ratio_diff = "CD8 Ratio Difference (T-S)",
  #                    exh_total_diff = "TIM3+ Difference (T-S)",
  #                    exhcd8_total_diff = "TIM3+CD8+ Difference (T-S)",
  #                    exh_ratio_diff =  "TIM3 Ratio Difference (T-S)",
  #                    exhcd8_ratio_diff = "TIM3 CD8 Ratio Difference (T-S)",
  #                    exh2_total_diff = "TIM3+PD1+ Difference (T-S)",
  #                    exh2cd8_total_diff = "TIM3+PD1+CD8+ Difference (T-S)",
  #                    exh2_ratio_diff = "TIM3+PD1 Ratio Difference (T-S)",
  #                    exh2cd8_ratio_diff = "TIM3+PD1 CD8 Ratio Difference (T-S)") %>%

  set_variable_labels(
                      # tumor_tcell="Tumor T cell",
                      # tumor_cyto_tcell="Tumor Cytotoxic T cell",
                      # tumor_cd8minus_tcell="Tumor CD8- T cell",
                      # tumor_bcell = "Tumor B cell",
                      # tumor_cyto_tcell_relative= "Tumor Cytotoxic T cell (% of T cells)",
                      # tumor_cd8minus_tcell_relative= "Tumor CD8- T cell (% of T cells)",
                      # tcell_exh2_grp = "Tumor TIM3+PD1+ Abundance",
                      # tcell_exh2_stroma_grp = "Stroma TIM3+PD1+ Abundance",
                      # tcell_exh2_interface_grp = "Interface TIM3+PD1+ Abundance",
                      # tcell_cyto_exh2_grp = "Tumor Cyto TIM3+PD1+ Abundance",
                      # tcell_cyto_exh2_stroma_grp = "Stroma Cyto TIM3+PD1+ Abundance",
                      # tcell_cyto_exh2_interface_grp = "Interface Cyto TIM3+PD1+ Abundance",
                      #
                      # stroma_tcell="Stroma T cell",
                      # stroma_cyto_tcell="Stroma Cytotoxic T cell",
                      # stroma_cd8minus_tcell="Stroma CD8- T cell",
                      # stroma_bcell = "Stroma B cell",
                      # stroma_cyto_tcell_relative= "Stroma Cytotoxic T cell (% of T cells)",
                      # stroma_cd8minus_tcell_relative= "Stroma CD8- T cell (% of T cells)",
                      # cd3p_cd3n_logratio = "Tumor CD3+/CD3- Ratio (Log)",
                      # cd3p_cd3n_stroma_logratio = "Stroma CD3+/CD3- Ratio (Log)",
                      # cd3p_cd3n_interface_logratio = "Interface CD3+/CD3- Ratio (Log)",
                      cd8p_cd8n_logratio = "Tumor CD8+/CD8- Ratio (Log)",
                      cd8p_cd8n_stroma_logratio = "Stroma CD8+/CD8- Ratio (Log)",
                      # cd8p_cd8n_interface_logratio = "Interface CD8+/CD8- Ratio (Log)",
                      # timp_timn_logratio = "Tumor TIM+/TIM- Ratio (Log)",
                      # timp_timn_stroma_logratio = "Stroma TIM+/TIM- Ratio (Log)",
                      # timp_timn_interface_logratio = "Interface TIM+/TIM- Ratio (Log)",
                      # timp_timn_cd8_logratio = "Tumor TIM+/TIM- CD8+ Ratio (Log)",
                      # timp_timn_cd8_stroma_logratio = "Stroma TIM+/TIM- CD8+ Ratio (Log)",
                      # timp_timn_cd8_interface_logratio = "Interface TIM+/TIM- CD8+ Ratio (Log)",

                      logit_prop = "Tumor T cell proportion (Logit)",
                      logit_stroma_prop = "Stroma T cell proportion (Logit)",
                      #  logit_interface_prop = "Interface T cell proportion (Logit)",
                      logit_cytoprop = "Tumor CD8+ T cell proportion (Logit)",
                      #  logit_exhprop = "Tumor Exhausted T cell proportion (Logit)",
                      #  logit_exhcytoprop = "Tumor Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_exhnocytoprop = "Tumor Exhausted CD8- T cell proportion (Logit)",
                      logit_stroma_cytoprop = "Stroma CD8+ T cell proportion (Logit)",
                      #  logit_stroma_exhprop ="Stroma Exhausted T cell proportion (Logit)",
                      #  logit_stroma_exhcytoprop = "Stroma Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_stroma_exhnocytoprop = "Stroma Exhausted CD8- T cell proportion (Logit)",
                      #  logit_interface_cytoprop = "Interface CD8+ T cell proportion (Logit)",
                      #  logit_interface_exhprop ="Interface Exhausted T cell proportion (Logit)",
                      #  logit_interface_exhcytoprop = "Interface Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_interface_exhnocytoprop = "Interface Exhausted CD8- T cell proportion (Logit)",
                      #
                      #  logit_exh2prop = "Tumor PD1 Exhausted T cell proportion (Logit)",
                      #  logit_exh2cytoprop = "Tumor PD1 Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_stroma_exh2prop ="Stroma PD1 Exhausted T cell proportion (Logit)",
                      #  logit_stroma_exh2cytoprop = "Stroma PD1 Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_interface_exh2prop ="Interface PD1 Exhausted T cell proportion (Logit)",
                      #  logit_interface_exh2cytoprop = "Interface PD1 Exhausted CD8+ T cell proportion (Logit)",
                      #
                      #  tumor_z="Tumor T cell abundance (1 SD)",
                      #  stroma_z ="Stroma T cell abundance (1 SD)",
                      #  interface_z ="Interface T cell abundance (1 SD)",
                      #  cyto_tumor_z="Tumor CD8+ abundance (1 SD)",
                      #  cyto_stroma_z ="Stroma CD8+ abundance (1 SD)",
                      #  cyto_interface_z ="Interface CD8+ abundance (1 SD)",
                      #  nocyto_tumor_z="Tumor CD8- abundance (1 SD)",
                      #  nocyto_stroma_z ="Stroma CD8- abundance (1 SD)",
                      #  tumor_noncyto_tcell_timexhausted= "Tumor CD8- Exhausted Proportion",
                      #  stroma_noncyto_tcell_timexhausted = "Stroma CD8- Exhausted Proportion",
                      #  interface_noncyto_tcell_timexhausted = "Interface CD8- Exhausted Proportion",
                      #
                      #  interface_tcell="Interface T cell",
                      #  interface_cyto_tcell="Interface Cytotoxic T cell",
                      #  interface_bcell = "Interface B cell",
                      #  interface_cyto_tcell_relative= "Interface Cytotoxic T cell (% of T cells)",
                      #
                      # tumor_tcell_timexhausted= "Tumor Exhausted T cell - TIM3+",
                      # stroma_tcell_timexhausted= "Stroma Exhausted T cell - TIM3+",
                      # interface_tcell_timexhausted= "Interface Exhausted T cell - TIM3+",
                      # tumor_cyto_tcell_timexhausted= "Tumor Exhausted Cyto T cell - TIM3+",
                      # stroma_cyto_tcell_timexhausted= "Stroma Exhausted Cyto T cell - TIM3+",
                      # interface_cyto_tcell_timexhausted= "Interface Exhausted Cyto T cell - TIM3+",
                      #
                      # tumor_tcell_bothexhausted= "Tumor Terminal Exhausted T cell - PD1+TIM3+",
                      # stroma_tcell_bothexhausted= "Stroma Terminal Exhausted T cell - PD1+TIM3+",
                      # interface_tcell_bothexhausted= "Interface Terminal Exhausted T cell - PD1+TIM3+",
                      # tumor_cyto_tcell_bothexhausted= "Tumor Terminal Exhausted Cyto T cell - PD1+TIM3+",
                      # stroma_cyto_tcell_bothexhausted= "Stroma Terminal Exhausted Cyto T cell - PD1+TIM3+",
                      # interface_cyto_tcell_bothexhausted= "Interface Terminal Exhausted Cyto T cell - PD1+TIM3+"
  )
```

# I. Patient characteristics
```{r Patient characteristics}
spatial_data %>% 
  select(refage, diagyear, stage_cat, grade, grade_cat, behavior, 
         histotype, histotype2, histo_grade,
         subtype, new_dblkstat_treat_ca125, vital_status_fin, int_to_event_mos) %>%
  tbl_summary() %>%
  bold_labels() %>%
  modify_header(label ~ "**Patient characteristics**")
```

```{r reset stage_cat}
spatial_data <- spatial_data %>% 
  mutate(stage_cat = case_when(
    is.na(stage_cat)                  ~ "Late", 
    TRUE                              ~ stage_cat)) #%>% 
  # set_variable_labels(stage_cat="Stage")
```

Brett classified patients with unknown stage to the late stage category.
```{r}
spatial_data %>% 
  select(stage_cat) %>%
  tbl_summary() %>%
  bold_labels() 
```

Just checking if i get the same Ns as Brett for high-grade serous histotype
```{r}
hgsoc_data <- spatial_data %>% 
  filter(histotype2=="high-grade serous")

hgsoc_data %>% 
  select(tumor_percent_cd3_cells) %>% 
  tbl_summary() %>%
  bold_labels()
```

# II. Effect of boundary length on immune abundance or ripley
```{r ggcorrplot}
spat_corr <- spatial_data %>%
  select(contains("_percent"),
         -c(contains("tim3"), contains("pd1"), contains("minus"), 
            contains("sdhr")),
         boundary_length
         ) %>% 
  `colnames<-`(str_remove_all(colnames(.), "_percent|_cells"))

spat_corr <- cor(spat_corr, use = "pairwise.complete.obs")
spat_corr2 <- spat_corr %>% as_tibble(rownames = "features")

spat_corr3 <- spat_corr2 %>%
  select(features, boundary_length) %>%
  filter(features != "boundary_length") %>%
  column_to_rownames(var = "features") %>%
  as.matrix()

ggcorrplot(spat_corr3, #hc.order = TRUE,
           method = "circle",
           # type = "lower",
           lab = TRUE,
           title = "Correlation between boundary length and immune abundance",
           show.legend = TRUE, legend.title = "Correlation", show.diag = TRUE,
           lab_col = "darkblue", lab_size = 3,
           sig.level = 0.05, insig = c("pch", "blank"), pch = 4, pch.col = "black", pch.cex = 10,
           tl.cex = 10, tl.col = "green", tl.srt = 40,
           digits = 2
)


spat_corr <- spatial_data %>%
  select(ends_with("_do_ce_k_border"),
         -c(#contains("tim3"), contains("pd1"), contains("minus"),
            contains("sdhr")),
         boundary_length
         ) %>% 
  `colnames<-`(str_remove_all(colnames(.), "_do_ce_k_border"))

spat_corr <- cor(spat_corr, use = "pairwise.complete.obs")
spat_corr2 <- spat_corr %>% as_tibble(rownames = "features")

spat_corr3 <- spat_corr2 %>%
  select(features, boundary_length) %>%
  filter(features != "boundary_length") %>%
  mutate(compartment = str_match(features, "(tumor|stroma|interface)_")[,2]) %>% 
  mutate(features = str_remove_all(features, "tumor_|stroma_|interface_")) %>% 
  pivot_wider(id_cols = features, names_from = compartment, values_from = boundary_length) %>% 
  column_to_rownames(var = "features") %>%
  as.matrix()

ggcorrplot(spat_corr3, #hc.order = TRUE,
           method = "circle",
           # type = "lower",
           lab = TRUE,
           title = "Correlation between boundary length in the 3 compartments and immune clustering",
           show.legend = TRUE, legend.title = "Correlation", show.diag = TRUE,
           lab_col = "darkblue", lab_size = 3,
           sig.level = 0.05, insig = c("pch", "blank"), pch = 4, pch.col = "black", pch.cex = 10,
           tl.cex = 10, tl.col = "green", tl.srt = 40,
           digits = 2
)
```

# III. Immune abundance
## A. Distributions
```{r Quick look at immune abundance}
tbl1 <- spatial_data %>% 
  select(boundary_length) %>% 
  tbl_summary() %>%
  bold_labels()
tbl2 <- spatial_data %>% 
  select(starts_with("tumor_percent_"), 
         starts_with("tumorkernel_"),
         tumor_total_cells, 
         -ends_with("_sdhr")) %>% 
  tbl_summary() %>%
  bold_labels()
tbl3 <- spatial_data %>% 
  select(starts_with("stroma_percent_"),
         starts_with("stromakernel_"),
         stroma_total_cells, 
         -ends_with("_sdhr")) %>% 
  tbl_summary() %>%
  bold_labels()
tbl4 <- spatial_data %>% 
  select(starts_with("interface_percent_"),
         starts_with("interfacekernel_"),
         interface_total_cells) %>% 
  tbl_summary() %>%
  bold_labels()

tbl_stack(list(tbl1, tbl2, tbl3, tbl4), group_header = c("Boundary", "Tumor",
                                                         "Stroma", "Interface")) %>%
  modify_header(label ~ "**Immune abundance**") %>% 
  modify_footnote_header(
    footnote = "**Zero tumor cells from HALO were categorized as stroma using Kernel <br>
    Zero stroma cells from HALO were categorized as tumor using Kernel**",
    columns = all_stat_cols(),
    replace = FALSE
  ) %>% 
  as_gt() %>%
  gt::tab_options(row_group.border.top.color = "black",
                  row_group.border.top.width = 3,
                  row_group.border.bottom.color = "darkgrey",
                  row_group.font.size = 15,
                  row_group.font.weight = "bold",
                  row_group.text_transform = "uppercase",
                  row_group.background.color = "grey")
```

## B. Final abundance multivariable used in Brett's paper - no interactions
It includes 2 logit transformation for abundance of T cell % in Tumor and T cell % in Stroma, and 2 logratio for Tumor and Stroma CD8+/CD8-.

logit_prop (T cell % in Tumor – logit transformed)   
logit_stroma_prop (T cell % in Stroma – logit transformed)   
cd8p_cd8n_logratio (Tumor CD8+/CD8- ratio – log transformed)   
cd8p_cd8n_stroma_logratio (Stroma CD8+/CD8- ratio – log transformed)   

```{r}
spatial_data %>%
  select(logit_prop, logit_stroma_prop, cd8p_cd8n_logratio, cd8p_cd8n_stroma_logratio,
         set, refage, stage_cat) %>%
  tbl_summary() %>%
  bold_labels()
```

The OS regression model results are as follow:
```{r Bretts final model}
fitwts <- coxph(Surv(int_to_event_mos, vital_status_fin) ~
                  logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio +
                  set + refage + stage_cat + histo_grade,
                data = spatial_data, x = TRUE)
#vif(fitwts)
fitwts %>%
  tbl_regression(exponentiate = TRUE,
                 include = c("logit_prop",  "logit_stroma_prop",
                             "cd8p_cd8n_logratio", "cd8p_cd8n_stroma_logratio")) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**MVA additionally adjusted for <br>
                set + refage + stage_cat + histo_grade**")
```







```{r}
# spatial_data %>% 
#   select(suid, contains("percent_"),
#          -c(contains("tim3"), contains("pd1"), contains("minus"), 
#             ends_with("_sdhr"))) %>% 
#   pivot_longer(cols = -c(suid), 
#                names_to = "immune_cell") %>% 
#   mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
#          compartment = factor(compartment, levels = c("tumor", "stroma", "interface"))) %>% 
#   mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|percent_|_cells|plus")) %>% 
#   mutate(unit = "percent") %>% 
#   ggplot(aes(y = value, x = unit, fill = compartment)) + 
#   geom_violin(alpha = 0.5) +
#   ggtitle("Compartment differences in immune cell abundance")+
#   labs(x = "", y = "Immune cells (%)", fill = "Compartment")+
#   facet_wrap(. ~ cell, scales = "free")
# 
# 
# library(ggridges)
# spatial_data %>% 
#   select(suid, contains("percent_"),
#          -c(contains("tim3"), contains("pd1"), contains("minus"), 
#             ends_with("_sdhr"))) %>% 
#   pivot_longer(cols = -c(suid), 
#                names_to = "immune_cell") %>% 
#   mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
#          compartment = factor(compartment, levels = c("interface", "stroma", "tumor"))) %>% 
#   mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|percent_|_cells|plus")) %>% 
#   
# 
#   ggplot(aes(x = value, y = compartment, fill = cell)) +
#   geom_density_ridges(alpha=0.2, panel_scaling = TRUE) +
#   theme_ridges()+
#   theme(#axis.title.y = element_blank(),
#         #axis.text.y = element_blank(),
#         panel.spacing = unit(0.1, "lines"),
#         # strip.text.x = element_text(size = 8),
#         legend.title = element_blank(),
#         legend.position = "bottom"
#   )+
#   scale_y_discrete(expand = c(0, 0.1))+
#   facet_wrap(~ cell, nrow = 1, scales = "free_x")
#   
# 
# spatial_data %>% 
#   select(suid, contains("percent_"),
#          -c(contains("tim3"), contains("pd1"), contains("minus"), 
#          ends_with("_sdhr"))) %>% 
#   pivot_longer(cols = -c(suid), 
#                names_to = "immune_cell") %>% 
#   mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
#          compartment = factor(compartment, levels = c("interface", "stroma", "tumor"))) %>% 
#   mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|percent_|_cells|plus")) %>% 
#   ggplot(aes(x = value, y = compartment, fill = cell)) +
#   geom_density_ridges(alpha=0.2, panel_scaling = TRUE) +
#   theme_ridges()+
#   theme(#axis.title.y = element_blank(),
#         #axis.text.y = element_blank(),
#         panel.spacing = unit(0.1, "lines"),
#         # strip.text.x = element_text(size = 8),
#         legend.title = element_blank(),
#         legend.position = "bottom"
#   )+
#   xlim(0,40)+
#   scale_y_discrete(expand = c(0, 0.1))+
#   facet_wrap(~ cell, nrow = 1)
```

<br>


# III. Spatial clustering
## A. Distributions by death status
Lower Ripley's (clustering) in patient who died.   
We cannot make any relevant conclusion as abundance needs to be incorporated into the analysis.
```{r Spatial clustering}
tbl1 <- spatial_data %>% 
  select(starts_with("tumor_") & ends_with("_do_ce_k_border"), vital_status_fin) %>% 
  tbl_summary(by = vital_status_fin) %>%
  bold_labels() %>% 
  add_p() %>% bold_p() %>% add_overall() %>% 
  modify_spanning_header(c(stat_1, stat_2) ~ "**Death**")
tbl2 <- spatial_data %>% 
  select(starts_with("stroma_") & ends_with("_do_ce_k_border"), vital_status_fin) %>% 
  tbl_summary(by = vital_status_fin) %>%
  bold_labels() %>% 
  add_p() %>% bold_p() %>% add_overall()
tbl3 <- spatial_data %>% 
  select(starts_with("interface_") & ends_with("_do_ce_k_border"), vital_status_fin) %>% 
  tbl_summary(by = vital_status_fin) %>%
  bold_labels() %>% 
  add_p() %>% bold_p() %>% add_overall()

tbl_stack(list(tbl1, tbl2, tbl3), group_header = c("Tumor",
                                                   "Stroma", "Interface")) %>% 
  modify_header(label ~ "**Raw Ripley's K measure**") %>% 
  as_gt() %>%
  gt::tab_options(row_group.border.top.color = "black",
                  row_group.border.top.width = 3,
                  row_group.border.bottom.color = "darkgrey",
                  row_group.font.size = 15,
                  row_group.font.weight = "bold",
                  row_group.text_transform = "uppercase",
                  row_group.background.color = "grey")

```

```{r Spatial clustering 1-SD}
# tbl1 <- spatial_data %>% 
#   select(starts_with("tumor_") & ends_with("_do_ce_k_border_sdhr"), vital_status_fin) %>% 
#   tbl_summary(by = vital_status_fin) %>%
#   bold_labels() %>% 
#   add_p() %>% bold_p() %>% add_overall() %>% 
#   modify_spanning_header(c(stat_1, stat_2) ~ "**Death**")
# tbl2 <- spatial_data %>% 
#   select(starts_with("stroma_") & ends_with("_do_ce_k_border_sdhr"), vital_status_fin) %>% 
#   tbl_summary(by = vital_status_fin) %>%
#   bold_labels() %>% 
#   add_p() %>% bold_p() %>% add_overall()
# tbl3 <- spatial_data %>% 
#   select(starts_with("interface_") & ends_with("_do_ce_k_border_sdhr"), vital_status_fin) %>% 
#   tbl_summary(by = vital_status_fin) %>%
#   bold_labels() %>% 
#   add_p() %>% bold_p() %>% add_overall()
# 
# tbl_stack(list(tbl1, tbl2, tbl3), group_header = c("Tumor",
#                                                    "Stroma", "Interface")) %>% 
#   modify_header(label ~ "**1-SD normalized Ripley's K**") %>% 
#   as_gt() %>%
#   gt::tab_options(row_group.border.top.color = "black",
#                   row_group.border.top.width = 3,
#                   row_group.border.bottom.color = "darkgrey",
#                   row_group.font.size = 15,
#                   row_group.font.weight = "bold",
#                   row_group.text_transform = "uppercase",
#                   row_group.background.color = "grey")

```

```{r}
spatial_data %>% 
  select(suid, ends_with("_do_ce_k_border"),
         -c(contains("tim3"), contains("pd1"), contains("minus"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("tumor", "stroma", "interface"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|_do_ce_k_border|_cells|plus")) %>% 
  mutate(unit = "ripley") %>% 
  ggplot(aes(y = value, x = unit, fill = compartment)) + 
  geom_violin(alpha = 0.5) +
  ggtitle("Compartment differences in immune cell clustering")+
  labs(x = "", y = "Ripley's K", fill = "Compartment")+
  facet_wrap(. ~ cell, scales = "free")


library(ggridges)
spatial_data %>% 
  select(suid, ends_with("_do_ce_k_border"),
         -c(contains("tim3"), contains("pd1"), contains("minus"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("interface", "stroma", "tumor"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|_do_ce_k_border|_cells|plus")) %>% 
  

  ggplot(aes(x = value, y = compartment, fill = cell)) +
  geom_density_ridges(alpha=0.2, panel_scaling = TRUE) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "bottom"
  )+
  scale_y_discrete(expand = c(0, 0.1))+
  facet_wrap(~ cell, nrow = 1, scales = "free_x")
  

spatial_data %>% 
  select(suid, ends_with("_do_ce_k_border"),
         -c(contains("tim3"), contains("pd1"), contains("minus"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("interface", "stroma", "tumor"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|_do_ce_k_border|_cells|plus")) %>% 
  ggplot(aes(x = value, y = compartment, fill = cell)) +
  geom_density_ridges(alpha=0.2, panel_scaling = TRUE) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "bottom"
  )+
  scale_x_continuous(limits = c(-0.00001,3e7), 
                     labels = function(x) format(x, scientific = TRUE))+
  scale_y_discrete(expand = c(0, 0.1))+
  facet_wrap(~ cell, nrow = 1)
```

## B. HR spatial + abundance
In an analysis including each separate spatial factor additionally adjusted for   
logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio + set + refage + stage_cat + histo_grade;   
We observe a positive association with:   
tumor_cd3plus_cd8minus_do_ce_k_border_sdhr (higher clustering of tumor CD3 cells is associated with worse survival)

We observe a negative association with:   
tumor_cd3plus_pd1plus_do_ce_k_border_sdhr (higher clustering of tumor CD3/PD1 cells is associated with better survival). This can make sense if we think that the exhausted cells are clustered while active CD3 are not. Not all CD3 are exhausted.   
stroma_cd3plus_do_ce_k_border_sdhr	
stroma_cd3plus_cd8plus_do_ce_k_border_sdhr
stroma_cd3plus_cd8minus_do_ce_k_border_sdhr	(higher clustering of stroma CD3 cells is associated with better survival). Biology?
BUT 
higher clustering of stroma CD3/CD8/PD1 cells is associated with better survival
```{r}
# tbl_uvregression(
#   spatial_data %>% select(int_to_event_mos, vital_status_fin, 
#                           contains("tumor") & contains("do_ce_k_border_sdhr"), 
#                           contains("stroma") & contains("do_ce_k_border_sdhr"),
#                           "interface_cd3plus_do_ce_k_border_sdhr", "interface_cd3plus_cd8plus_do_ce_k_border_sdhr"),
#   method = coxph,
#   y = Surv(time = int_to_event_mos, event = vital_status_fin),
#   exponentiate = TRUE#,
#   # include = c("tumor_cd3plus_do_ce_k_border_sdhr", "tumor_cd3plus_cd8plus_do_ce_k_border_sdhr", 
#   #             "stroma_cd3plus_do_ce_k_border_sdhr", "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr", 
#   #             "interface_cd3plus_do_ce_k_border_sdhr", "interface_cd3plus_cd8plus_do_ce_k_border_sdhr"
#   # )
#   ) %>%
#   bold_p(t = .05) %>%
#   add_nevent(location = "level") %>% add_n(location = "level") %>%
#   bold_labels() %>% italicize_levels() %>%
#   modify_header(label ~ "**UVA**")



spat_vars <- spatial_data %>% select(
                          contains("tumor") & contains("do_ce_k_border_sdhr"), 
                          contains("stroma") & contains("do_ce_k_border_sdhr"),
                          -ends_with("_5cat")) %>%
  names()

# Create a list to store the results
cox_mva_fct <- lapply(spat_vars, function(var) {
  # Build the formula dynamically
  formula <- as.formula(paste("Surv(int_to_event_mos, vital_status_fin) ~", 
                              var, 
                              "+ logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio",
                              "+ set + refage + stage_cat + histo_grade"))
  
  # Fit the Cox model
  cox_model <- coxph(formula, data = spatial_data)
  
  # Create a regression table for each model
  tbl_regression(cox_model, exponentiate = TRUE, include = var) %>%
    bold_p(t = 0.05) %>%
    add_nevent(location = "level") %>%
    add_n(location = "level") %>%
    bold_labels() %>%
    italicize_levels()
})
tbl_stack(cox_mva_fct) %>%
  modify_header(label ~ "**Separate MVA additionally adjusted for <br>
                logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio + <br>
                set + refage + stage_cat + histo_grade**") %>% as_gt() %>% 
  gt::tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "tumor_cd3plus_cd8minus_do_ce_k_border_sdhr")
  ) %>% 
  gt::tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "tumor_cd3plus_pd1plus_do_ce_k_border_sdhr")
  ) %>% 
  gt::tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr")
  ) %>% 
  gt::tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "stroma_cd3plus_cd8minus_do_ce_k_border_sdhr")
  )# %>% 
  # gt::tab_style(
  #   style = cell_fill(color = "orange"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "tumor_cd3plus_cd8plus_do_ce_k_border_sdhr")
  # ) %>% 
  # gt::tab_style(
  #   style = cell_fill(color = "orange"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "tumor_cd3plus_cd8minus_pd1plus_do_ce_k_border_sdhr")
  # ) %>% 
  # gt::tab_style(
  #   style = cell_fill(color = "orange"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "stroma_cd3plus_cd8minus_pd1plus_do_ce_k_border_sdhr")
  # ) %>% 
  # gt::tab_style(
  #   style = cell_fill(color = "orange"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "stroma_pd1plus_do_ce_k_border_sdhr")
  # )
```

1) Using the corresponding variable from Brett's model.   
Analysis including the mutually adjusted spatial factor
also adjusted for   
logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio + set + refage + stage_cat + histo_grade
```{r}
mod <- coxph(Surv(int_to_event_mos, vital_status_fin) ~ tumor_cd3plus_do_ce_k_border_sdhr +
        stroma_cd3plus_do_ce_k_border_sdhr +
        tumor_cd3plus_cd8plus_do_ce_k_border_sdhr + stroma_cd3plus_cd8plus_do_ce_k_border_sdhr +
        
        logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio +
        set + refage + stage_cat + histo_grade,
      data = spatial_data, x = TRUE) 
mod %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("tumor_cd3plus_do_ce_k_border_sdhr",
                             "stroma_cd3plus_do_ce_k_border_sdhr",
                             "tumor_cd3plus_cd8plus_do_ce_k_border_sdhr", 
                             "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr")) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**Full MVA**")
```

Checking Variance Inflation Factor (VIF) as factors like CD3 and CD3/CD8 overlap and might be correlated.   
The smallest possible value of VIF is one (absence of multicollinearity). As a rule of thumb, a VIF value that exceeds 5 or 10 indicates a problematic amount of collinearity.
```{r}
# Check VIF
vif_values <- vif(mod)
vif_values
vif_values %>% 
  as_tibble(rownames = "factors") %>% 
  ggplot(aes(x = factors, y = GVIF))+
  geom_bar(stat = "identity", fill = "blue")+
  geom_hline(yintercept = 5, linetype = 2)+
  coord_flip()
```


But the cell types overlap in the above model.   
Below, test 2 models with no overlapping cell types.
```{r}
coxph(Surv(int_to_event_mos, vital_status_fin) ~ tumor_cd3plus_do_ce_k_border_sdhr +
        stroma_cd3plus_do_ce_k_border_sdhr +
        tumor_cd3plus_cd8plus_do_ce_k_border_sdhr + stroma_cd3plus_cd8plus_do_ce_k_border_sdhr +
        
        logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio +
        set + refage + stage_cat + histo_grade,
      data = spatial_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("tumor_cd3plus_do_ce_k_border_sdhr",
                             "stroma_cd3plus_do_ce_k_border_sdhr")) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**Full MVA**")
```

```{r}
coxph(Surv(int_to_event_mos, vital_status_fin) ~ tumor_cd3plus_do_ce_k_border_sdhr +
        stroma_cd3plus_do_ce_k_border_sdhr +
        tumor_cd3plus_cd8plus_do_ce_k_border_sdhr + stroma_cd3plus_cd8plus_do_ce_k_border_sdhr +
        
        logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio +
        set + refage + stage_cat + histo_grade,
      data = spatial_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("tumor_cd3plus_cd8plus_do_ce_k_border_sdhr", 
                             "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr")) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**Full MVA**")
```

Below is a model including the CD8 subpopulation created after chatting with Brett.
```{r}
coxph(Surv(int_to_event_mos, vital_status_fin) ~ tumor_cd3plus_cd8plus_do_ce_k_border_sdhr +
        tumor_cd3plus_cd8minus_do_ce_k_border_sdhr +
        stroma_cd3plus_cd8plus_do_ce_k_border_sdhr +
        stroma_cd3plus_cd8minus_do_ce_k_border_sdhr +
        
        logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio +
        set + refage + stage_cat + histo_grade,
      data = spatial_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("tumor_cd3plus_cd8plus_do_ce_k_border_sdhr", 
                             "tumor_cd3plus_cd8minus_do_ce_k_border_sdhr", 
                             "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr", 
                             "stroma_cd3plus_cd8minus_do_ce_k_border_sdhr")) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**Full MVA**")
```

2) Analysis including the factors associated with survival in an single adjusted analysis, in mutually adjusted MVA additionally adjusted for logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio + set + refage + stage_cat + histo_grade
```{r}
coxph(Surv(int_to_event_mos, vital_status_fin) ~ tumor_cd3plus_cd8minus_do_ce_k_border_sdhr +
        tumor_cd3plus_pd1plus_do_ce_k_border_sdhr +
        stroma_cd3plus_cd8plus_do_ce_k_border_sdhr + stroma_cd3plus_cd8minus_do_ce_k_border_sdhr +
        
        logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio +
        set + refage + stage_cat + histo_grade,
      data = spatial_data, x = TRUE) %>% 
  tbl_regression(exponentiate = TRUE,
                 include = c("tumor_cd3plus_cd8minus_do_ce_k_border_sdhr",
                             "tumor_cd3plus_pd1plus_do_ce_k_border_sdhr",
                             "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr", 
                             "stroma_cd3plus_cd8minus_do_ce_k_border_sdhr")) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>%
  modify_header(label ~ "**Full MVA**")
```
<br>









































# 5 categorical analysis
I did this with the idea to identify non-linear relationship. The issue is that if I make a category based on Ripley's measure for cell type A, often there will be an inverse distribution for another cell type. This is easily visible in the density plots. 
## Distribution
```{r table 5 cat}
spatial_data %>%
  select(ends_with("_5cat")) %>%
  tbl_summary() %>%
  bold_labels()
```


## HR
```{r UVA 5 cat}
# tbl_uvregression(
#   spatial_data %>% select(int_to_event_mos, vital_status_fin,
#                            ends_with("_5cat")),
#   method = coxph,
#   y = Surv(time = int_to_event_mos, event = vital_status_fin),
#   exponentiate = TRUE
# ) %>%
#   bold_p(t = .05) %>%
#   add_nevent(location = "level") %>% add_n(location = "level") %>%
#   bold_labels() %>% italicize_levels() %>%
#   modify_header(label ~ "**UVA**")


spat_vars <- spatial_data %>% 
  select(ends_with("_5cat")) %>%
  names()

# Create a list to store the results
cox_mva_fct <- lapply(spat_vars, function(var) {
  # Build the formula dynamically
  formula <- as.formula(paste("Surv(int_to_event_mos, vital_status_fin) ~", 
                              var, 
                              "+ logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio",
                              "+ set + refage + stage_cat + histo_grade"))
  
  # Fit the Cox model
  cox_model <- coxph(formula, data = spatial_data)
  
  # Create a regression table for each model
  tbl_regression(cox_model, exponentiate = TRUE, include = var) %>%
    bold_p(t = 0.05) %>%
    add_nevent(location = "level") %>%
    add_n(location = "level") %>%
    bold_labels() %>%
    italicize_levels()
})
tbl_stack(cox_mva_fct) %>%
  modify_header(label ~ "**MVA <br>
                logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio + <br>
                set + refage + stage_cat + histo_grade**") %>% as_gt() %>% 
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "tumor_cd3plus_cd8plus_do_ce_k_border_sdhr_5cat")
  ) %>% 
  tab_style(
    style = cell_fill(color = "green"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "tumor_cd3plus_cd8minus_pd1plus_do_ce_k_border_sdhr_5cat")
  ) %>% 
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "stroma_cd3plus_do_ce_k_border_sdhr_5cat")
  ) %>% 
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr_5cat")
  ) %>% 
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "stroma_cd3plus_cd8plus_pd1plus_do_ce_k_border_sdhr_5cat")
  ) %>% 
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "stroma_cd3plus_cd8minus_do_ce_k_border_sdhr_5cat")
  )
  


```
<br>

In the density plots, the y‑axis categories are defined based on Ripley’s measure. The x‑axis represents cell abundances.   
For example, in the third panel, categories are defined based on Ripley’s measure for CD3⁺/CD8⁺ cells. In Category 1, where CD3⁺/CD8⁺ cells are dispersed, a high abundance of CD3⁺/CD8⁺ cells (light blue) is observed. In contrast, in the highly clustered category, a high abundance of CD19 cells (red) is observed.   
Also, defining categories in this manner does not weight cell abundance. For example, in the third panel, as Ripley’s measure increases, the abundance of the cell type used to define the categories decreases.

```{r fig.height=40}
spatial_data %>%
  select(suid,
         tumor_cd19plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8plus_pd1plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8minus_pd1plus_do_ce_k_border_sdhr_5cat,
         tumor_cd8plus_do_ce_k_border_sdhr_5cat,
         tumor_pd1plus_do_ce_k_border_sdhr_5cat,
         # stroma_cd19plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8plus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8minus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd3plus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd8plus_do_ce_k_border_5cat,
         tumor_percent_cd19_cells,
         tumor_percent_cd3plus_cd8plus_cells,
         tumor_percent_cd3plus_cd8plus_pd1plus_cells,
         tumor_percent_cd3plus_cd8minus_pd1plus_cells#,
         # tumor_percent_cd8_cells,
         # tumor_percent_pd1_cells
         ) %>%

  pivot_longer(cols = c(contains("_5cat")),
               values_to = "cluster_cat",
               names_to = "cluster_type") %>%

  pivot_longer(cols = c(contains("percent")),
               values_to = "cell_abundance",
               names_to = "cell_type") %>%
  mutate(cluster_cell_type = str_remove_all(cluster_type, "tumor_|stroma_|interface_|_do_ce_k_border_sdhr_5cat")) %>%
  mutate(cell_type_abundance = str_remove_all(cell_type, "tumor_|stroma_|interface_|percent_|_cells")) %>%
  filter(!is.na(cluster_cat)) %>%


  ggplot(aes(x = cell_abundance, y = cluster_cat, color = cell_type_abundance, fill = cell_type_abundance)) +
  geom_density_ridges(alpha=0.1, scale = 1) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "inside",
        legend.position.inside = c(0.3, 0.4)
  )+
  scale_y_discrete(expand = c(0, 0.1))+
  facet_wrap(~ cluster_cell_type, ncol = 1, scales = "free_x")
```

```{r ridges 5cat, fig.height=20, fig.width=10}
spatial_data %>%
  select(suid,
         tumor_cd19plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8plus_pd1plus_do_ce_k_border_sdhr_5cat,
         tumor_cd3plus_cd8minus_pd1plus_do_ce_k_border_sdhr_5cat,
         tumor_cd8plus_do_ce_k_border_sdhr_5cat,
         tumor_pd1plus_do_ce_k_border_sdhr_5cat,
         # stroma_cd19plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8plus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd3plus_cd8minus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd3plus_pd1plus_do_ce_k_border_5cat,
         # stroma_cd8plus_do_ce_k_border_5cat,
         tumor_percent_cd19_cells,
         tumor_percent_cd3plus_cd8plus_cells,
         tumor_percent_cd3plus_cd8plus_pd1plus_cells,
         tumor_percent_cd3plus_cd8minus_pd1plus_cells#,
         # tumor_percent_cd8_cells,
         # tumor_percent_pd1_cells
         ) %>%

  pivot_longer(cols = c(contains("_5cat")),
               values_to = "cluster_cat",
               names_to = "cluster_type") %>%

  pivot_longer(cols = c(contains("percent")),
               values_to = "cell_abundance",
               names_to = "cell_type") %>%
  mutate(cluster_cell_type = str_remove_all(cluster_type, "tumor_|stroma_|interface_|_do_ce_k_border_sdhr_5cat")) %>%
  mutate(cell_type_abundance = str_remove_all(cell_type, "tumor_|stroma_|interface_|percent_|_cells")) %>%
  filter(!is.na(cluster_cat)) %>%


  ggplot(aes(x = cell_abundance, y = cluster_cat, color = cell_type_abundance, fill = cell_type_abundance)) +
  geom_density_ridges(alpha=0.1, scale = 1) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "bottom"
  )+
  scale_y_discrete(expand = c(0, 0.1))+
  scale_color_discrete(guide = guide_legend(nrow = 2))+
  facet_wrap(cell_type_abundance ~ cluster_cell_type, ncol = 6, scales = "free_x")
```























