---
title: "Immune Spatial–Abundance Crosstalk Analysis"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library, echo = FALSE}
library(tidyverse)
library(labelled)
library(ggcorrplot)
library(gtsummary)
library(survminer)
library(survival)
library(car)
library(gt)
theme_gtsummary_compact()
theme_set(theme_classic())
```

```{r load}
path_raw <- fs::path("","Volumes","Peres_Research", "AACES")

spatial_data <- 
  read_rds(paste0(here::here(), 
                  "/data/raw_data",
                  # path_raw, "/Panel 2 Immune Markers/Survival Analysis",
                  "/compartment_spatial_dataset_06Mar2026.rds")) %>% 
  janitor::clean_names()
# immune_subtypes <- 
#   read_csv(paste0(path_raw, 
#                   "/Panel 2 Immune Markers/Clinical data",
#                   "/panel2_bothphases_11072024.csv")) %>% 
#   select(suid, subtype)
# race_dat <- 
#   read_csv(paste0(path_raw, 
#                   "/AACES_NCOCS_data/clinical data/processed_data",
#                   "/case_ctrl_data_2024-08-16.csv")) %>% 
#   select(suid, site, race)
```

```{r fix grade}
spatial_data <- spatial_data %>% 
  # collapse histotype/grade, assume missing grade is HIGH
  mutate(histo_grade = case_when(histotype2=="high-grade serous" ~ histotype2,
                                 grade_cat == "Low grade"   ~ "other, low-grade",
                                 TRUE  ~ "other, high-grade")) %>% 
  mutate(stage_cat = case_when(
    is.na(stage_cat)                  ~ "Late", 
    TRUE                              ~ stage_cat))
```

```{r Labeling data}
var_label(spatial_data) <- list(refage = "Patient age",
                                diagyear = "Diagnosis Year", 
                                histotype = "Histotype",
                                histotype2 = "Histotype, collapsed", 
                                histo_grade = "Histotype/Grade",
                                # sex = "Sex",
                                # race_cr_src_desc_1	 = "Race (raw)",
                                # race = "Race", ethnicity_src_desc = "Ethnicity",
                                # histology_desc = "Histology",
                                # bmi = "BMI (continuous)",
                                # bmi_cat = "BMI (categorical)",
                                grade_cat = "Grade",
                                new_dblkstat_treat_ca125 = "debulking status",
                                
                                int_to_event_mos = "Follow-up time from interview (months)",
                                vital_status_fin = "Vital Status"
                                # os_event = "Deaths"
)
```


```{r}
# interface_cd3plus_cells
# interface_percent_cd3plus_cells
# stroma_cd3_cells
# stroma_percent_cd3plus_cells
# tumor_cd3_cells
# tumor_percent_cd3plus_cells
# 
# "interface_tumor_cells"                                    
#  [21] "interface_stroma_cells"                                    "interface_cells"
# 
# "stroma_tumor_cells"                                        "stroma_stroma_cells"                                      
#  [57] "stroma_total_cells"
# 
# "tumor_tumor_cells"                                        
#  [91] "tumor_stroma_cells"                                        "tumor_total_cells"
# 
# [109] "boundary_length"
# 
# [111] "interface_cd3plus_do_ce_k_border"                          
# "tumor_cd3plus_do_ce_k_border"                             
# [129] "stroma_cd3plus_do_ce_k_border"                                   
# [137] "cd3 high/low"                                              "cd3+ cd8+ high/low"                                       
# [139] "Stroma cd3 high/low"                                       "Stroma cd3+ cd8+ high/low"                                
# [141] "Interface cd3 high/low"                                    "Interface cd3+ cd8+ high/low"
```

```{r reorder vars}
spatial_data <- spatial_data %>% 
  select(suid, 
         starts_with("tumor") & ends_with("do_ce_k_border"),
         starts_with("stroma") & ends_with("do_ce_k_border"),
         starts_with("interface") & ends_with("do_ce_k_border"),
         starts_with("tumor") & contains("percent_"),
         starts_with("stroma") & contains("percent_"),
         starts_with("interface") & contains("percent_"),
         
         starts_with("tumor") & ends_with("_cells"),
         tumorkernel_tumorhalo_cells = tumor_tumor_cells,
         tumorkernel_stromahalo_cells = tumor_stroma_cells, 
         tumor_total_cells,
         starts_with("stroma") & ends_with("_cells"),
         stromakernel_tumorhalo_cells = stroma_tumor_cells, 
         stromakernel_stromahalo_cells = stroma_stroma_cells,
         stroma_total_cells,
         starts_with("interface") & ends_with("_cells"),
         interfacekernel_tumorhalo_cells = interface_tumor_cells,
         interfacekernel_stromahalo_cells = interface_stroma_cells, 
         interface_total_cells = interface_cells,
         
         boundary_length,
         
         image_tag,
         
         tumor_cd3_high_low = cd3_high_low, tumor_cd3_cd8_high_low = cd3_cd8_high_low,
         stroma_cd3_high_low, stroma_cd3_cd8_high_low,
         interface_cd3_high_low, interface_cd3_cd8_high_low,
         
         set : ncol(.)
  ) %>% 
  `colnames<-`(str_replace_all(colnames(.), "cd3_", "cd3plus_") %>% 
                 str_replace_all(., "cd8_", "cd8plus_") %>% 
                 str_replace_all(., "pd1_", "pd1plus_"))
```

# Data available
```{r}
colnames(spatial_data)[2:134]
```

```{r create sdhr vars}
spatial_data <- spatial_data %>% 
  mutate(across(c(ends_with("_high_low"), -contains("cd8")), 
                ~ factor(., levels = 
                           c("Low (<=2%)", "High (>2%)")))) %>% 
  mutate(across(c(ends_with("_high_low") & contains("cd8")), 
                ~ factor(., levels = 
                           c("Low (<=1%)", "High (>1%)")))) %>% 
  mutate(across(c(contains("tumor") & contains("percent"), 
                  contains("stroma") & contains("percent"),
                  contains("tumor") & contains("do_ce_k_border"), 
                  contains("stroma") & contains("do_ce_k_border"),
                  "interface_cd3plus_do_ce_k_border", "interface_cd3plus_cd8plus_do_ce_k_border"
  ), ~ . / sd(., na.rm = TRUE) , .names = "{.col}_sdhr")) %>% 
  mutate(across(c(ends_with("do_ce_k_border_sdhr")), ~ case_when(
    ntile(., 5) == 1 ~ "1",
    ntile(., 5) == 2 ~ "2",
    ntile(., 5) == 3 ~ "3",
    ntile(., 5) == 4 ~ "4",
    ntile(., 5) == 5 ~ "High clustered"),
    .names = "{.col}_5cat"))
```


# Inclusion
```{r}
spatial_data <- spatial_data %>% 
  filter(phase == 1 & #sample_type=="Compartment" & 
           (neoadj_treat != "yes" | is.na(neoadj_treat))) %>% 
  filter(histotype2=="high-grade serous")
```

<!-- Inclusion of WTS with more than 2% (tumor_percent_cd3plus_cells > 2) of T cell in the tumor compartment -->
<!-- ```{r} -->
<!-- spatial_data <- spatial_data %>%  -->
<!--   filter(tumor_percent_cd3plus_cells > 2) -->
<!-- ``` -->


```{r Brett logit transf}
eps <- 1e-7
spatial_data <- spatial_data %>%
  mutate(
  #        cyto_stroma_z = scale(stroma_cyto_tcell, center = TRUE, scale = TRUE),
  #        cyto_tumor_z =  scale(tumor_cyto_tcell, center = TRUE, scale = TRUE),
  #        cyto_interface_z =  scale(interface_cyto_tcell, center = TRUE, scale = TRUE),
  #        stroma_z = scale(stroma_tcell, center = TRUE, scale = TRUE),
  #        tumor_z =  scale(tumor_tcell, center = TRUE, scale = TRUE),
  #        interface_z =  scale(interface_tcell, center = TRUE, scale = TRUE),
  #        stroma_ex_z = scale(stroma_percent_cd3plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        tumor_ex_z =  scale(tumor_percent_cd3plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        interface_ex_z =  scale(interface_percent_cd3plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        stroma_cytoex_z = scale(stroma_percent_cd3plus_cd8plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        tumor_cytoex_z =  scale(tumor_percent_cd3plus_cd8plus_tim3plus_cells, center = TRUE, scale = TRUE),
  #        interface_cytoex_z =  scale(interface_percent_cd3plus_cd8plus_tim3plus_cells, center = TRUE, scale = TRUE)) %>%
  # #same for CD8- subtype for models with CD8- and CD8+ both included - sensitivity analyses
  # mutate(nocyto_tumor_z = scale(tumor_percent_cd3plus_cd8minus_cells, center=TRUE, scale=TRUE ),
  #        tumor_nocytoex_z = scale(tumor_percent_cd3plus_cd8minus_tim3plus_cells, center=TRUE, scale=TRUE)) %>%
  # mutate(nocyto_stroma_z = scale(stroma_percent_cd3plus_cd8minus_cells, center=TRUE, scale=TRUE ),
  #        stroma_nocytoex_z = scale(stroma_percent_cd3plus_cd8minus_tim3plus_cells, center=TRUE, scale=TRUE)) %>%
  # mutate(tumor_noncyto_tcell_timexhausted = (  ifelse(tumor_cd3plus_cd8minus_cells != 0,
  #                                                     (tumor_cd3plus_cd8minus_tim3plus_cells  /
  #                                                        tumor_cd3plus_cd8minus_cells)*100, 0) )) %>%
  # mutate(stroma_noncyto_tcell_timexhausted = (  ifelse(stroma_cd3plus_cd8minus_cells != 0,
  #                                                      (stroma_cd3plus_cd8minus_tim3plus_cells  /
  #                                                         stroma_cd3plus_cd8minus_cells)*100, 0) )) %>%
  # mutate(interface_noncyto_tcell_timexhausted = (  ifelse(interface_cd3plus_cd8minus_cells != 0,
  #                                                         (interface_cd3plus_cd8minus_tim3plus_cells  /
  #                                                            interface_cd3plus_cd8minus_cells)*100, 0) )
  ) %>%
  # scale
  mutate(tumor_cd3plus_do_ce_k_border_scale = scales::rescale(tumor_cd3plus_do_ce_k_border, to=c(0,1)),
         stroma_cd3plus_do_ce_k_border_scale = scales::rescale(stroma_cd3plus_do_ce_k_border, to=c(0,1)),
         tumor_cd3plus_cd8plus_do_ce_k_border_scale = scales::rescale(tumor_cd3plus_cd8plus_do_ce_k_border, to=c(0,1)),
         stroma_cd3plus_cd8plus_do_ce_k_border_scale = scales::rescale(stroma_cd3plus_cd8plus_do_ce_k_border, to=c(0,1))) %>%
  ## RATIOS
  mutate(
         # cd3p_cd3n_ratio = (tumor_cd3_cells + eps) / ((tumor_total_cells - tumor_cd3_cells) + eps),
         # cd3p_cd3n_logratio = log(cd3p_cd3n_ratio),
         # cd3p_cd3n_stroma_ratio = (stroma_cd3_cells + eps) / ((stroma_total_cells - stroma_cd3_cells) + eps),
         # cd3p_cd3n_stroma_logratio = log(cd3p_cd3n_stroma_ratio),
         # cd3p_cd3n_interface_ratio = (interface_cd3plus_cells + eps) / ((interface_total_cells - interface_cd3plus_cells) + eps),
         # cd3p_cd3n_interface_logratio = log(cd3p_cd3n_interface_ratio),
         cd8p_cd8n_ratio = (tumor_cd3plus_cd8plus_cells + eps) / (tumor_cd3plus_cd8minus_cells + eps),
         cd8p_cd8n_logratio = log(cd8p_cd8n_ratio),
         cd8p_cd8n_stroma_ratio = (stroma_cd3plus_cd8plus_cells + eps) / (stroma_cd3plus_cd8minus_cells + eps),
         cd8p_cd8n_stroma_logratio = log(cd8p_cd8n_stroma_ratio),
         #    cd8p_cd8n_interface_ratio = (interface_cd3plus_cd8plus_cells + eps) / (interface_cd3plus_cd8minus_cells + eps),
         #    cd8p_cd8n_interface_logratio = log(cd8p_cd8n_interface_ratio),
         logit_prop = qlogis(pmin(pmax(tumor_percent_cd3plus_cells/100, eps), 1-eps)),
         logit_stroma_prop = qlogis(pmin(pmax(stroma_percent_cd3plus_cells/100, eps), 1-eps)),
         #    logit_interface_prop = qlogis(pmin(pmax(interface_percent_cd3plus_cells/100, eps), 1-eps)),
         logit_cytoprop = qlogis(pmin(pmax(tumor_percent_cd3plus_cd8plus_cells/100, eps), 1-eps)),
         logit_stroma_cytoprop = qlogis(pmin(pmax(stroma_percent_cd3plus_cd8plus_cells/100, eps), 1-eps)),
         #    logit_interface_cytoprop = qlogis(pmin(pmax(interface_percent_cd3plus_cd8plus_cells/100, eps), 1-eps)),
         #    logit_cytoprop_relative = qlogis(pmin(pmax(tumor_cyto_tcell_relative/100, eps), 1-eps)),
         #    logit_stroma_cytoprop_relative = qlogis(pmin(pmax(stroma_cyto_tcell_relative/100, eps), 1-eps)),
         #    logit_interface_cytoprop_relative = qlogis(pmin(pmax(interface_cyto_tcell_relative/100, eps), 1-eps)),
         #    tumor_cd3plus_tim3minus_cells = tumor_cd3_cells -  tumor_cd3plus_tim3plus_cells,
         #    stroma_cd3plus_tim3minus_cells = stroma_cd3_cells -  stroma_cd3plus_tim3plus_cells,
         #    interface_cd3plus_tim3minus_cells = interface_cd3plus_cells -  interface_cd3plus_tim3plus_cells,
         #    timp_timn_ratio =  (tumor_cd3plus_tim3plus_cells + eps) / (tumor_cd3plus_tim3minus_cells + eps),
         #    timp_timn_logratio = log(timp_timn_ratio),
         #    timp_timn_stroma_ratio =  (stroma_cd3plus_tim3plus_cells + eps) / (stroma_cd3plus_tim3minus_cells + eps),
         #    timp_timn_stroma_logratio = log(timp_timn_stroma_ratio),
         #    timp_timn_interface_ratio =  (interface_cd3plus_tim3plus_cells + eps) / (interface_cd3plus_tim3minus_cells + eps),
         #    timp_timn_interface_logratio = log(timp_timn_interface_ratio),
         #    logit_exhprop = qlogis(pmin(pmax(tumor_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_stroma_exhprop = qlogis(pmin(pmax(stroma_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_interface_exhprop = qlogis(pmin(pmax(interface_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_exh2prop = qlogis(pmin(pmax(tumor_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_stroma_exh2prop = qlogis(pmin(pmax(stroma_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_interface_exh2prop = qlogis(pmin(pmax(interface_tcell_bothexhausted/100, eps), 1-eps)),
         #   logit_exh2prop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_pd1plus_tim3plus_cells /100, eps), 1-eps)),
         #    logit_stroma_exh2prop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_pd1plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exh2prop_total = qlogis(pmin(pmax(interface_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         # logit_exh2cytoprop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_cd8plus_pd1plus_tim3plus_cells /100, eps), 1-eps)),
         #    logit_stroma_exh2cytoprop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_cd8plus_pd1plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exh2cytoprop_total = qlogis(pmin(pmax(interface_percent_cd3plus_cd8plus_pd1plus_tim3plus_cells/100, eps), 1-eps)),
         # logit_exh2prop_relative = qlogis(pmin(pmax(tumor_pd1_relative/100, eps), 1-eps)),
         #    logit_stroma_exh2prop_relative = qlogis(pmin(pmax(stroma_pd1_relative/100, eps), 1-eps)),
         #    logit_interface_exh2prop_relative = qlogis(pmin(pmax(interface_pd1_relative/100, eps), 1-eps)),
         #   logit_exh2cytoprop_relative = qlogis(pmin(pmax(tumor_pd1_cyto_relative/100, eps), 1-eps)),
         #    logit_stroma_exh2cytoprop_relative = qlogis(pmin(pmax(stroma_pd1_cyto_relative/100, eps), 1-eps)),
         #    logit_interface_exh2cytoprop_relative = qlogis(pmin(pmax(interface_pd1_cyto_relative/100, eps), 1-eps)),
         #    tumor_cd3plus_cd8plus_tim3minus_cells = tumor_cd3plus_cd8plus_cells - tumor_cd3plus_cd8plus_tim3plus_cells,
         #    stroma_cd3plus_cd8plus_tim3minus_cells = stroma_cd3plus_cd8plus_cells - stroma_cd3plus_cd8plus_tim3plus_cells,
         #    interface_cd3plus_cd8plus_tim3minus_cells = interface_cd3plus_cd8plus_cells - interface_cd3plus_cd8plus_tim3plus_cells,
         #    timp_timn_cd8_ratio = (tumor_cd3plus_cd8plus_tim3plus_cells + eps) / (tumor_cd3plus_cd8plus_tim3minus_cells + eps),
         #    timp_timn_cd8_logratio = log(timp_timn_cd8_ratio),
         #    timp_timn_cd8_stroma_ratio = (stroma_cd3plus_cd8plus_tim3plus_cells + eps) / (stroma_cd3plus_cd8plus_tim3minus_cells + eps),
         #    timp_timn_cd8_stroma_logratio = log(timp_timn_cd8_stroma_ratio),
         #    timp_timn_cd8_interface_ratio = (interface_cd3plus_cd8plus_tim3plus_cells + eps) / (interface_cd3plus_cd8plus_tim3minus_cells + eps),
         #    timp_timn_cd8_interface_logratio = log(timp_timn_cd8_interface_ratio),
         # logit_exhprop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_stroma_exhprop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exhprop_total = qlogis(pmin(pmax(interface_percent_cd3plus_tim3plus_cells/100, eps), 1-eps)),
         #   logit_exhcytoprop_total = qlogis(pmin(pmax(tumor_percent_cd3plus_cd8plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_stroma_exhcytoprop_total = qlogis(pmin(pmax(stroma_percent_cd3plus_cd8plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_interface_exhcytoprop_total = qlogis(pmin(pmax(interface_percent_cd3plus_cd8plus_tim3plus_cells/100, eps), 1-eps)),
         #    logit_exhcytoprop = qlogis(pmin(pmax(tumor_cyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_stroma_exhcytoprop = qlogis(pmin(pmax(stroma_cyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_interface_exhcytoprop = qlogis(pmin(pmax(interface_cyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_exh2cytoprop = qlogis(pmin(pmax(tumor_cyto_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_stroma_exh2cytoprop = qlogis(pmin(pmax(stroma_cyto_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_interface_exh2cytoprop = qlogis(pmin(pmax(interface_cyto_tcell_bothexhausted/100, eps), 1-eps)),
         #    logit_exhnocytoprop = qlogis(pmin(pmax(tumor_noncyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_stroma_exhnocytoprop = qlogis(pmin(pmax(stroma_noncyto_tcell_timexhausted/100, eps), 1-eps)),
         #    logit_interface_exhnocytoprop = qlogis(pmin(pmax(interface_noncyto_tcell_timexhausted/100, eps), 1-eps))


         # logit_cd3tumor_ripley = qlogis(pmin(pmax(tumor_cd3plus_do_ce_k_border_scale, eps), 1-eps)),
         # logit_cd3stroma_ripley = qlogis(pmin(pmax(stroma_cd3plus_do_ce_k_border_scale, eps), 1-eps)),
         # logit_cd8tumor_ripley = qlogis(pmin(pmax(tumor_cd3plus_cd8plus_do_ce_k_border_scale, eps), 1-eps)),
         # logit_cd8stroma_ripley = qlogis(pmin(pmax(stroma_cd3plus_cd8plus_do_ce_k_border_scale, eps), 1-eps))
  ) %>%
  #Paired Differences
  # mutate(cd3_total_diff = logit_prop - logit_stroma_prop,
                     # cd8_total_diff = logit_cytoprop - logit_stroma_cytoprop,
                     # cd8_ratio_diff = cd8p_cd8n_logratio - cd8p_cd8n_stroma_logratio,
                     # exh_total_diff = logit_exhprop_total - logit_stroma_exhprop_total,
                     # exhcd8_total_diff = logit_exhcytoprop_total - logit_stroma_exhcytoprop_total,
                     # exh_ratio_diff = timp_timn_logratio - timp_timn_stroma_logratio,
                     # exhcd8_ratio_diff = timp_timn_cd8_logratio - timp_timn_cd8_stroma_logratio,
                     # exh2_total_diff = logit_exh2prop_total - logit_stroma_exh2prop_total,
                     # exh2cd8_total_diff = logit_exh2cytoprop_total - logit_stroma_exh2cytoprop_total,
                     # exh2_ratio_diff = logit_exh2prop - logit_stroma_exh2prop,
                     # exh2cd8_ratio_diff = logit_exh2cytoprop - logit_stroma_exh2cytoprop) %>%
  # set_variable_labels(cd3_total_diff = "CD3+ Difference (T-S)",
  #                    cd8_total_diff = "CD8+ Difference (T-S)",
  #                    cd8_ratio_diff = "CD8 Ratio Difference (T-S)",
  #                    exh_total_diff = "TIM3+ Difference (T-S)",
  #                    exhcd8_total_diff = "TIM3+CD8+ Difference (T-S)",
  #                    exh_ratio_diff =  "TIM3 Ratio Difference (T-S)",
  #                    exhcd8_ratio_diff = "TIM3 CD8 Ratio Difference (T-S)",
  #                    exh2_total_diff = "TIM3+PD1+ Difference (T-S)",
  #                    exh2cd8_total_diff = "TIM3+PD1+CD8+ Difference (T-S)",
  #                    exh2_ratio_diff = "TIM3+PD1 Ratio Difference (T-S)",
  #                    exh2cd8_ratio_diff = "TIM3+PD1 CD8 Ratio Difference (T-S)") %>%

  set_variable_labels(
                      # tumor_tcell="Tumor T cell",
                      # tumor_cyto_tcell="Tumor Cytotoxic T cell",
                      # tumor_cd8minus_tcell="Tumor CD8- T cell",
                      # tumor_bcell = "Tumor B cell",
                      # tumor_cyto_tcell_relative= "Tumor Cytotoxic T cell (% of T cells)",
                      # tumor_cd8minus_tcell_relative= "Tumor CD8- T cell (% of T cells)",
                      # tcell_exh2_grp = "Tumor TIM3+PD1+ Abundance",
                      # tcell_exh2_stroma_grp = "Stroma TIM3+PD1+ Abundance",
                      # tcell_exh2_interface_grp = "Interface TIM3+PD1+ Abundance",
                      # tcell_cyto_exh2_grp = "Tumor Cyto TIM3+PD1+ Abundance",
                      # tcell_cyto_exh2_stroma_grp = "Stroma Cyto TIM3+PD1+ Abundance",
                      # tcell_cyto_exh2_interface_grp = "Interface Cyto TIM3+PD1+ Abundance",
                      #
                      # stroma_tcell="Stroma T cell",
                      # stroma_cyto_tcell="Stroma Cytotoxic T cell",
                      # stroma_cd8minus_tcell="Stroma CD8- T cell",
                      # stroma_bcell = "Stroma B cell",
                      # stroma_cyto_tcell_relative= "Stroma Cytotoxic T cell (% of T cells)",
                      # stroma_cd8minus_tcell_relative= "Stroma CD8- T cell (% of T cells)",
                      # cd3p_cd3n_logratio = "Tumor CD3+/CD3- Ratio (Log)",
                      # cd3p_cd3n_stroma_logratio = "Stroma CD3+/CD3- Ratio (Log)",
                      # cd3p_cd3n_interface_logratio = "Interface CD3+/CD3- Ratio (Log)",
                      cd8p_cd8n_logratio = "Tumor CD8+/CD8- Ratio (Log)",
                      cd8p_cd8n_stroma_logratio = "Stroma CD8+/CD8- Ratio (Log)",
                      # cd8p_cd8n_interface_logratio = "Interface CD8+/CD8- Ratio (Log)",
                      # timp_timn_logratio = "Tumor TIM+/TIM- Ratio (Log)",
                      # timp_timn_stroma_logratio = "Stroma TIM+/TIM- Ratio (Log)",
                      # timp_timn_interface_logratio = "Interface TIM+/TIM- Ratio (Log)",
                      # timp_timn_cd8_logratio = "Tumor TIM+/TIM- CD8+ Ratio (Log)",
                      # timp_timn_cd8_stroma_logratio = "Stroma TIM+/TIM- CD8+ Ratio (Log)",
                      # timp_timn_cd8_interface_logratio = "Interface TIM+/TIM- CD8+ Ratio (Log)",

                      logit_prop = "Tumor T cell proportion (Logit)",
                      logit_stroma_prop = "Stroma T cell proportion (Logit)",
                      #  logit_interface_prop = "Interface T cell proportion (Logit)",
                      logit_cytoprop = "Tumor CD8+ T cell proportion (Logit)",
                      #  logit_exhprop = "Tumor Exhausted T cell proportion (Logit)",
                      #  logit_exhcytoprop = "Tumor Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_exhnocytoprop = "Tumor Exhausted CD8- T cell proportion (Logit)",
                      logit_stroma_cytoprop = "Stroma CD8+ T cell proportion (Logit)",
                      #  logit_stroma_exhprop ="Stroma Exhausted T cell proportion (Logit)",
                      #  logit_stroma_exhcytoprop = "Stroma Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_stroma_exhnocytoprop = "Stroma Exhausted CD8- T cell proportion (Logit)",
                      #  logit_interface_cytoprop = "Interface CD8+ T cell proportion (Logit)",
                      #  logit_interface_exhprop ="Interface Exhausted T cell proportion (Logit)",
                      #  logit_interface_exhcytoprop = "Interface Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_interface_exhnocytoprop = "Interface Exhausted CD8- T cell proportion (Logit)",
                      #
                      #  logit_exh2prop = "Tumor PD1 Exhausted T cell proportion (Logit)",
                      #  logit_exh2cytoprop = "Tumor PD1 Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_stroma_exh2prop ="Stroma PD1 Exhausted T cell proportion (Logit)",
                      #  logit_stroma_exh2cytoprop = "Stroma PD1 Exhausted CD8+ T cell proportion (Logit)",
                      #  logit_interface_exh2prop ="Interface PD1 Exhausted T cell proportion (Logit)",
                      #  logit_interface_exh2cytoprop = "Interface PD1 Exhausted CD8+ T cell proportion (Logit)",
                      #
                      #  tumor_z="Tumor T cell abundance (1 SD)",
                      #  stroma_z ="Stroma T cell abundance (1 SD)",
                      #  interface_z ="Interface T cell abundance (1 SD)",
                      #  cyto_tumor_z="Tumor CD8+ abundance (1 SD)",
                      #  cyto_stroma_z ="Stroma CD8+ abundance (1 SD)",
                      #  cyto_interface_z ="Interface CD8+ abundance (1 SD)",
                      #  nocyto_tumor_z="Tumor CD8- abundance (1 SD)",
                      #  nocyto_stroma_z ="Stroma CD8- abundance (1 SD)",
                      #  tumor_noncyto_tcell_timexhausted= "Tumor CD8- Exhausted Proportion",
                      #  stroma_noncyto_tcell_timexhausted = "Stroma CD8- Exhausted Proportion",
                      #  interface_noncyto_tcell_timexhausted = "Interface CD8- Exhausted Proportion",
                      #
                      #  interface_tcell="Interface T cell",
                      #  interface_cyto_tcell="Interface Cytotoxic T cell",
                      #  interface_bcell = "Interface B cell",
                      #  interface_cyto_tcell_relative= "Interface Cytotoxic T cell (% of T cells)",
                      #
                      # tumor_tcell_timexhausted= "Tumor Exhausted T cell - TIM3+",
                      # stroma_tcell_timexhausted= "Stroma Exhausted T cell - TIM3+",
                      # interface_tcell_timexhausted= "Interface Exhausted T cell - TIM3+",
                      # tumor_cyto_tcell_timexhausted= "Tumor Exhausted Cyto T cell - TIM3+",
                      # stroma_cyto_tcell_timexhausted= "Stroma Exhausted Cyto T cell - TIM3+",
                      # interface_cyto_tcell_timexhausted= "Interface Exhausted Cyto T cell - TIM3+",
                      #
                      # tumor_tcell_bothexhausted= "Tumor Terminal Exhausted T cell - PD1+TIM3+",
                      # stroma_tcell_bothexhausted= "Stroma Terminal Exhausted T cell - PD1+TIM3+",
                      # interface_tcell_bothexhausted= "Interface Terminal Exhausted T cell - PD1+TIM3+",
                      # tumor_cyto_tcell_bothexhausted= "Tumor Terminal Exhausted Cyto T cell - PD1+TIM3+",
                      # stroma_cyto_tcell_bothexhausted= "Stroma Terminal Exhausted Cyto T cell - PD1+TIM3+",
                      # interface_cyto_tcell_bothexhausted= "Interface Terminal Exhausted Cyto T cell - PD1+TIM3+"
  )
```

```{r Labeling data2}
var_label(spatial_data) <- list(refage = "Patient age",
                                diagyear = "Diagnosis Year", 
                                histotype = "Histotype",
                                histotype2 = "Histotype, collapsed", 
                                histo_grade = "Histotype/Grade",
                                # sex = "Sex",
                                # race_cr_src_desc_1	 = "Race (raw)",
                                # race = "Race", ethnicity_src_desc = "Ethnicity",
                                # histology_desc = "Histology",
                                # bmi = "BMI (continuous)",
                                # bmi_cat = "BMI (categorical)",
                                grade_cat = "Grade",
                                new_dblkstat_treat_ca125 = "debulking status",
                                
                                int_to_event_mos = "Follow-up time from interview (months)",
                                vital_status_fin = "Vital Status",
                                # os_event = "Deaths"
                                logit_prop = "Tumor CD3 proportion",
                                logit_stroma_prop = "Stroma CD3 proportion",
                                cd8p_cd8n_logratio = "Tumor CD8+/CD8- Ratio (Log)",
                                cd8p_cd8n_stroma_logratio = "Stroma CD8+/CD8- Ratio (Log)",
                                tumor_cd3plus_do_ce_k_border_sdhr = "Tumor CD3 (K)",
                                tumor_cd3plus_cd8plus_do_ce_k_border_sdhr = "Tumor CD3+/CD8+ (K)",
                                tumor_cd3plus_cd8minus_do_ce_k_border_sdhr = "Tumor CD3+/CD8- (K)",
                                stroma_cd3plus_do_ce_k_border_sdhr = "Stroma CD3 (K)",
                                stroma_cd3plus_cd8plus_do_ce_k_border_sdhr = "Stroma CD3+/CD8+ (K)",
                                stroma_cd3plus_cd8minus_do_ce_k_border_sdhr = "Stroma CD3+/CD8- (K)",
                                tumor_cd3plus_cd8plus_pd1plus_do_ce_k_border_sdhr = "Tumor CD3+/CD8+/PD1+ (K)",
                                tumor_cd3plus_cd8minus_pd1plus_do_ce_k_border_sdhr = "Tumor CD3+/CD8-/PD1+ (K)",
                                tumor_cd3plus_pd1plus_do_ce_k_border_sdhr = "Tumor CD3+/PD1+ (K)",
                                tumor_cd8plus_do_ce_k_border_sdhr = "Tumor CD8+ (K)",
                                tumor_pd1plus_do_ce_k_border_sdhr = "Tumor PD1+ (K)",
                                stroma_cd3plus_cd8plus_pd1plus_do_ce_k_border_sdhr = "Stroma CD3+/CD8+/PD1+ (K)",
                                stroma_cd3plus_cd8minus_pd1plus_do_ce_k_border_sdhr = "Stroma CD3+/CD8-/PD1+ (K)",
                                stroma_cd3plus_pd1plus_do_ce_k_border_sdhr = "Stroma CD3+/PD1+ (K)",
                                stroma_cd8plus_do_ce_k_border_sdhr = "Stroma CD8+ (K)",
                                stroma_pd1plus_do_ce_k_border_sdhr = "Stroma PD1+ (K)"
)
```

# I. Patient characteristics
```{r Patient characteristics}
spatial_data %>% 
  select(refage, diagyear, stage_cat, grade, grade_cat, behavior, 
         histotype, histotype2, histo_grade,
         subtype, new_dblkstat_treat_ca125, vital_status_fin, int_to_event_mos) %>%
  tbl_summary() %>%
  bold_labels() %>%
  modify_header(label ~ "**Patient characteristics**")
```

# II. Immune abundance
## A. Distributions
```{r Quick look at immune abundance}
tbl1 <- spatial_data %>% 
  select(boundary_length) %>% 
  tbl_summary() %>%
  bold_labels()
tbl2 <- spatial_data %>% 
  select(starts_with("tumor_percent_"), 
         starts_with("tumorkernel_"),
         tumor_total_cells, 
         -ends_with("_sdhr")) %>% 
  tbl_summary() %>%
  bold_labels()
tbl3 <- spatial_data %>% 
  select(starts_with("stroma_percent_"),
         starts_with("stromakernel_"),
         stroma_total_cells, 
         -ends_with("_sdhr")) %>% 
  tbl_summary() %>%
  bold_labels()
tbl4 <- spatial_data %>% 
  select(starts_with("interface_percent_"),
         starts_with("interfacekernel_"),
         interface_total_cells) %>% 
  tbl_summary() %>%
  bold_labels()

tbl_stack(list(tbl1, tbl2, tbl3, tbl4), group_header = c("Boundary", "Tumor",
                                                         "Stroma", "Interface")) %>%
  modify_header(label ~ "**Immune abundance**") %>% 
  modify_footnote_header(
    footnote = "**Zero tumor cells from HALO were categorized as stroma using Kernel <br>
    Zero stroma cells from HALO were categorized as tumor using Kernel**",
    columns = all_stat_cols(),
    replace = FALSE
  ) %>% 
  as_gt() %>%
  gt::tab_options(row_group.border.top.color = "black",
                  row_group.border.top.width = 3,
                  row_group.border.bottom.color = "darkgrey",
                  row_group.font.size = 15,
                  row_group.font.weight = "bold",
                  row_group.text_transform = "uppercase",
                  row_group.background.color = "grey")
```

## B. Tumor/Stroma/Interface calls discrepancies between Halo/Kernel
```{r}
spatial_data %>% 
  select(contains("kernel")) %>% 
  tbl_summary() %>%
  bold_labels() %>% 
  modify_header(label ~ "**Raw count**")

spatial_data %>% 
  mutate(across(contains("kernel"), ~ . / tumor_total_cells * 100, 
                .names = "{.col}_percent")) %>% 
  select(contains("kernel") & 
           contains("percent")) %>% 
  tbl_summary() %>%
  bold_labels() %>% 
  modify_header(label ~ "**Count (%)**")

spatial_data %>% 
  mutate(across(contains("kernel"), ~ . / tumor_total_cells * 100, 
                .names = "{.col}_percent")) %>% 
  select(stromakernel_stromahalo_cells_percent,
         interfacekernel_tumorhalo_cells_percent,
         interfacekernel_stromahalo_cells_percent) %>% 
  pivot_longer(cols = everything(), names_to = "Cell_type", values_to = "Percent") %>% 
  ggplot(aes(x = Percent))+
  geom_histogram(binwidth = 10, boundary = 0, fill = "grey")+
  facet_wrap(. ~ Cell_type, ncol = 1, scales = "free")
```

## C. Final abundance multivariable used in Brett's paper - no interactions
It includes 2 logit transformation for abundance of T cell % in Tumor and T cell % in Stroma, and 2 logratio for Tumor and Stroma CD8+/CD8-.

logit_prop (T cell % in Tumor – logit transformed)   
logit_stroma_prop (T cell % in Stroma – logit transformed)   
cd8p_cd8n_logratio (Tumor CD8+/CD8- ratio – log transformed)   
cd8p_cd8n_stroma_logratio (Stroma CD8+/CD8- ratio – log transformed)   

```{r}
spatial_data %>%
  select(logit_prop, logit_stroma_prop, cd8p_cd8n_logratio, cd8p_cd8n_stroma_logratio,
         set, refage, stage_cat) %>%
  tbl_summary() %>%
  bold_labels()
```

<!-- The OS regression model results are as follow: -->
<!-- ```{r Bretts final model, echo = TRUE, class.source = 'fold-hide'} -->
<!-- fitwts <- coxph(Surv(int_to_event_mos, vital_status_fin) ~ -->
<!--                   logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio + -->
<!--                   set + refage + stage_cat + histo_grade, -->
<!--                 data = spatial_data, x = TRUE) -->
<!-- fitwts %>% -->
<!--   tbl_regression(exponentiate = TRUE, -->
<!--                  include = c("logit_prop",  "logit_stroma_prop", -->
<!--                              "cd8p_cd8n_logratio", "cd8p_cd8n_stroma_logratio")) %>% -->
<!--   bold_p(t = .05) %>% -->
<!--   add_nevent(location = "level") %>% add_n(location = "level") %>% -->
<!--   bold_labels() %>% italicize_levels() %>% -->
<!--   modify_header(label ~ "**MVA model include the 4 mutually adjusted Ripley's K features and additionally adjusted for <br> -->
<!--                 set + refage + stage_cat + histo_grade**") -->
<!-- ``` -->

<br>


# III. Spatial clustering
## A. Distributions by death status
Lower Ripley's (clustering) in patient who died.   
We cannot make any relevant conclusion as abundance needs to be incorporated into the analysis.
```{r Spatial clustering}
tbl1 <- spatial_data %>% 
  select(starts_with("tumor_") & ends_with("_do_ce_k_border"), vital_status_fin) %>% 
  tbl_summary(by = vital_status_fin) %>%
  bold_labels() %>% 
  add_p() %>% bold_p() %>% add_overall() %>% 
  modify_spanning_header(c(stat_1, stat_2) ~ "**Death**")
tbl2 <- spatial_data %>% 
  select(starts_with("stroma_") & ends_with("_do_ce_k_border"), vital_status_fin) %>% 
  tbl_summary(by = vital_status_fin) %>%
  bold_labels() %>% 
  add_p() %>% bold_p() %>% add_overall()
tbl3 <- spatial_data %>% 
  select(starts_with("interface_") & ends_with("_do_ce_k_border"), vital_status_fin) %>% 
  tbl_summary(by = vital_status_fin) %>%
  bold_labels() %>% 
  add_p() %>% bold_p() %>% add_overall()

tbl_stack(list(tbl1, tbl2, tbl3), group_header = c("Tumor",
                                                   "Stroma", "Interface")) %>% 
  modify_header(label ~ "**Raw Ripley's K measure**") %>% 
  as_gt() %>%
  gt::tab_options(row_group.border.top.color = "black",
                  row_group.border.top.width = 3,
                  row_group.border.bottom.color = "darkgrey",
                  row_group.font.size = 15,
                  row_group.font.weight = "bold",
                  row_group.text_transform = "uppercase",
                  row_group.background.color = "grey")

```



```{r}
spatial_data %>% 
  select(suid, ends_with("_do_ce_k_border"),
         -c(contains("tim3"), contains("pd1"), 
            contains("cd19"), contains("minus"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("tumor", "stroma", "interface"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|_do_ce_k_border|_cells|plus")) %>% 
  mutate(unit = "ripley") %>% 
  ggplot(aes(y = value, x = unit, fill = compartment)) + 
  geom_violin(alpha = 0.5, bounds = c(0, Inf)) +
  ggtitle("Compartment differences in immune cell clustering")+
  labs(x = "", y = "Ripley's K", fill = "Compartment")+
  facet_wrap(. ~ cell, scales = "free")


library(ggridges)
spatial_data %>% 
  select(suid, ends_with("_do_ce_k_border"),
         -c(contains("tim3"), contains("pd1"), 
            contains("cd19"), contains("minus"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("interface", "stroma", "tumor"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|_do_ce_k_border|_cells|plus")) %>% 
  

  ggplot(aes(x = value, y = compartment, fill = cell)) +
  geom_density_ridges(alpha=0.2, panel_scaling = TRUE) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "bottom"
  )+
  scale_y_discrete(expand = c(0, 0.1))+
  facet_wrap(~ cell, nrow = 1, scales = "free_x")
  

spatial_data %>% 
  select(suid, ends_with("_do_ce_k_border"),
         -c(contains("tim3"), contains("pd1"), 
            contains("cd19"), contains("minus"))) %>% 
  pivot_longer(cols = -c(suid), 
               names_to = "immune_cell") %>% 
  mutate(compartment = str_match(immune_cell, "(tumor|stroma|interface)_*")[,2],
         compartment = factor(compartment, levels = c("interface", "stroma", "tumor"))) %>% 
  mutate(cell = str_remove_all(immune_cell, "tumor_|stroma_|interface_|_do_ce_k_border|_cells|plus")) %>% 
  ggplot(aes(x = value, y = compartment, fill = cell)) +
  geom_density_ridges(alpha=0.2, panel_scaling = TRUE) +
  theme_ridges()+
  theme(#axis.title.y = element_blank(),
        #axis.text.y = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        # strip.text.x = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "bottom"
  )+
  scale_x_continuous(limits = c(-0.00001,3e7), 
                     labels = function(x) format(x, scientific = TRUE))+
  scale_y_discrete(expand = c(0, 0.1))+
  facet_wrap(~ cell, nrow = 1)
```

## B. HR spatial + abundance
Limit to WTS with ≥ 8 cells





### 1. Single Ripley's K feature analysis as covariate in Brett's model
additionally adjusted for `Tumor CD3 proportion` + `Stroma CD3 proportion`  + `Tumor CD8+/CD8- Ratio (Log)`  + `Stroma CD8+/CD8- Ratio (Log)` + set + refage + stage_cat + histo_grade
```{r echo = TRUE, class.source = 'fold-hide'}
spat_vars <- spatial_data %>% select(
                          contains("tumor") & contains("do_ce_k_border_sdhr"), 
                          contains("stroma") & contains("do_ce_k_border_sdhr"),
                          -ends_with("_5cat"), 
                          -tumor_cd19plus_do_ce_k_border_sdhr,
                          -stroma_cd19plus_do_ce_k_border_sdhr) %>%
  names()

# Create a list to store the results
cox_mva_fct <- lapply(spat_vars, function(var) {
  
  cell <- str_remove(var, "_do_ce_k_border_sdhr")
  cell2 <- paste0(cell, "_cell", collapse = "")
  print(cell2)
  spatial_data <- spatial_data %>% 
    filter(cell2 >=8)
  # Build the formula dynamically
  formula <- as.formula(paste("Surv(int_to_event_mos, vital_status_fin) ~", 
                              var, 
                              "+ logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio",
                              "+ set + refage + stage_cat"))
  
  # Fit the Cox model
  cox_model <- coxph(formula, data = spatial_data)
  
  # Create a regression table for each model
  tbl_regression(cox_model, exponentiate = TRUE, include = var) %>%
    bold_p(t = 0.05) %>%
    add_nevent(location = "level") %>%
    add_n(location = "level") %>%
    bold_labels() %>%
    italicize_levels()
})
tbl_stack(cox_mva_fct) %>%
  modify_header(label ~ "**MVA for each separate Ripley's K feature additionally adjusted<br>
                `Tumor CD3 proportion` + `Stroma CD3 proportion`  + `Tumor CD8+/CD8- Ratio (Log)`  + `Stroma CD8+/CD8- Ratio (Log)` + <br>
                set + refage + stage_cat + histo_grade**") #%>% as_gt() %>% 
  # gt::tab_style(
  #   style = cell_fill(color = "yellow"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "tumor_cd3plus_cd8minus_do_ce_k_border_sdhr")
  # ) %>% 
  # gt::tab_style(
  #   style = cell_fill(color = "yellow"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "tumor_cd3plus_pd1plus_do_ce_k_border_sdhr")
  # ) %>% 
  # gt::tab_style(
  #   style = cell_fill(color = "yellow"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr")
  # ) %>% 
  # gt::tab_style(
  #   style = cell_fill(color = "yellow"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "stroma_cd3plus_cd8minus_do_ce_k_border_sdhr")
  # )
```

### 2. Single Ripley's K feature analysis as covariate with corresponding abundance var
additionally adjusted for Tumor/Stroma corresponding cell type + set + refage + stage_cat + histo_grade
```{r echo = TRUE, class.source = 'fold-hide'}
spat_vars <- spatial_data %>% select(
                          contains("tumor") & contains("do_ce_k_border_sdhr"), 
                          contains("stroma") & contains("do_ce_k_border_sdhr"),
                          -ends_with("_5cat"), 
                          -tumor_cd19plus_do_ce_k_border_sdhr,
                          -stroma_cd19plus_do_ce_k_border_sdhr) %>%
  names()

# Create a list to store the results
cox_mva_fct <- lapply(spat_vars, function(var) {
  
  cell <- str_remove(var, "_do_ce_k_border_sdhr")
  compartment <- str_match(cell, "(tumor|stroma|interface)_*")[,2]
  cell_type <- str_remove(cell, "tumor|stroma")
  
  cell2 <- paste0(compartment, "_percent", cell_type, "_cells_sdhr", collapse = "")
  print(cell2)
  spatial_data <- spatial_data %>% 
    filter(cell2 >=8)
  # Build the formula dynamically
  formula <- as.formula(paste("Surv(int_to_event_mos, vital_status_fin) ~", 
                              var, "+", cell2,
                              # "+ logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio",
                              "+ set + refage + stage_cat"))
  
  # Fit the Cox model
  cox_model <- coxph(formula, data = spatial_data)
  
  # Create a regression table for each model
  tbl_regression(cox_model, exponentiate = TRUE, include = var) %>%
    bold_p(t = 0.05) %>%
    add_nevent(location = "level") %>%
    add_n(location = "level") %>%
    bold_labels() %>%
    italicize_levels()
})
tbl_stack(cox_mva_fct) %>%
  modify_header(label ~ "**MVA for each separate Ripley's K feature additionally adjusted<br>
                `Percent abundance of the corresponding cell type` + <br>
                set + refage + stage_cat + histo_grade**") %>% as_gt() %>% 
  gt::tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "tumor_cd3plus_cd8minus_do_ce_k_border_sdhr")
  ) %>% 
  gt::tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "tumor_cd3plus_pd1plus_do_ce_k_border_sdhr")
  ) %>% 
  gt::tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr")
  ) %>% 
  gt::tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(
      columns = everything(),
      rows = variable == "stroma_cd3plus_cd8minus_do_ce_k_border_sdhr")
  )
```



### 3. Single Ripley's K feature analysis as interaction term
additionally adjusted for `Tumor CD3 proportion` + `Stroma CD3 proportion`  + `Tumor CD8+/CD8- Ratio (Log)`  + `Stroma CD8+/CD8- Ratio (Log)` + set + refage + stage_cat + histo_grade
```{r echo = TRUE, class.source = 'fold-hide'}
spat_vars <- spatial_data %>% select(
                          contains("tumor") & contains("do_ce_k_border_sdhr"), 
                          contains("stroma") & contains("do_ce_k_border_sdhr"),
                          -ends_with("_5cat"), 
                          -tumor_cd19plus_do_ce_k_border_sdhr,
                          -stroma_cd19plus_do_ce_k_border_sdhr) %>%
  names()
spat_vars2 <- spat_vars %>% str_remove(., "_do_ce_k_border_sdhr")
# Create a list to store the results
cox_mva_fct <- lapply(spat_vars, function(var) {
  
  cell <- str_remove(var, "_do_ce_k_border_sdhr")
  compartment <- str_match(cell, "(tumor|stroma|interface)_*")[,2]
  cell_type <- str_remove(cell, "tumor|stroma")
  cell2 <- paste0(compartment, "_percent", cell_type, "_cells_sdhr", collapse = "")
  
  spatial_data <- spatial_data %>% 
    filter(cell2 >=8)
  # Build the formula dynamically
  formula <- as.formula(paste("Surv(int_to_event_mos, vital_status_fin) ~", 
                              var, "*", cell2,
                              # "+ logit_prop + logit_stroma_prop  + cd8p_cd8n_logratio  + cd8p_cd8n_stroma_logratio",
                              "+ set + refage + stage_cat"))
  
  # Fit the Cox model
  cox_model <- coxph(formula, data = spatial_data)
  
  # Create a regression table for each model
  tbl_regression(cox_model, exponentiate = TRUE#, 
                 # include = c(everything, -var, cell2)
                 ) %>%
    bold_p(t = 0.05) %>%
    add_nevent(location = "level") %>%
    add_n(location = "level") %>%
    bold_labels() %>%
    italicize_levels() %>% 
    remove_row_type(set, type = "all") %>% 
    remove_row_type(refage, type = "all") %>% 
    remove_row_type(stage_cat, type = "all")
})
tbl_stack(cox_mva_fct, group_header = c(spat_vars2)) %>%
  modify_header(label ~ "**MVA for each separate Ripley's K feature additionally adjusted<br>
                `Percent abundance of the corresponding cell type` + <br>
                set + refage + stage_cat + histo_grade**") %>% as_gt() %>% 
  gt::tab_options(row_group.border.top.color = "black",
                  row_group.border.top.width = 3,
                  row_group.border.bottom.color = "darkgrey",
                  row_group.font.size = 15,
                  row_group.font.weight = "bold",
                  row_group.text_transform = "uppercase",
                  row_group.background.color = "grey")
  # gt::tab_style(
  #   style = cell_fill(color = "yellow"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "tumor_cd3plus_cd8minus_do_ce_k_border_sdhr")
  # ) %>% 
  # gt::tab_style(
  #   style = cell_fill(color = "yellow"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "tumor_cd3plus_pd1plus_do_ce_k_border_sdhr")
  # ) %>% 
  # gt::tab_style(
  #   style = cell_fill(color = "yellow"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "stroma_cd3plus_cd8plus_do_ce_k_border_sdhr")
  # ) %>% 
  # gt::tab_style(
  #   style = cell_fill(color = "yellow"),
  #   locations = cells_body(
  #     columns = everything(),
  #     rows = variable == "stroma_cd3plus_cd8minus_do_ce_k_border_sdhr")
  # )
```

# Categorical
```{r}

```

### 3. Single Ripley's K feature analysis as interaction term







































